### 2012-02-10 p.tex
init

\documentclass[preprint]{sigplanconf}

\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{proof}
\usepackage{stmaryrd}
\usepackage{url}
\usepackage{xspace}
\usepackage{multicol}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{proposition}[theorem]{Proposition}
\newenvironment{proof}[1][Proof.]{\begin{trivlist}\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}

\newcommand{\arrow}[1]{\mathtt{#1}}

\begin{document}

\conferenceinfo{ICFP'12}{}
\CopyrightYear{}
\copyrightdata{}
\titlebanner{}
\preprintfooter{}

\title{Continuations via Negative Information Flow}
\authorinfo{Roshan P. James}
           {Indiana University}
           {rpjames@indiana.edu}
\authorinfo{Amr Sabry}
           {Indiana University}
           {sabry@indiana.edu}
\maketitle

\begin{abstract}

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\category{D.3.1}{Formal Definitions and Theory}{}
\category{F.3.2}{Semantics of Programming Languages}{}
\category{F.3.3}{Studies of Program Constructs}{Type structure}

\terms
Languages, Theory

\keywords Arrows, Linear logic, Quantum computing, Reversible logic.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

Blah\cite{infeffects}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}

\acks This project was partially funded by Indiana University's Office of the
Vice President for Research and the Office of the Vice Provost for Research
through its Faculty Research Support Program.  We also acknowledge support
from Indiana University's Institute for Advanced Study.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{small}
\bibliographystyle{abbrvnat}
\bibliography{cites}
\end{small}

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%