### 2012-03-14 iso-int/rc-iso-int.tex
worked on a few pieces

\documentclass{llncs}
\usepackage{url}
\usepackage{proof}
\usepackage{amssymb}
\usepackage{stmaryrd}
\usepackage{listings}
\usepackage{graphicx}

\newcommand{\todo}[1]{\textbf{TODO: #1}}

%subcode-inline{bnf-inline} name langRev
%! swap+ = \mathit{swap}^+
%! swap* = \mathit{swap}^*
%! dagger =  ^{\dagger}
%! assocl+ = \mathit{assocl}^+
%! assocr+ = \mathit{assocr}^+
%! assocl* = \mathit{assocl}^*
%! assocr* = \mathit{assocr}^*
%! identr* = \mathit{uniti}
%! identl* = \mathit{unite}
%! dist = \mathit{distrib}
%! factor = \mathit{factor}
%! (o) = \fatsemi
%! (;) = \fatsemi
%! (*) = \times
%! (+) = +

%subcode-inline{bnf-inline} regex \{\{(((\}[^\}])|[^\}])*)\}\} name main include langRev
%! [^ = \ulcorner
%! ^] = \urcorner
%! [v = \llcorner
%! v] = \lrcorner
%! |-->* = \mapsto^{*}
%! |-->> = \mapsto_{\ggg}
%! |-->let = \mapsto_{let}
%! |--> = \mapsto
%! <--| = \mapsfrom
%! |- = \vdash
%! in = \!\!\in\!\!
%! <=> = \Longleftrightarrow
%! <-> = \leftrightarrow
%! ~> = \leadsto
%! ::= = ::=
%! /= = \neq
%! vi = v_i
%! di = d_i
%! si = s_i
%! sj = s_j
%! F = \texttt{F}
%! T = \texttt{T}
%! forall = \forall
%! exists = \exists
%! empty = \emptyset
%! eta = \eta
%! where = \textbf{where}
%! epsilon = \varepsilon
%! least = \phi
%! trace+ = trace
%! trace* = trace_{\times}
%! loop+ = loop_{+}
%! loop* = loop_{\times}
%! CatC = {\mathcal C}
%! CatA = {\mathcal A}
%! gamma = \gamma
%! {[ = \{
%! ]} = \}
%! elem = \in
%! dagger = ^\dagger
%! alpha = \alpha
%! beta = \beta
%! rho = \rho
%! @@ = \mu
%! @ = \,@\,
%! langRev = \Pi
%! langRevT = \Pi^{o}
%! langRevEE = \Pi^{\eta\epsilon}_{+}
%! bullet = \bullet

\urldef{\mails}\path|{rpjames, sabry}@indiana.edu|

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\title{Isomorphic Interpreters from Logically Reversible Abstract Machines} 
\titlerunning{On the construction of Isomorphic Interpreters} 
\author{Roshan P. James \and Amr Sabry}
\institute{School of Informatics and Computing, Indiana University\\
\mails}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
% Previous work on {{langRevT}} \cite{rc2011,infeffects} studied its
% properties and various constructions in it. However, for a
% computational model, there has been little focus on actually writing
% programs directly in {{langRevT}}. This paper aims to remedy that by
% presenting a systematic technique of expressing logically reversible
% small-step abstract machines in {{langRevT}}. In other words, this
% paper shows that once we have a logically reversible machine in a
% notation of our choice, expressing the machine as an isomorphic
% interpreter in {{langRevT}} can be done systematically and does not
% present any significant conceptual difficulties. This gives us a means
% of developing large {{langRevT}} programs with the ease of reasoning
% at the level of a conventional small-step semantics.  We develop
% several simple interpreters over numbers and addition, move onto tree
% traversal and finish with an interpreter for {{langRevT}} expressed in
% {{langRevT}}.
  In our previous work [RC'11, POPL'12], we developed a reversible
  programming language and established that every computation in it is a
  (partial) isomorphism that is reversible and that preserves
  information. The language is founded on type isomorphisms that have a clear
  categorical semantics but that are awkward as a notation for writing actual
  programs, especially recursive programs. This paper remedies this aspect by
  presenting a systematic technique for developing a large and expressive
  class of reversible recursive programs, that of logically reversible
  small-step abstract machines. In other words, this paper shows that once we
  have a logically reversible machine in a notation of our choice, expressing
  the machine as an isomorphic interpreter can be done systematically and
  does not present any significant conceptual difficulties. Concretely, we
  develop several simple interpreters over numbers and addition, move on to
  tree traversals, and finish with a meta-circular interpreter for our
  reversible language. Generally, this gives us a means of developing large
  reversible programs with the ease of reasoning at the level of a
  conventional small-step semantics.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

Reversible languages

Main points: 
\begin{itemize}
\item Make the case for reversibility via type isomorphisms
\item Type isomorphisms between finite types are straightforward (but
  tedious)
\item Type isomorphisms between recursive types are quite difficult. This is
  what we focus on.
\end{itemize}
To understand the problem in more detail, consider the problem of converting
a reversible recursive function into a partial isomorphism. In general such a
function takes an input, makes several recursive calls, and produces an
answer. To implement this in a reversible way, the structure of the recursion
and the number of recursive calls need to be exposed. The key trick is to
express tease the function into two components: an outer driver that performs
iteration and an inner state transition function that realizes one
``unfolding'' of the recursion. In program transformation world, this is
called, light-weight fission. Using this idea, one can realize any recursive
function as a partial isomorphism. We show a few examples of the construction
culminating with the most challenging recursive function: a full interpreter.

Previous work on {{langRevT}} \cite{rc2011,infeffects} studied its
properties and various constructions in it. However, for a
computational model, there has been little focus on actually writing
programs directly in {{langRevT}}. This paper aims to remedy that by
presenting a systematic technique of expressing logically reversible
small-step abstract machines in {{langRevT}}. This paper shows that
once we have a logically reversible machine in a notation of our
choice, expressing the machine as an isomorphic interpreter in
{{langRevT}} can be done systematically and does not present any
significant conceptual difficulties.

The constructions in this paper apply only to a class of interpreters
that can be represented by logically reversible small-step operational
semantics. While, in a formal sense, they are related to Abramsky's
``bi-orthogonal automata''\cite{abramsky2005structural}, we will try
to give them a more direct characterization purely in terms of logical
reversibility of the small-step semantics.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Review of the Reversible Language: {{langRevT}} }
\label{sec:pi}

We briefly review the reversible language {{langRevT}} introduced in our
previous work~\cite{rc2011,infeffects}. The set of types includes the empty
type {{0}}, the unit type {{1}}, sum types {{b1+b2}}, product types
{{b1*b2}}, and recursive types {{@@x.b}}. The set of values {{v}} includes
{{()}} which is the only value of type {{1}}, {{left v}} and {{right v}}
which inject {{v}} into a sum type, {{(v1,v2)}} which builds a value of
product type, and {{<v>}} which is used to construct recursive values. There
are no values of type {{0}}:
%subcode{bnf} include main
% value types, b ::= 0 | 1 | b + b | b * b | x | @@x.b
% values, v ::= () | left v | right v | (v, v) | <v>

The expressions of {{langRevT}} are witnesses to the following type
isomorphisms:
%subcode{bnf} include main
%! columnStyle = r@{\hspace{-0.5pt}}c@{\hspace{-0.5pt}}l
%zeroe :&  0 + b <-> b &: zeroi
%swap+ :&  b1 + b2 <-> b2 + b1 &: swap+
%assocl+ :&  b1 + (b2 + b3) <-> (b1 + b2) + b3 &: assocr+
%identl* :&  1 * b <-> b &: identr*
%swap* :&  b1 * b2 <-> b2 * b1 &: swap*
%assocl* :&  b1 * (b2 * b3) <-> (b1 * b2) * b3 &: assocr*
%dist0 :& 0 * b <-> 0 &: factor0
%dist :&~ (b1 + b2) * b3 <-> (b1 * b3) + (b2 * b3)~ &: factor
%fold :&~ b[@@x.b/x] <-> @@x.b ~&: unfold
Each line of the above table introduces one or two combinators that witness
the isomorphism in the middle. Collectively the isomorphisms state that the
structure {{(b,+,0,*,1)}} is a \emph{commutative semiring}, i.e., that each
of {{(b,+,0)}} and {{(b,*,1)}} is a commutative monoid and that
multiplication distributes over addition. The last isomorphism witnesses the
equivalence of a value of recursive type with all its ``unrollings.'' The
isomorphisms are extended to form a congruence relation by adding the
following constructors that witness equivalence and compatible closure. In
addition, the language includes a {{trace}} operator to express looping:
%subcode{proof} include main
%@  ~
%@@ id : b <-> b 
%
%@ c : b1 <-> b2
%@@ sym c : b2 <-> b1
%
%@ c1 : b1 <-> b2
%@ c2 : b2 <-> b3
%@@ c1(;)c2 : b1 <-> b3
%---
%@ c1 : b1 <-> b3
%@ c2 : b2 <-> b4
%@@ c1 (+) c2 : b1 + b2 <-> b3 + b4
%
%@ c1 : b1 <-> b3
%@ c2 : b2 <-> b4
%@@ c1 (*) c2 : b1 * b2 <-> b3 * b4
%
%@ c : b1 + b2 <-> b1 + b3
%@@ trace c : b2 <-> b3

\paragraph*{Adjoint.} 
An important property of the language is that every combinator {{c}} has an
adjoint {{c{dagger}}} that reverses the action of {{c}}. This is evident by
construction for the primitive isomorphisms. For the closure combinators, the
adjoint is homomorphic except for the case of sequencing in which the order
is reversed, i.e., {{(c1 (;) c2){dagger} = (c2{dagger}) (;) (c1{dagger}) }}.

\paragraph*{Graphical Language.}
The syntactic notation above is often obscure and hard to read.  Following
the tradition established for monoidal
categories~\cite{springerlink:10.1007/978-3-642-12821-94}, we present a
graphical language that conveys the semantics of the language.  The general
idea of the graphical notation is that combinators are modeled by ``wiring
diagrams'' or ``circuits'' and that values are modeled as ``particles'' or
``waves'' that may appear on the wires. Evaluation therefore is modeled by
the flow of waves and particles along the wires.

\begin{itemize}
\item The simplest sort of diagram is the {{id : b <-> b}} combinator which
  is simply represented as a wire labeled by its type {{b}}, as shown on the
  left. In more complex diagrams, if the type of a wire is obvious from the
  context, it may be omitted. When tracing a computation, one might imagine a
  value {{v}} of type {{b}} on the wire, as shown on the right.

  \begin{multicols}{2}
\begin{center}
\scalebox{0.95}{
%%subcode-line{pdfimage}[diagrams/thesis/b-wire.pdf]
\includegraphics{diagrams/thesis/b-wire.pdf}
}
\end{center}
\begin{center}
\scalebox{0.95}{
%%subcode-line{pdfimage}[diagrams/thesis/b-wire.pdf]
\includegraphics{diagrams/thesis/b-wire-value.pdf}
}
\end{center}
  \end{multicols}

\item The product type {{b1*b2}} may be represented using either one wire
  labeled {{b1*b2}} or two parallel wires labeled {{b1}} and {{b2}}. In the
  case of products represented by a pair of wires, when tracing execution
  using particles, one should think of one particle on each wire or
  alternatively as in folklore in the literature on monoidal categories as a
  ``wave.''
\begin{multicols}{2}
\begin{center}
\scalebox{0.95}{
%%subcode-line{pdfimage}[diagrams/thesis/pair-one-wire.pdf]
\includegraphics{diagrams/thesis/product-one-wire.pdf}
}
\end{center}
\begin{center}
\scalebox{0.95}{
\includegraphics{diagrams/thesis/product-one-wire-value.pdf}
}
\end{center}
\end{multicols}
\begin{multicols}{2}
\begin{center}
\scalebox{0.95}{
%%%subcode-line{pdfimage}[diagrams/thesis/pair-of-wires.pdf]
\includegraphics{diagrams/thesis/product-two-wires.pdf}
}
\end{center}
\begin{center}
\scalebox{0.95}{
\includegraphics{diagrams/thesis/product-two-wires-value.pdf}
}
\end{center}
\end{multicols}

\item Sum types may similarly be represented by one wire or using
  parallel wires with a {{+}} operator between them. When tracing the
  execution of two additive wires, a value can reside on only one of the two
  wires.
\begin{multicols}{2}
\begin{center}
\scalebox{0.95}{
%%subcode-line{pdfimage}[diagrams/thesis/sum-one-wire.pdf]
\includegraphics{diagrams/thesis/sum-one-wire.pdf}
}
\end{center}
\begin{center}
\scalebox{0.95}{
\includegraphics{diagrams/thesis/sum-two-wires-left-value.pdf}
}
\end{center}
\end{multicols}
\begin{multicols}{2}
\begin{center}
\scalebox{0.95}{
%%subcode-line{pdfimage}[diagrams/thesis/sum-of-wires.pdf]
\includegraphics{diagrams/thesis/sum-two-wires.pdf}
}
\end{center}
\begin{center}
\scalebox{0.95}{
\includegraphics{diagrams/thesis/sum-two-wires-right-value.pdf}
}
\end{center}
\end{multicols}

%% \item
%% When representing complex types like {{(b1*b2)+b3}} some visual
%% grouping of the wires may be done to aid readability. The exact type
%% however will always be clarified by the context of the diagram.

%% \begin{center}
%% \scalebox{0.95}{
%% %subcode-line{pdfimage}[diagrams/thesis/complex-type-crop.pdf]
%% }
%% \end{center}

\item Associativity is implicit in the graphical language. Three parallel
  wires represent {{b1*(b2*b3)}} or {{(b1*b2)*b3}}, based on the context.
\begin{center}
\scalebox{0.95}{
%%subcode-line{pdfimage}[diagrams/thesis/associate.pdf]
\includegraphics{diagrams/thesis/assoc.pdf}
}
\end{center}

\item Commutativity is represented by crisscrossing wires.
\begin{multicols}{2}
\begin{center}
\scalebox{0.95}{
%%subcode-line{pdfimage}[diagrams/thesis/swap-pair.pdf]
\includegraphics{diagrams/thesis/swap_times.pdf}
}
\end{center}
\begin{center}
\scalebox{0.95}{
%%subcode-line{pdfimage}[diagrams/thesis/swap-sum.pdf]
\includegraphics{diagrams/thesis/swap_plus.pdf}
}
\end{center}
\end{multicols}

By visually tracking the flow of particles on the wires, one can
verify that the expected types for commutativity are satisfied.

\begin{multicols}{2}
\begin{center}
\scalebox{0.95}{
\includegraphics{diagrams/thesis/swap_times_value.pdf}
}
\end{center}
\begin{center}
\scalebox{0.95}{
\includegraphics{diagrams/thesis/swap_plus_value.pdf}
}
\end{center}
\end{multicols}

\item The morphisms that witness that {{0}} and {{1}} are the additive and
  multiplicative units are represented as shown below. Note that since there
  is no value of type 0, there can be no particle on a wire of type {{0}}.
  Also since the monoidal units can be freely introduced and eliminated, in
  many diagrams they are omitted and dealt with explicitly only when they are
  of special interest.
\begin{multicols}{2}
\begin{center}
\scalebox{0.95}{
%%subcode-line{pdfimage}[diagrams/thesis/identr1.pdf]
\includegraphics{diagrams/thesis/uniti.pdf}
}
\end{center}
\begin{center}
\scalebox{0.95}{
%%subcode-line{pdfimage}[diagrams/thesis/identl1.pdf]
\includegraphics{diagrams/thesis/unite.pdf}
}
\end{center}  
\end{multicols}
\begin{multicols}{2}
\begin{center}
\scalebox{0.95}{
%%subcode-line{pdfimage}[diagrams/thesis/identr0.pdf]
\includegraphics{diagrams/thesis/zeroi.pdf}
}
\end{center}
\columnbreak
\begin{center}
\scalebox{0.95}{
%%subcode-line{pdfimage}[diagrams/thesis/identl0.pdf]
\includegraphics{diagrams/thesis/zeroe.pdf}
}
\end{center}
\end{multicols}

\item Distributivity and factoring are represented using the dual
  boxes shown below:
\begin{multicols}{2}
\begin{center}
  \includegraphics{diagrams/thesis/dist.pdf}
\end{center}
\begin{center}
  \includegraphics{diagrams/thesis/factor.pdf}
\end{center}
\end{multicols}

Distributivity and factoring are interesting because they represent
interactions between sum and pair types. Distributivity should
essentially be thought of as a multiplexer that redirects the flow of
{{v:b}} depending on what value inhabits the type {{b1+b2}}, as shown
below. Factoring is the corresponding adjoint operation.

\begin{multicols}{2}
\begin{center}
  \includegraphics{diagrams/thesis/dist-wire-value1.pdf}
\end{center}
\begin{center}
  \includegraphics{diagrams/thesis/dist-wire-value2.pdf}
\end{center}
\end{multicols}

\item \todo{fold, unfold, and trace}
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Simple Bounded Number Iteration}

We illustrate the main concepts and constructions using two simple examples
that essentially count $n$ steps. The first machine does nothing else; the
second uses this counting ability to add two numbers.

%%%%%%%%%%%%%%%%%%%
\subsection{Counting}

The first machine is defined as follows:

\begin{multicols}{2}
%subcode{bnf} include main
% Numbers, n, m = 0 | n + 1
% Machine states = <n, n>

%subcode{bnf} include main
% Start state = <n, 0>
% Stop State = <0, n>
\end{multicols}
% Small-step operational semantics:

%subcode{opsem} include main
% <n+1, m> |--> <n, m+1>

\noindent The machine is started with a number {{n}} in the first position
and {{0}} in the second. Each reduction step decrements the first number and
increments the second. The machine stops when the first number reaches 0,
thereby taking exactly {{n}} steps. For example, if the machine is started in
the configuration {{<3,0>}}, it would take exactly 3 steps to reach the final
configuration: {{ <3,0> |--> <2,1> |--> <1,2> |--> <0,3> }}. Dually, since
the machine transition is clearly reversible, the machine can execute
backwards from the final configuration to reach the initial configuration in
also 3 steps.

Our goal is to implement this abstract machine in {{langRevT}}, i.e., our
goal is to write a combinator {{c : nat * nat <-> nat * nat}} where {{nat}}
is an abbreviation for the recursive type {{@@x.1+x}}. The combinator should
map {{(n,0)}} to {{(0,n)}} by precisely mimicking the steps of the abstract
machine. Let us analyze this desired combinator {{c}} in detail starting from
its interface:

\begin{center}
  \includegraphics{diagrams/nat-nat1.pdf}
\end{center}

The reduction step of the machine examines the first {{nat}} and reconstructs
the second {{nat}}. In other words, the first {{nat}} is unfolded to examine
its structure and the second {{nat}} is folded to reconstruct it:

\begin{center}
  \includegraphics{diagrams/nat-nat3.pdf}
\end{center}

The triangle on the left denotes the {{unfold}} isomorphism, which
works as follows: If the number {{n}} is zero, the output is the top
branch which has type {{1}}. If the number was non-zero, the output is
on the bottom branch and has value {{n-1}}. The triangle on the right
is the dual {{fold}} isomorphism.  Let us examine all the possible
machine states by introducing distribution operations and any required
swaps:

%% by distributing on the
%% input. Dually let us do the same with the output, but this time also
%% be a little bit explicit about the swap operations required so the
%% right types are in place.

\begin{center}
  \includegraphics{diagrams/nat-nat4.pdf}
\end{center}

The construction above already gives us some insight into how we can
implement our machine's one reduction step. We want the output {{n'}} to be
{{n+1}} where~{{n}} is the first component of the input and similarly the
output {{m'}} to be {{m+1}} where {{m}} is the second component of the input.
We can express these equalities by connecting wires, i.e., by connecting the
{{n-1}} wire to {{n'}} and the {{m}} wire to {{m'-1}}:

\begin{center}
  \includegraphics{diagrams/nat-nat5.pdf}
\end{center}

%% We are a few steps away from an interpreter at this point and the
%% final steps rely on the answers to a few questions.

%% \begin{enumerate}
%% \item How do relate the outputs of the combinator to the inputs, such
%%   that the machine correctly takes multiple steps?

%% \item

%%  What do we do about the two inner branches?

%% \end{enumerate}

The branch labeled {{((), m)}} corresponds to the machine state {{<0,m>}} and
hence looks like the stop state of the machine. Similarly the branch labeled
{{((), n')}} corresponds to the start state of the machine. We are close to
the completion of the interpreter at this point. The last step in the
construction involves introducing a {{trace}] operation for iteration of
reduction steps and exposing the start and stop states. We do this by
``sliding'' the two sections of the diagram as shown below:

\vspace{-20px}
\begin{multicols}{2}
\begin{center}
\scalebox{0.5}{
  \includegraphics{diagrams/nat-nat6.pdf}
}
\end{center}
\columnbreak
\begin{center}
\scalebox{0.6}{
  \includegraphics{diagrams/nat-nat7.pdf}
}
\end{center}  
\end{multicols}

The introduced {{trace}} feeds back related machine states. The
exposed start state essentially requires the initial value of {{n}} to
start the machine.

\begin{center}
\scalebox{1.0}{ \includegraphics{diagrams/nat-nat8.pdf} }
\end{center}

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \section{Abstracting out the Construction}

% The construction of the isomorphic interpreter essentially exploits
% the structure of most abstract machines -- the LHS of the transition
% rules determines a machine state by pattern machine. The RHS does the
% dual namely term reconstruction. The fact that the machine is
% logically reversible ensures that the computation performed by each
% reduction is essentially captured by an isomorphism.

% Consider a machine with the following two reduction steps
% {{<n+1,m>|--> <n, m+1>}} and {{<n+1,m>|--><n,m>}}. Can we create an
% isomorphic interpreter for such a machine? We can't, because for any
% given machine state, it is not obvious what reduction rule to
% apply. We could reverse this machine if there was a clear
% non-ambiguity in which machine rule was chosen.

% Consider a machine, with the following rewrite rules
% {{<n+1,m,p>|--><n,m,p+1>}} and {{<0,m+1,p>|--><0,m,p+1>}}. Can we
% create an isomorphic interpreter for this machine? It turns out we
% can't because for any given output state {{<0,m,p+1>}} it is not
% obvious which input state it came from - i.e. there are two
% possibilities {{<1,m,p>}} or there is {{<0,m+1,p>}}.

% Consider a machine that has the addition rule {{<''x+y''>|--><''z''>}}
% where {{z}} is the sum of {{x}} and {{y}}. This machine again cannot
% be turned into an isomorphic interpreter because the one reduction
% rule does not have a logically inverse -- in other words, given a
% {{z}}, we are unable to determine the original {{x}} and {{y}}.

% If we had only the reduction relation {{<n+1> |--> <n>}}, can we
% construct an isomorphic interpreter? After all, every rewrite rule is
% logically reversible. In fact, this is not possible because there is
% no distinguished start state for such an interpreter and hence one
% does not know when to stop during reverse execution.

% \paragraph*{Conditions that must be satisfied for constructing a isomorphic
% interpreter}:
% \begin{enumerate}
% \item Distinguishable start and stop states. There may be more than one
%   valid start state and more than one valid stop state.
% \item Each valid machine state must match a unique reduction LHS. Each
%   valid machine state must machine a unique reduction RHS.
% \item Every reduction step must be computable (in {{langRevT}}) and
%   must be logically reversible -- i.e. it must be possible to reduce
%   from right to left.
% \end{enumerate}

% \paragraph*{High level steps in the translation:}
% \label{sec:steps}
% \begin{enumerate}
% \item Expand the input such that every input term that the LHS of the
%   interpreter pattern matches on is exposed.
% \item Expand the output such that every output term that it
%   reconstructed by the interpreter is exposed.
% \item Shuffle matching LHS term to matching RHS terms, inserting any
%   appropriate mediating computations. In the previous example, this
%   was only a swap operation.
% \item Flip the construction inside out introducing a {{trace}} and
%   exposing the start and stop states. 
% \end{enumerate}

\subsection{Adder}

Although trivial, the previous example captures the fundamental steps of our
construction. For reference, these steps are:
\begin{enumerate}
\item Expand the input to expose enough structure to distinguish the left
  hand side of each machine state; 
\item Expand the output to expose enough structure to distinguish the right
  hand side of each machine state;
\item Shuffle matching input terms to output terms, inserting any appropriate
  mediating computations. 
\item Slide the two sections to expose the start and stop state and introduce
  a {{trace}} to iterate the construction.
\end{enumerate}

Let us apply these steps to the slightly more interesting example of an
adder:
\begin{multicols}{2}  
%subcode{bnf} include main
% Numbers, n, m,p = 0 | n + 1
% Machine states = <e, e, e>

%subcode{bnf} include main
% Start state = <n, n, 0>
% Stop State = <0, n, n>
\end{multicols}
%subcode{opsem} include main
% <n+1, p, m> |--> <n, p+1, m+1>

The idea of the machine is to start with 3 numbers: the two numbers to add
and an accumulator initialized to 0. Each step of the machine, decrements one
of the numbers and increments the second number and the accumulator. For
example, {{<3,4,0> |--> <2,5,1> |--> <1,6,2> |--> <0,7,3>}}. In general, the
sum of the two numbers will be in the second component, and the last
component is supposed to record enough information to make the machine
reversible.

A closer look however reveals that the machine only defines a \emph{partial}
isomorphism: not all valid final states can be mapped to valid start
states. Indeed consider the configuration {{<0,2,3>}} which is a valid final
state. Going backwards, the transitions start as follows
{{<0,2,3> |--> <1,1,2> |--> <2,0,1>}} at which point, the machine 
gets stuck at a state that is not a valid start state. 
There are several ways to make this machine reversible. The simplest is to
add another possible kind of start state {{<n,0,n>}}. Our construction now
proceeds as expected. After expanding the input to expose $n$ and the output
to expose both $p$ and $m$, we reach the following configuration:

% \begin{center}
%   \includegraphics{diagrams/adder1.pdf}
% \end{center}

% The equalities we would like to express at this point are {{n-1=n'}},
% {{m=m'-1}} and that {{p+1=p'}}. Fortunately {{langRevT}} provides the
% {{add1: nat <-> nat}} combinator and so we could just use that to
% mediate between {{p}} and {{p'}}. This is one solution and this gives
% us:

% \begin{center}
%   \includegraphics{diagrams/adder2.pdf}
% \end{center}

% This can be inverted as before, to give us an interpreter. While this
% is a correct implementation of the small-step semantics, it is
% somewhat unsatisfying for the reason that it does not see very
% systematic. After all we chose to use an {{add1}} for {{p}} -- why
% didn't we do the same for {{m}}? More importantly, the fact that we
% chose used {{add1}} is also hiding something -- namely that {{add1}}
% is a partial isomorphism. Its inverse is not defined when applied to
% {{0}} i.e. it will go into an infinite loop.

% In a sense this implicit asymmetry of the abstract machine was being
% hidden by the {{add1}} operation. So instead of choosing to use
% {{add1}} let us chose to expand out {{p'}}, as recommended by step 2
% in section \ref{sec:steps}. This gives us the following: 
%
% We can very that the connected wires do indeed give us the desired
% equalities between the inputs and outputs. Also we recognize our
% expected start state and expected stop state. What is the dangling
% state?

% The dangling state was what has hidden by the use of {{add1}} and
% denotes the fact that the abstract machine did not define any valid
% backward reduction for the following valid machine state
% {{<n,0,m+1>}}. To take a backward step, both {{m}} and {{p}} have to
% be simultaneously decremented and this is not possible if {{p}} is
% {{0}}.  At this point we have two choices:

\begin{center}
\scalebox{0.9}{
  \includegraphics{diagrams/adder3.pdf}
}
\end{center}

We have also connected the matching input and output wires. At this point, we
have exposed both start states and the stop state. 
% \begin{enumerate}
% \item We can hide the use the combinator {{omega : b <-> 0}}
%   expressible in {{langRevT}}. This combinator is basically an
%   infinite loop on all inputs and hence would allow us to entirely
%   discard one possible branch. In other words, we are explicitly
%   admitting that our machine is undefined on some inputs.
% \item We can admit that there are two possible start states to this
%   abstract machine {{<n, m, 0>}} and {{<n,0,m>}}. One of this is an
%   artifact of the fact that backward execution encounters stuck
%   states. Dually, if there were stuck states for forward execution,
%   they would be reflected in the fact that there were multiple stop
%   states -- only one of which denotes a desirable or correct execution
%   of the machine.
% \end{enumerate}
We can slide the two sides of the diagram and tie the knot using {{trace}} to
get the desired combinator:

\begin{center}
\scalebox{0.9}{
  \includegraphics{diagrams/adder4.pdf}
}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Advanced Examples}

We show the generality of our construction by applying it to two non-trivial
examples: a tree traversal and a meta-circulator interpreter for
{{langRevT}}.

%%%%%%%%%%%%%%%%%%
\subsection{Tree Traversal}

The type of binary trees we use is {{@@x.nat+x*x}}, i.e., binary trees with
no information at the nodes and with natural numbers at the leaves. To define
the abstract machine, we need a notion of \emph{tree contexts} to track which
subtree is currently being explored. The definitions are:

%subcode{bnf} include main
% Tree, t = L n | N t t
% Tree Contexts, c = Lft c t | Rgt t c | []
%
% Machine states = <t, c> | {[c, t]}
% Start state = <t, []>
% Stop State = {[[], t]}

Our example machine below traverses a given tree and increments every leaf
value:

%subcode{opsem} include main
% <L n, c> &|-->& {[c, L (n+1)]}
% <N t1 t2, c> &|-->& <t1, Lft c t2>
% {[Lft c t2, t1]} &|-->& <t2, Rgt t1 c>
% {[Rgt t1 c, t2]} &|-->& {[N t1 t2, c]}

The machine here is a little richer than the ones dealt with
previously. In particular, we have a two types of machine states --
{{<t,c>}} and {{ {[t, c]} }}. The first of these corresponds to walking
down a tree, building up the context in the process. The second
corresponds to reconstructing the tree from the context and also
switching focus to any unexplored subtrees in the process. There are
also two syntactic categories to deal with (trees and tree contexts)
where previously we only had numbers. The {{fold}} and {{unfold}}
isomorphisms that we need for trees and tree contexts are:

%subcode{bnf} include main
%! columnStyle = rrcll
% unfold_t :& ~~~~t &<->& n + t * t &: fold_t
% unfold_c :& c &<->& 1 + c * t + t * c ~~~~~~&: fold_c

\paragraph*{New Notation.}
To make the diagrams easier to understand, we introduce a syntactic
convenience which combines the first steps of the construction that consist
of {{fold}} / {{unfold}} and {{distribute}} / {{factor}}. We collectively
represent these steps using thin vertical rectangle. Also we will introduce
the convention that the component that is being expanded or collapsed will be
marked by using {{[^ ^]}} and the components that are being generated will be
marked by {{[v v]}}. For example, given a value of type {{tree*nat}}, the
diagram below shows how to first expand the {{tree}} component and then in
one of the generated branches, expand the {{nat}} component:

\begin{center}
\scalebox{1.0}{
  \includegraphics{diagrams/syntax1.pdf}
}
\end{center}

We can now apply our construction. The first step to developing the
isomorphic interpreter is to recognize that the two possible kinds of machine
states simply hide an implicit {{bool}}. We make this explicit:

%subcode{opsem} include main
%! columnStyle = rclr
% <T, L n, c> &|-->& <F, L (n+1), c> &~~~~~ rule 1
% <T, N t1 t2, c> &|-->& <T, t1, Lft c t2> & rule 2 
% <F, t1, Lft c t2> &|-->& <T, t2, Rgt t1 c> & rule 3
% <F, t2, Rgt t1 c> &|-->& <F, c, N t1 t2> & rule 4

And so machine states have the form:

%subcode{bnf} include main
% Machine states = <bool, t, c>
% Start state = <T, t, []>
% Stop state =  <F, t, []>

And we start examining the machine components as before:

\begin{center}
\scalebox{0.9}{
  \includegraphics{diagrams/tree1.pdf}
}
\end{center}

On the input side, for {{T}} states, we have expanded the tree component and
for {{F}} states we have expanded the tree contexts. Similarly we have done
the opposite on the output side exactly matching up what the abstract machine
does. One thing to note is that we dropped the {{1}} introduced by expanding
the {{bool}} and instead just labeled the {{true}} and {{false}} branches.

We are ready to start connecting the machine states corresponding to
the reductions that we would like:
\begin{enumerate}
\item When we encounter a leaf we would like to increment its value and move
  to the corresponding {{F}} machine state. For the sake of simplicity in
  this interpreter we won't be concerned with exposing stuck states on
  reverse execution caused by {{sub1 0}}. So we are free to use the {{add1}}
  combinator.

\item For all the other reduction rules, it is a straightforward mapping of
  related states following the reduction rules.

\end{enumerate}

The required reduction rules have been labeled on the construction
belong. We have also included subscripts on the various {{t}}s to
indicate any swaps that may need to be inserted.

\begin{center}
\scalebox{0.9}{
  \includegraphics{diagrams/tree2.pdf}
}
\end{center}

This essentially completes the construction of the (partially)
isomorphic tree-traversal interpreter. (Partially because of the
implicit non-termination introduced by the {{add1}}).

%%%%%%%%%%%%%%%%%%%%%
\subsection{A {{langRevT}} Interpreter}

%subcode{bnf} include main
% Combinators, c = iso | c (;) c | c (*) c | c (+) c | trace c
% Combinator Contexts, cc = [] | Fst cc c | Snd c cc 
%                  &|& LeftTimes cc c v | RightTimes c v cc 
%                  &|& LeftPlus cc c | RightPlus c cc | Trace cc
% Values, v = () | (v, v) | L v | R v
%
% Machine states = <c, v, cc> | {[c, v, cc]}
% Start state = <c, v, []> 
% Stop State = {[c, v, []]}

%subcode{opsem} include main
%! columnStyle = rclr
% <iso, v, cc> &|-->& {[iso, iso(v), cc]} &~~~~~~~~~~~~ rule 1
% <c1(;)c2, v, cc> &|-->& <c1, v, Fst cc c2> &~~~~~~~~~~~~ rule 2
% {[c1, v, Fst cc c2]} &|-->& <c2, v, Snd c1 cc> &~~~~~~~~~~~~ rule 3
% {[c2, v, Snd c1 cc]} &|-->& {[ c1(;)c2, v, cc ]} &~~~~~~~~~~~~ rule 4
% <c1(+)c2, L v, cc> &|-->& <c1, v, LeftPlus cc c2> &~~~~~~~~~~~~ rule 5
% {[ c1, v, LeftPlus cc c2 ]} &|-->& {[c1 (+) c2, L v, cc ]} &~~~~~~~~~~~~ rule 6
% <c1(+)c2, R v, cc> &|-->& <c2, v, RightPlus c1 cc> &~~~~~~~~~~~~ rule 7
% {[ c2, v, RightPlus c1 cc ]} &|-->& {[c1 (+) c2, R v, cc ]} &~~~~~~~~~~~~ rule 8
% <c1(*)c2, (v1, v2), cc> &|-->& <c1, v1, LeftTimes cc c2 v2> &~~~~~~~~~~~~ rule 9
% {[ c1, v1, LeftTimes cc c2 v2 ]} &|-->& <c2, v2, RightTimes c1 v1 cc> &~~~~~~~~~~~~ rule 10
% {[ c2, v2, RightTimes c1 v1 cc ]} &|-->& {[ c1 (*) c2, (v1, v2), cc ]} &~~~~~~~~~~~~ rule 11
% < trace c, v, cc> &|-->& <c, R v, Trace cc> &~~~~~~~~~~~~ rule 12
% {[c, L v, Trace cc]} &|-->& <c, L v, Trace cc> &~~~~~~~~~~~~ rule 13
% {[c, R v, Trace cc]} &|-->& {[trace c, R v, cc]} &~~~~~~~~~~~~ rule 14

\begin{center}
\scalebox{0.9}{
  \includegraphics{diagrams/pi0-1.pdf}
}
\end{center}

\begin{center}
\scalebox{0.9}{
  \includegraphics{diagrams/pi0-2.pdf}
}
\end{center}

% TODO : The real difficulty with this interpreter is to argue why it is
% not full of undefined states. For instance, why the machines will
% never reach stuck states such as {{<c1(+)c2, (), ..>}}.  The right way
% out of that is to make the type system of the abstract machine rich
% enough that the structure of the abstract automatically rules out this
% sort of thing.

% %subcode{bnf} include main
% % Combinators, c(t1<->t2) = 
% %             &|& c(t<->t) = iso 
% %             &|& c(t1<->t2) = c(t1<->t3) (;) c(t3<->t2) 
% %             &|& c(t1*t2<->t3*t4) = c(t1<->t3) (*) c(t2<->t4) 
% %             &|& c(t1+t2<->t3+t4) = c(t1<->t3) (+) c(t2<->t4)
% %
% % Combinator Contexts, cc = [] | Fst cc c | Snd c cc 
% %                  &|& LeftTimes cc c v | RightTimes c v cc 
% %                  &|& LeftPlus cc c | RightPlus c cc 
% %
% % Type Indexed Values, v(t) =
% %  &|& v(1) = () 
% %  &|& v(t1*t2) = (v(t1), v(t2)) 
% %  &|& v(t1+t2) = L v(t1) | R v(t2)
% %
% % Machine states = <c, v, cc> | {[c, v, cc]}
% % Start state = <c, v, []> 
% % Stop State = {[c, v, []]}

% However this begs the question - does {{langRevT}} need all these GADT
% types to reason about all these things? Something to think about. If
% we do, then there are two ways out

% (1) add GADTs to {{langRevT}} and keep the structure of things
% simple. They can always be added opaquely for any given types -- we
% only need to add the proper fold/unfold morphisms for those types
% (maybe).

% (2) compile down to {{langRevT}} as it is. In this case, {{langRevT}}
% is indeed a relatively less typed language than the source language
% and will hence have to treat as stuck-states the non-type correct
% machine states. We could apply omega to them and make them vanish --
% for example. This is similar to the handling of the {{sub1}} in the
% previous example.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{splncs03} 
\bibliography{cites}
\end{document}