### 2012-10-24 r.tex

\documentclass[preprint]{sigplanconf}

\usepackage{graphicx}
\usepackage{comment}
\usepackage{amsmath}
\usepackage{xspace}
\usepackage{amssymb}
\usepackage{stmaryrd}
\usepackage{proof}
\usepackage{multicol}
\usepackage[nodayofweek]{datetime}
\usepackage{etex}
\usepackage[all, cmtip]{xy}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{multicol}
\newcommand\hmmax{0} % default \newcommand\bmmax{0} % default 4
\usepackage{bm}
\usepackage{cmll}

\newcommand{\vecspace}[1]{\mathbb{V}_{#1}}
\renewcommand{\vec}[2]{\begin{bmatrix} #1 \\ #2 \end{bmatrix}}
\newcommand{\Fpx}[1]{\mathbb{F}_{#1}}
\newcommand{\modulusp}[1]{\left | #1 \right |}
\newcommand{\fname}[1]{\ulcorner #1 \urcorner}
\newcommand{\fconame}[1]{\llcorner #1 \lrcorner}

\newcommand{\xcomment}[2]{\textbf{#1:~\textsl{#2}}}
\newcommand{\amr}[1]{\xcomment{Amr}{#1}}
\newcommand{\roshan}[1]{\xcomment{Roshan}{#1}}
\newcommand{\jacques}[1]{\xcomment{Jacques}{#1}}

\newcommand{\asterix}[0]{*}

\newcommand{\ie}{\textit{i.e.}\xspace}
\newcommand{\eg}{\textit{e.g.}\xspace}

\newcommand{\lcal}{\ensuremath{\lambda}-calculus\xspace}
\newcommand{\G}{\ensuremath{\mathcal{G}}\xspace}

\newcommand{\code}[1]{\lstinline[basicstyle=\small]{#1}\xspace}
\newcommand{\name}[1]{\code{#1}}

\def\newblock{}

\newenvironment{floatrule}
    {\hrule width \hsize height .33pt \vspace{.5pc}}
    {\par\addvspace{.5pc}}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{proposition}[theorem]{Proposition}
\newenvironment{proof}[1][Proof.]{\begin{trivlist}\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}

\newcommand{\arrow}[1]{\mathtt{#1}}

\newcommand{\dgm}[2][0.95]{
\begin{center}
\scalebox{#1}{
\includegraphics{diagrams/#2.pdf}
}
\end{center}
}

%subcode-inline{bnf-inline} name langRev
%! swap+ = \mathit{swap}^+
%! swap* = \mathit{swap}^*
%! dagger =  ^{\dagger}
%! assocl+ = \mathit{assocl}^+
%! assocr+ = \mathit{assocr}^+
%! assocl* = \mathit{assocl}^*
%! assocr* = \mathit{assocr}^*
%! identr* = \mathit{uniti}
%! identl* = \mathit{unite}
%! dist = \mathit{distrib}
%! factor = \mathit{factor}
%! eta = \eta
%! eps = \epsilon
%! eta+ = \eta^+
%! eps+ = \epsilon^+
%! eta* = \eta^{\times}
%! eps* = \epsilon^{\times}
%! trace+ = trace^+
%! trace* = trace^{\times}
%! (o) = \circ
%! (;) = \fatsemi
%! (*) = \times
%! (+) = +
%! LeftP = L^+
%! RightP = R^+
%! LeftT = L^{\times}
%! RightT = R^{\times}
%! alpha = \alpha
%! bool = \textit{bool}
%! color = \textit{color}
%! Gr = G

%subcode-inline{bnf-inline} regex \{\{(((\}[^\}])|[^\}])*)\}\} name main include langRev
%! Gx = \Gamma^{\times}
%! G = \Gamma
%! [] = \Box
%! |-->* = \mapsto^{\asterix}
%! |-->> = \mapsto_{\ggg}
%! |--> = \mapsto
%! <--| = \mapsfrom
%! |- = \vdash
%! <><> = \approx
%! ==> = \Longrightarrow
%! <== = \Longleftarrow
%! <=> = \Longleftrightarrow
%! <-> = \leftrightarrow
%! ~> = \leadsto
%! -o+ = \multimap^{+}
%! -o* = \multimap^{\times}
%! -o = \multimap
%! ::= = &::=&
%! /= = \neq
%! @@ = \mu
%! [^ = \lceil
%! ^] = \rceil
%! {| = \{
%! |} = \}
%! bot = \bot
%! bot1 = \bot_1
%! dots = \ldots
%! ^^^ = ^{\dagger}
%! elem = \in
%! forall = \forall
%! exists = \exists
%! empty = \epsilon
%! Pi = \Pi
%! Pi0 = \Pi^{o}
%! PiEE* = \Pi^{\eta\epsilon}_{*}
%! PiEE+ = \Pi^{\eta\epsilon}_{+}
%! PiEE = \Pi^{\eta\epsilon}
%! CatSet = \textbf{Set}
%! theseus = Theseus
%! sqrt(x) = \sqrt{#x}
%! surd(p,x) = \sqrt[#p]{#x}
%! inv(x) = \frac{1}{#x}
%! frac(x,y) = \frac{#x}{#y}
%! * = \times

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\conferenceinfo{ICFP'13}{}
\CopyrightYear{}
\copyrightdata{}
\titlebanner{}
\preprintfooter{}

\title{Negative and Fractional Types}

\authorinfo{Roshan P. James}
           {Indiana University}
           {rpjames@indiana.edu}
\authorinfo{Zachary Sparks} 
           {Indiana University}
           {zasparks@indiana.edu}
\authorinfo{Jacques Carette} 
           {McMaster University}
           {carette@mcmaster.ca}
\authorinfo{Amr Sabry}
           {Indiana University}
           {sabry@indiana.edu}

\maketitle

\begin{abstract}

\end{abstract}

\category{D.3.1}{Formal Definitions and Theory}{}
\category{F.3.2}{Semantics of Programming Languages}{}
\category{F.3.3}{Studies of Program Constructs}{Type structure}

\terms
Languages, Theory

\keywords

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

\nocite{infeffects}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Core Reversible Language: {{Pi}} }
\label{sec:pi}

There are two ways to read types and Curry-Howard: 
\begin{enumerate}
\item One first concentrates on the \emph{types}, as propositions.  We can
  come up with a ``type system'' which reflects a logic.  Then, as a second
  step, choose the proofs which inhabit these types/propositions, and call
  these proofs \emph{terms}.
\item Alternatively, one first concentrates on the \emph{terms}, focusing on
  their operational semantics as primary.  Then, as a second step, one
  designs a ``type system'' to classify the terms, so as to be able to prove
  progress and preservations theorems, to ensure that well-typed terms
  ``don't go wrong''.
\end{enumerate}

Under the first interpretation, coming up with an operational interpretation
of the terms can be challenging: witness the ongoing effort to understand
aspects of classical logic as a type system for a (programming) calculus.
Similarly, various modal logics as well as linear logic are still being given
different interpretations as type systems for calculi, to different degrees
of effectiveness.

The second interpretation is not straightforward either.  The definition of
``doing wrong'' frequently gets expanded to covering effects, permissions,
memory access, and so forth, which can seriously affect the decidability
properties of the type system (not to mention efficiency when decidability is
not an issue).

Even in the first interpretation, the understanding is that terms represent
\emph{values}, which are irreducible terms (proof terms reduce principally
via cut elimination, but there are other reduction rules for projections,
etc).

We wish to shift the focus and consider \emph{equivalences} as our types.
This will mean, under the first interpretation, that our terms will be
witnesses for these equivalences.  The underlying reason is that we wish to
examine \emph{reversible computation}.  From an information-theoretic point
of view, this means that information \emph{must be preserved}, although it
can certainly be rearranged in many different ways.

In other words, we wish to regard types with shape $T_1\equiv T_2$, where
$T_1$ and $T_2$ are usual value-types.  In fact, we want more: we wish to
regard these as \emph{oriented}: the eventual operational interpretation will
regard equivalence witnesses as \emph{inducing} a transformation from values
of types $T_1$ to values of type $T_2$.  Of course, as these are
equivalences, there will always be a term which inhabits $T_2\equiv T_1$
whenever there is one inhabiting $T_1 \equiv T_2$.  This will be clearly
visible when we give rules for terms, as these will always be symmetric (and
will provably induce invertible transformations).

So we start with a type theory $\mathcal{T}$, and move to another type theory
whose only shape is $T_1\equiv T_2$.  The constants of this new type theory
(denoted $\mathcal{T}^{\equiv}$) will be proofs of type equivalences of the
original theory.  Note that this is a very general set-up: given any
equational theory, we can ``reify'' the axioms as constants.  [Insert the
example of doing this for Monoid here].  If some of the constructors of
$\mathcal{T}$ respect equivalences, then this induces combinators on
$\mathcal{T}^{\equiv}$; the rules of equivalences (reflexivity, transitivity,
symmetry) induce more.

Note that it is very important that the theories be equational: the only
thing we can talk about are total equivalences.  One of the most natural
theory to consider, at least in the usual setting of type theories
$\mathcal{T}$ with sums and products, is that of \emph{semirings} (sometimes
jokingly called \emph{rigs}), as the rules of semirings turns out to exactly
reflect the judgments of such a type theory.

What happens in equational theories is that we declare a number of values to
be \emph{equal}.  Moving to the categorical setting, we are no longer
interested in equalities, but just equivalences.  In equational theories, the
equality witnesses are not particularly interesting, as they have no
computational content -- if $a = b$, there is no way we can distinguish $a$
and $b$, whereas if we only have $a \equiv b$, that may not be the case.
[The difference is that between syntax and semantics, where we have
equivalences on the syntax which reflect equilities in the semantics.  In
other words, plus(one,one) and two are easily distinguished, but 1+1 and 2
are not.]

Furthermore, we can use these equivalences in an effective way: we can create
a (total!) function $ev : T_1 \equiv T_2 \rightarrow T_1 \rightarrow T_2$.
In other words, given a proof of an equivalence between terms, and a value of
the first type, we can get a value of the second type.

It is important to understand that $ev$ \emph{mixes} two levels: that of
(proofs of) equivalence and that of values.  $ev$ is an \emph{induced
  action}.

One further remark: our language of equivalences is actually a \emph{free}
theory.  There are absolutely no equalities (or equivalences) given between
its terms.  Of course, some of its terms are ``observationally equivalent'',
in the sense that $ev$ will not be able to witness any difference.
Nevertheless, we will not regard such terms as equal.

What needs to be shown then is the relationship between the equational theory
of semirings and the first-order type theory with sums and products.

Then we can investigate how adding further axioms to our equational theory
(and corresponding witnesses to our type theory) changes things.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\acks We thank Michael Adams, Will Byrd, Lindsey Kuper, and Yin Wang for
helpful comments and questions. This project is supported by NSF grant
1217454 on ``Information Effects.'' Initial stages of the research were
partially funded by Indiana University's Office of the Vice President for
Research and the Office of the Vice Provost for Research through its Faculty
Research Support Program.  We also acknowledge support from Indiana
University's Institute for Advanced Study.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{small}
\bibliographystyle{abbrvnat}
\bibliography{cites}
\end{small}

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%