### 2013-02-24 frac.tex
global revision

\documentclass{llncs}
\usepackage{url}
\usepackage{proof}
\usepackage{amssymb}
\usepackage{stmaryrd}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{comment}

\newcommand{\dgm}[2][1.5]{
\begin{center}
\scalebox{#1}{
\includegraphics{diagrams/#2.pdf}
}
\end{center}
}

\newcommand{\todo}[1]{\textbf{TODO:} #1}
\newcommand{\jacques}[1]{\textsc{Jacques says:} #1}
\newcommand{\amr}[1]{\textsc{Amr says:} #1}

\newcommand{\roshan}[1]{} 
%% \newcommand{\roshan}[1]{\textsc{Roshan says:} 
%%   \textit{#1}
%% }

\hyphenation{a-reas}

%subcode-inline{bnf-inline} name Pi
%! swap+ = \mathit{swap}^+
%! swap* = \mathit{swap}^*
%! dagger =  ^{\dagger}
%! assocl+ = \mathit{assocl}^+
%! assocr+ = \mathit{assocr}^+
%! assocl* = \mathit{assocl}^*
%! assocr* = \mathit{assocr}^*
%! identr* = \mathit{uniti}
%! identl* = \mathit{unite}
%! dist = \mathit{distrib}
%! factor = \mathit{factor}
%! (o) = \fatsemi
%! (;) = \fatsemi
%! (*) = \times
%! (+) = +
%! foldB = fold_B
%! unfoldB = unfold_B
%! foldN = fold_N
%! unfoldN = unfold_N
%! trace+ = \mathit{trace}^{+}
%! trace* = \mathit{trace}^{\times}
%! :-* = \multimap
%! :-+ = \multimap^{+}
%! emptyset = \emptyset

%subcode-inline{bnf-inline} regex \{\{(((\}[^\}])|[^\}])*)\}\} name main include Pi
%! [^ = \ulcorner
%! ^] = \urcorner
%! [v = \llcorner
%! v] = \lrcorner
%! [[ = \llbracket
%! ]] = \rrbracket
%! ^^^ = ^{\dagger}
%! eta* = \eta
%! eps* = \epsilon
%! Union = \bigcup
%! in = \in
%! |-->* = \mapsto^{*}
%! |-->> = \mapsto_{\ggg}
%! |-->let = \mapsto_{let}
%! |--> = \mapsto
%! <--| = \mapsfrom
%! |- = \vdash
%! <=> = \Longleftrightarrow
%! <-> = \leftrightarrow
%! -> = \rightarrow
%! ~> = \leadsto
%! ::= = ::=
%! /= = \neq
%! vi = v_i
%! di = d_i
%! si = s_i
%! sj = s_j
%! F = \texttt{F}
%! T = \texttt{T}
%! forall = \forall
%! exists = \exists
%! empty = \emptyset
%! eta = \eta
%! where = \textbf{where}
%! epsilon = \varepsilon
%! least = \phi
%! loop+ = loop_{+}
%! loop* = loop_{\times}
%! CatC = {\mathcal C}
%! CatA = {\mathcal A}
%! gamma = \gamma
%! {[ = \{
%! ]} = \}
%! elem = \in
%! dagger = ^\dagger
%! alpha = \alpha
%! beta = \beta
%! rho = \rho
%! @@ = \mu
%! @ = \,@\,
%! Pow = \mathcal{P}
%! Pi = \Pi
%! PiT = \Pi^{o}
%! PiEE = \Pi^{\eta\epsilon}_{+}
%! PiT = \Pi^{o}
%! PiTF = \Pi^{/}
%! bullet = \bullet
%! * = \times

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\title{Fractional Types}
\author{Roshan P. James$^{1}$ \and Zachary Sparks$^{1}$ \and Jacques Carette$^{2}$ \and Amr Sabry$^{1}$}
\institute{$^{(1)}$~Indiana University \qquad $^{(2)}$~McMaster University}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  In previous work, we developed a \emph{first-order},
  information-preserving, and reversible programming language {{Pi}} founded
  on type isomorphisms. Being restricted to first-order types limits the
  expressiveness of the language: it is not possible, for example, to
  abstract common program fragments into a higher-level combinator. In this
  paper, we introduce a higher-order extension of {{Pi}} based on the novel
  concept of \emph{fractional types} {{1/b}}. Intuitively, a value of a
  fractional type {{1/v}} represents \emph{negative} information. A
  function is modeled by a pair {{(1/v1,v2)}} with {{1/v1}}
  representing the needed argument and {{v2}} representing the
  result. Fractional values are first-class: they can be freely propagated
  and transformed but must ultimately --- in a complete program --- be offset
  by the corresponding amount of positive information.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

We are witnessing a convergence of ideas from several distinct research
communities (physics, mathematics, and computer science) towards replacing
\emph{equalities} by \emph{isomorphisms}. The combined programme has sparked
a significant amount of research that unveiled new and surprising connections
between geometry, algebra, logic, and computation (see~\cite{baez2011physics}
for an overview of some of the connections).

In the physics community, Landauer~\cite{Landauer:1961,Landauer},
Feynman~\cite{springerlink:10.1007/BF02650179}, and others have interpreted
the laws of physics as fundamentally related to computation. The great
majority of these laws are formulated as equalities between different
physical observables which is unsatisfying: \emph{different} physical
observables should not be related by an \emph{equality}. It is more
appropriate to relate them by an \emph{isomorphism} that witnesses, explains,
and models the process of transforming one observable to the other.

In the mathematics and logic community, Martin-L\"of developed an extension
of the simply typed $\lambda$-calculus originally intended to provide a
rigorous framework for constructive
mathematics~\cite{citeulike:7374951}. This theory has been further extended
with \emph{identity types} representing the proposition that two terms are
``equal.'' (See~\cite{streicher,warren} for a survey.) Briefly speaking,
given two terms $a$ and $b$ of the same type $A$, one forms the type
$\texttt{Id}_A(a,b)$ representing the proposition that~$a$ and~$b$ are equal:
in other words, a term of type $\texttt{Id}_A(a,b)$ witnesses, explains, and
models the process of transforming $a$ to $b$ and vice-versa.

\roshan{The above para is too technical in a weird way. Its unclear
  what it contributes. If one wants to make a point of computational
  witnesses of isomorphisms its easier to explain and cite in terms of
  Andreas Blass and Fiore -- and more generally by Barry Mazur. I say
  this also because it seems that HOTT is more concerned with
  topological structure of equalities, even though they tend to start
  their literature with some talk of intentional specifications.}

%% AMR: I am not sure how to take this comment into account. I think we
%% should make a specific connection to Martin-Lof type theory to at least
%% justify the mention of groupoids in the conclusion. I agree though that
%% this is a very technical point and I do find it hard to express the point
%% in a couple of sentences. Suggestions welcome.

In the computer science community, the theory and practice of type
isomorphisms is well-established. Originally, such type isomorphisms were
motivated by the pragmatic concern of searching large libraries of functions
by providing one of the many possible isomorphic types for the desired
function~\cite{Rittri:1989:UTS:99370.99384}. More recently, type isomorphisms
have taken a more central role as \emph{the} fundamental computational
mechanism from which more conventional, i.e., irreversible computation, is
derived. In our own previous
work~\cite{James:2012:IE:2103656.2103667,rc2011,rc2012} we started with the
notion of type isomorphism and developed from it a family of programming
languages, {{Pi}} with various superscripts, in which computation is an
isomorphism preserving the information-theoretic entropy.
%% Moreover these languages have models in various flavors of symmetric
%% monoidal categories (which are the backbone of models for linear logic and
%% quantum computation).
%% We have also invested a significant amount of time in using {{Pi}} as
%% an actual programming language and established several idioms for
%% programming large classes of interesting (recursive) programs in
%% {{Pi}}.

A major open problem remains, however: a higher-order extension of
{{Pi}}. This extension is of fundamental importance in all the originating
research areas.
%% In physics, it allows a process or observer to be treated as
%% ``data'' that can be transformed by higher-level processes or observed by
%% meta-level observers. 
In physics, it allows for quantum states to be viewed as processes and
processes to be viewed as states, such as with the Choi-Jamiolkowski
isomorphism~\cite{choi1975completely,jamiolkowski1972linear}.  In mathematics
and logic, it allows the equivalence between different proofs of type
$\texttt{Id}_A(a,b)$ to itself be expressed as an isomorphism (of a higher
type) $\texttt{Id}_{\texttt{Id}_A(a,b)}(p,q)$. Finally, in computer science,
higher-order types allow code to abstract over other code fragments as well
as the manipulation of code as data and data as code.

\roshan{Again, HOTT seems jarring. Saying something about logic in
  this way however seems useful. Maybe one can use one of the examples
  of equivalence between proofs that Baez's Rosetta stone uses. }

Technically speaking, obtaining a higher-order extension requires the
construction of a \emph{closed category} from the underlying monoidal
category for~{{Pi}}. Although the general idea of such a construction is
well-understood, the details of adapting it to an actual programming language
are subtle.  Our main novel technical device to achieving the higher-order
extension is a \emph{fractional type} which represents \emph{negative
  information} and which is so named because of its duality with conventional
product types.  
%% By adding a formal (linear) dual to (type theoretic) products, we can
%% internalize the notion that ``taking an input'' corresponds
%% to negative information, which turns out to be the basis of first-class
%% functions.  
%% \roshan{The last sentence is terrible.}
The remainder of the paper reviews {{Pi}} and then introduces the syntax and
semantics of the extension with fractional types. We then study the
properties of the extended language and establish its expressiveness via
several constructions and examples that exploit its ability to model
higher-order computations.

%% AMR:
%% \roshan{It is unclear what the Id types refer to. We should provide
%%   some more context here}
%% I think the previous paragraph in the intro that talks about Martin-Lof type 
%% theory provides the context. Anything else??

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Background: {{Pi}} }
\label{sec:pi}

We review our language {{Pi}} providing the necessary background and
context for our higher-order extension.\footnote{The presentation in this
  section focuses on the simplest version of {{Pi}}. Other versions
  include the empty type, recursive types, and trace operators but these
  extensions are orthogonal to the higher-order extension emphasized in this
  paper.} The terms of {{Pi}} are not classical values and functions;
rather, the terms are isomorphism witnesses.  In other words, the terms of
{{Pi}} are proofs that certain ``shapes of values'' are isomorphic.
And, in classical Curry-Howard fashion, our operational semantics shows how
these proofs can be directly interpreted as actions on ordinary values which
effect this shape transformation. Of course, ``shapes of values'' are very
familiar already: they are usually called \emph{types}.  But frequently one
designs a type system as a method of classifying terms, with the eventual
purpose to show that certain properties of well-typed terms hold, such as
safety.  Our approach is different: we start from a type system, and then
present a term language which naturally inhabits these types, along with an
appropriate operational semantics.

\paragraph*{Data.}
We view {{Pi}} as having two levels:  it has traditional values, given by:
%subcode{bnf} include main
% values, v ::= () | left v | right v | (v, v)

\noindent and these are classified by ordinary types:
%subcode{bnf} include main
% value types, b ::= 1 | b + b | b * b

\noindent 
Types include the unit type {{1}}, sum types {{b1+b2}}, and
products types {{b1*b2}}.  Values includes {{()}} which is the only value of
type {{1}}, {{left v}} and {{right v}} which inject~{{v}} into a sum type,
and {{(v1,v2)}} which builds a value of product type. But these values should
be regarded as largely ancillary: we do not treat them as first-class citizens,
and they only occur when observing the effect of an isomorphism.

\paragraph*{Isomorphisms.} The important terms of {{Pi}} are witnesses to
(value) type isomorphisms.  They are also typed, by the shape of the (value)
type isomorphism they witness {{b <-> b}}.  Specifically, they are witnesses to
the following isomorphisms:
%subcode{bnf} include main
%! columnStyle = r@{\hspace{-0.5pt}}r c l@{\hspace{-0.5pt}}l
%swap+ :&  b1 + b2 & <-> & b2 + b1 &: swap+
%assocl+ :&  b1 + (b2 + b3) & <-> & (b1 + b2) + b3 &: assocr+
%identl* :&  1 * b & <-> & b &: identr*
%swap* :&  b1 * b2 & <-> & b2 * b1 &: swap*
%assocl* :&  b1 * (b2 * b3) & <-> & (b1 * b2) * b3 &: assocr*
%dist :&~ (b1 + b2) * b3 & <-> & (b1 * b3) + (b2 * b3)~ &: factor

\noindent Each line of the above table introduces a pair of dual
constants\footnote{where {{swap*}} and {{swap+}} are self-dual.} that witness
the type isomorphism in the middle.  These are the base (non-reducible) terms
of the second, principal level of {{Pi}}. Note how the above has two
readings: first as a set of typing relations for a set of constants. Second,
if these axioms are seen as universally quantified, orientable statements,
they also induce transformations of the (traditional) values. The
(categorical) intuition here is that these axioms have computational content
because they witness isomorphisms rather than merely stating an extensional
equality.

The isomorphisms are extended to form a congruence relation by adding the
following constructors that witness equivalence and compatible closure:

%subcode{proof} include main
%@  ~
%@@ id : b <-> b 
%
%@ c : b1 <-> b2
%@@ sym c : b2 <-> b1
%
%@ c1 : b1 <-> b2
%@ c2 : b2 <-> b3
%@@ c1(;)c2 : b1 <-> b3
%---
%@ c1 : b1 <-> b3
%@ c2 : b2 <-> b4
%@@ c1 (+) c2 : b1 + b2 <-> b3 + b4
%
%@ c1 : b1 <-> b3
%@ c2 : b2 <-> b4
%@@ c1 (*) c2 : b1 * b2 <-> b3 * b4
\noindent The syntax is overloaded: we use the same symbol at the value-type level
and at the isomorphism-type level for denoting sums and products.  Hopefully
this will not cause undue confusion.

It is important to note that ``values'' and ``isomorphisms'' are completely
separate syntactic categories which do not intermix. The semantics of the
language come when these are made to interact at the ``top level'' via
\emph{application}:
%subcode{bnf} include main
% top level term, l ::= c v

% \noindent
% To summarize, the syntax of {{Pi}} is given as follows.

% \begin{definition}{(Syntax of {{Pi}})}
% \label{def:Pi}
% %subcode{bnf} include main
% % value types, b ::= 1 | b+b | b*b 
% % values, v ::= () | left v | right v | (v,v) 
% %
% % iso.~types, t ::= b <-> b
% % base iso ::= swap+ | assocl+ | assocr+ 
% %     &|& unite | uniti | swap* | assocl* | assocr* 
% %     &|& dist | factor 
% % iso comb., c ::= iso | id | sym c | c (;) c | c (+) c | c (*) c 
% % top level term, l ::= c v
% \end{definition}

The language presented above, at the type level, models a commutative ringoid
where the multiplicative structure forms a commutative monoid, but the
additive structure is just a commutative semigroup.  Note that the version of
{{Pi}} that include the empty type with its usual laws exactly captures, at
the type level, the notion of a \emph{semiring} (occasionally called a
\emph{rig}) where we replace equality by isomorphism.  Semantically, {{Pi}}
models a \emph{bimonoidal category} whose simplest example is the category of
finite sets and bijections. In that interpretation, each value type denotes a
finite set of a size calculated by viewing the types as natural numbers and
each combinator {{c : b1 <-> b2}} denotes a bijection between the sets
denoted by~{{b1}} and~{{b2}}. (We discuss the operational semantics in
conjuction with our extension in the next section.)

% Operationally, the
% semantics of {{Pi}} is given using two mutually recursive interpreters: one
% going forward and one going backwards. The use of {{sym}} switches control
% from one evaluator to the other. We will present the operational semantics in
% the next section along with the extension with fractional types and
% values. For now, we state without proof that the evaluation of well-typed
% combinators always terminates and that {{Pi}} is logically reversible, i.e.,
% that for all combinators {{c : b1 <-> b2}} and values {{v1 : b1}} and {{v2 :
%     b2}} we have the forward evaluation of {{c v1}} produces {{v2}} iff the
% backwards evaluation of {{c v2}} produces {{v1}}.

%% AMR:
%% \jacques{Since we have Agda code for the above, should we refer to it
%% here, and make it available?}
%% That is a great idea but it would take a significant amount of time to
%% clean up the code and make it presentable. If someone is willing to do
%% that, then we should definitely add a reference to the code and make it
%% available.
%% Jacques:  Right.  I'll definitely work on cleaning up the paper first,
%% then see if there is time left for the code.

%% As is usual with type theories,
%% the coherence conditions (pentagonal and triangle identity) are not explicit,
%% but would correspond to certain identity types being trivial.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Language: {{PiTF}} }

The language {{Pi}} models isomorphisms of values rather well, and as we
established before~\cite{James:2012:IE:2103656.2103667,rc2011,rc2012}, it is
logically reversible and each computation it expresses preserves the
information-theoretic entropy. However, purposefully, it has no distinguished
notion of \emph{input} or \emph{output}; more precisely, because of strict
preservation of information, these two concepts coincide.  This is the root
of reversibility. But if we want to model functions of any flavor as values,
we need to differentiate between these notions. Our idea, inspired by
computational dualities~\cite{Filinski:1989:DCI:648332.755574,Curien:2000},
polarity~\cite{Girard87tcs,10.1109/LICS.2010.23}, and the categorical notion
of \emph{compact
  closure}~\cite{Selinger:2007:DCC:1229185.1229207,Abramsky:2004:CSQ:1018438.1021878},
is to introduce a \emph{formal dual} for our types.  We thus consider a type
in a negative position to possess negative information, as it is really a
request for information.  Since information is logarithmic, this means that a
request for information should behave like a \emph{fractional type}.

% \roshan{It is unclear how fractions help differentiate between inputs
%   and outputs. On first reading this seemed promising, but nothing
%   in the rest of the paper substantiates this point of view.}
% AMR: a function a -o b is modeled by (1/a,b) not (a,b) 
% Jacques: Exactly!  1/a is negative information is input, b is positive is
%    output.
%

% \roshan{I think we should arrive at relational programming as a means
%   of giving semantics to fractionals, rather than the other way
%   round. To show that relational programming is reversible, one does
%   not need to resort to fractionals. Everyone understand that
%   relations can be reversed and none of this setup with {{Pi}} is
%   required for that. Our key point should be to communicate that
%   fractionals have a clean semantics in the setting of relational
%   programming rather than the other way round.}
% AMR: relations are indeed only presented as a way to explain the semantics

The extension of the sets of types and values from {{Pi}} to {{PiTF}} is
simple:
%subcode{bnf} include main
% value types, b ::= ... | 1/b
% values, v ::= ... | 1/v

\noindent For a given type {{b}}, the values of type {{1/b}} are of the form
{{1/v}} where {{v : b}}. Note that {{1/v}} is purely formal, as is
{{1/b}}.

Semantically, we are extending the symmetric monoidal category modeling
{{Pi}} to a \emph{compact closed} one, i.e., a category in which morphisms
are representable as objects. In such a setting, the new dual objects (i.e.,
the fractionals) must satisfy the following isomorphism witnessed by two new
type indexed combinators {{eta*_b}} and {{eps*_b}}:
\begin{multicols}{3}
  \dgm{eta_times2}

\columnbreak

%subcode{bnf} include main
%! columnStyle = r r c l l
%eta*_b :&  1 & <-> & 1/b * b &: eps*_b
%% XXX: diagrams need unit wires(?)
%% Roshan: We should just be clear that unit wires are dropped. Dgms
%% including them dont communicate any additional information and
%% appear crowded.

\columnbreak

  \dgm{eps_times2}  
\end{multicols}

\vspace{-10pt}
\noindent From a programming perspective, we think of a value {{1/v}} as a
\emph{first-class constraint} that can only be satisfied if it is matched
with an actual value {{v}} or as a \emph{pattern} representing the absence of
some information of a specific shape (i.e., \emph{negative information}),
that can only be reconciliated by an actual value {{v}}. The combinator
{{eta*_b : 1 <-> 1/b * b}} thus represents a fission point which creates ---
out of no information --- an equal amount of negative and positive
information. Symmetrically, the combinator {{eps*_b : 1/b * b <-> 1}} matches
up an equal amount of negative and positive information producing no residual
information.

Historically, versions of compact closed categories were introduced by
Abramsky and Coecke~\cite{Abramsky:2004:CSQ:1018438.1021878} and by
Selinger~\cite{Selinger:2007:DCC:1229185.1229207} as the generalization of
monoidal categories to model quantum computation.  The first and most
intuitive example of such categories is the category of finite sets and
relations. In that case, the dual operation (i.e., the fractional type)
collapses by defining {{1/b}} to be {{b}}. In spite of this degenerate
interpretation of fractionals, this is still an interesting category as it
provides a model for a reversible higher-order language. Indeed, any relation
is reversible via the standard \emph{converse} operation on a relation.
Furthermore, as any relation between {{b1}} and {{b2}} can be represented as
a subset of {{b1 * b2}}, i.e., as a value.

In this section, we provide a simple semantics of {{PiTF}} in the category of
finite sets and relations. We interpret each combinator {{c : b1 <-> b2}} as
a \emph{relation} between the sets denoted by {{b1}} and {{b2}}
respectively. As we illustrate, this semantics allows \emph{arbitrary}
relations (including the empty relation and relations that are not
isomorphisms) to be represented as values in {{PiTF}}.  In Sec.~\ref{sec:cat}
we discuss more refined semantics that are more suitable for richer
categories.

\begin{definition}[Denotation of Value Types]
\label{chx:def:denot}
Each type denotes a finite set of values as follows:
%subcode{opsem} include main
%! <- = \leftarrow
%! union = \cup
% [[1]] '= {[ () ]}
% [[b1 + b2]] '= {[ left v ~|~ v <- [[b1]] ]} union {[ right v ~|~ v <- [[b2]] ]}
% [[b1 * b2]] '= {[ (v1, v2) ~|~ v1 <- [[b1]], v2 <- [[b2]] ]}
% [[1/b]] '= {[ 1/v ~|~ v <- [[b]] ]}
\end{definition}

We specify
relations using a deductive system whose judgments are of the form 
{{ v1 ~~c~~ v2 }} indicating that the pair {{(v1,v2)}} is in the relation 
denoted by {{c}}.

\begin{definition}[Relational semantics]
\label{def:relational-PiTF}
Each combinator {{c : b1 <-> b2}} in {{PiTF}} denotes a relation
as specified below.~\footnote{
In the interest of brevity, we have treated the type annotations in the base
isomorphisms as implicit; for example, the rule for {{assocr*}} should really
read:

%subcode{proof} include main
%@ v1 in b1
%@ v2 in b2
%@ v3 in b3
%@@ ((v1, v2), v3) ~~assocr*_{b1, b2, b3}~~ (v1, (v2, v3))
}
%subcode{proof} include main
%@ ~
%@@ (left v) ~~swap+~~ (right v)
%
%@ ~
%@@ (right v) ~~swap+~~ (left v)
%----
%@ ~
%@@ (left v) ~~assocl+~~ (left (left v))
%
%@ ~
%@@ (right (left v)) ~~assocl+~~ (left (right v))
%----
%@ ~
%@@ (right (right v)) ~~assocl+~~ (right v)
%
%@ ~
%@@ (left (left v)) ~~assocr+~~ (left v)
%----
%@ ~
%@@ (left (right v)) ~~assocr+~~ (right (left v))
%
%@ ~
%@@ (right v) ~~assocr+~~ (right (right v))
%----
%@ ~
%@@ ((), v) ~~unite~~ v
%
%@ ~
%@@ v ~~uniti~~ ((), v)
%----
%@ ~
%@@ (v1, v2) ~~swap*~~ (v2, v1)
%----
%@ ~
%@@ (v1, (v2, v3)) ~~assocl*~~ ((v1, v2), v3)
%
%@ ~
%@@ ((v1, v2), v3) ~~assocr*~~ (v1, (v2, v3))
%----
%@ ~
%@@ (left v1, v3) ~~dist~~ (left (v1, v3))
%
%@ ~
%@@ (right v2, v3) ~~dist~~ (right (v2, v3))
%----
%@ ~
%@@ (left (v1, v3)) ~~factor~~ (left v1, v3)
%
%@ ~
%@@ (right (v2, v3)) ~~factor~~ (right v2, v3)
%----
%@ ~
%@@ v ~~id~~ v
%
%@ v' ~~c~~ v
%@@ v ~~(sym c)~~ v'
%
%@ v1 ~~c1~~ v2
%@ v2 ~~c2~~ v3
%@@ v1 ~~(c1(;)c2)~~ v3
%----
%@ v ~~c1~~ v'
%@@ (left v) ~~(c1 (+) c2)~~ (left v')
%
%@ v ~~c2~~ v'
%@@ (right v) ~~(c1 (+) c2)~~ (right v')
%----
%@ v1 ~~c1~~ v1'
%@ v2 ~~c2~~ v2'
%@@ (v1, v2) ~~(c1 (*) c2)~~ (v1', v2')
%----
%@ ~
%@@ () ~~eta*_b~~ (1/v,v)
%
%@ ~
%@@ (1/v,v) ~~eps*_b~~ ()
\end{definition}

The semantics above defines a relation rather than a function. It is still
reversible in the sense that it ``mode-checks''. In other words, for any
combinator {{c}} we can treat either its left or its right side as an input;
this allows us to define {{sym}} as just relational converse, which amounts
to an argument swap. We can then give an operational interpretation to the
semantic relation by defining two functions: a forward evaluator and a
backwards evaluator that exercise the relation in opposite directions, but
which now can return a \emph{set} of results. This operational view exploits
the usual conversion of a relation from a subset of {{A * B}} to a function
from {{A}} to the powerset of {{B}} whose composition is given by the Kleisli
composition in the powerset monad. It is straightforward to check that if the
forward evaluation of {{c v1}} produces {{v2}} as a possible answer then the
backwards evaluation of {{c v2}} produces {{v1}} as a possible answer.  One
can check that all relations for {{Pi}} are total functional relations whose
converse are also total functional relations -- aka isomorphisms.  However
{{eta*_b}} relates {{()}} to multiple values (one for each value of {{b}}),
and is thus not functional; {{eps*_b}} is exactly its relational converse, as
expected.

% \roshan{ In the rules above it is unclear as to which v is chosen in
%   the rule for eta. It is unclear that these rules are logically
%   reversible.  }

% \begin{proposition}[Logical Reversibility]
% \label{chx:prop:logrev-tracep}
% For all combinators {{c : b1 <-> b2}} and values {{v1 : b1}} and 
% {{v2 : b2}} we have {{v1 ~~c~~ v2}} iff {{v2 ~~(sym c)~~ v1}}.
% \end{proposition}

% Some thoughts on a relational semantics. We define {{ a <-> b }}
% $\equiv$ {{a R b}}

% \noindent
% We now have primitives:
% %subcode{opsem} include main
% % swap*:b1*b2<->b2*b1 '= {[ ((v1, v2), (v2, v1)) ~|~ v1 in b1, v2 in b2 ]}
% % swap+:b1+b2<->b2+b1 '= {[ ((left v1), (right v1)) and ((right v2), (left v2)) 
% %                     && ~~~| v1 in b1, v2 in b2]}
% % assocl*:b1*(b2*b3)<->(b1*b2)*b3 '= {[ ((v1, (v2, v3)), ((v1, v2), v3)) 
% %                   & & ~~~|~ v1 in b1, v2 in b2, v3 in b3]}
% etc..

% \noindent
% closure combinators:
% %subcode{opsem} include main
% % sym c '= {[(v2, v1) ~|~ (v1, v2) in c ]}
% % c1 (;) c2 '= {[(v1, v3) ~|~ v2 in b2, (v1, v2) in c1, (v2, v3) in c2 ]}
% % c1 (*) c2 '= {[((v1, v2), (v3, v4)) ~|~ (v1, v3) in c1, (v2, v4) in c2 ]}
% etc..

% \noindent
% Now we have 
% %subcode{opsem} include main
% % eta* '= {[ ((), (v, 1/v)) ~|~ v in b ]}
% % eps* '= {[ ((v, 1/v), ()) ~|~ v in b ]}

% Now {{PiTF}} is logically reversible in that for all {{c:b1<->b2}} we
% have {{c^{dagger}: b2 <-> b1}} such that 
% %
% {{c^{dagger} = {[(v2, v1)~|~(v1,v2) in c ]} }}. This is
% straightforward from the definitions above.

% I think defining eval is an open problem.  Consider a few relations:

% \begin{enumerate}
% \item 
% {{empty = {[ ]} }}

% \item 
% {{full =  {[ R, G, B ]} * {[ R, G, B ]} }}

% \item 
% {{id  =  {[ (R, R), (G, G), (B, B) ]} }}

% \item 
% {{cycle = {[ (R, R), (R, G), (G, G), (G, B), (B, B), (B, R) ]} }}
% \end{enumerate}

% Can we define an {{eval}} that meaningfully distinguishes these? What
% can we say about {{(c (;) c^^^) v}} based on {{eval}}?  Here are
% several possible {{eval}} definitions:

% \begin{enumerate}
% \item {{ c v = v'}} where {{ v' in {[ v' ~|~ (v, v') in c ]} }}. The
%   eval maps {{b1 -> b2}}.

% \item {{ c v = {[ v' ~|~ (v, v') in c ]} }}. The eval maps {{b1 -> Pow(b2)}}.

% \item {{ c vs = {[ v' ~|~ (v, v') in c, v in vs ]} }}. where
%   {{vs:Pow(b1)}}. The eval maps {{Pow(b1) -> Pow(b2)}}.

% \end{enumerate}

% For the relations above, none of the suggested eval definitions give a
% logically reversible semantics.

% \roshan{If {{|b1|=|b2|}}, can we define all relations between {{b1}} and
% {{b2}}? If yes, then we have a model for {{PiTF}} where the objects
% are sets {{b}} and the morphisms are the relations between objects of
% equal size.}

% \roshan{ I suspect that all relations can be constructed, but we have
%   to show this. This is not obvious.}

% \roshan{In the judgments below it is misleading that only {{eta*}}
%   appears indexed by a type. To construct each family of relations,
%   the corresponding definition must be indexed by their types. In
%   other words, there is nothing special about {{eta*}}. Further, I am
%   unsure of how to format this so that the combinators are correctly
%   indexed by their types for both primitives and for composition
%   combinators. We could do the following, but it seems to be a
%   overkill.}

% %subcode{proof} include main
% %@ v1 in [[b1]]
% %@ v2 in [[b2]]
% %@ v3 in [[b3]]
% %@@ (v1, (v2, v3)) ~assocl*~ ((v1, v2), v3)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Expressiveness and Examples}

Having introduced the syntax of our extended language, its semantics in the
category of sets and relations, and its operational semantics using forward
and backwards interpreters, we now illustrate its expressiveness as a
higher-order programming language. In the following presentation consisting
of numerous programming examples, we use {{bool}} as an abbreviation of
{{1+1}} with {{true}} as an abbreviation for {{left ()}} and {{false}} as an
abbreviation for {{right ()}}. In addition, instead of presenting the code
using the syntax of {{Pi}}, we use circuit diagrams that are hopefully more
intuitive. Each diagrams represents a combinator whose evaluation consists of
propagating values along the wires.  For the sake of readability, we omit
obvious re-shuffling circuitry and elide trivial unit wires which carry no
information.\footnote{The full code is available upon request. We plan on
  releasing it as soon as possible.}

%%%%%%%%%%%%
\subsection{First-Class Relations}

The most basic additional expressiveness of {{PiTF}} over {{Pi}} is the
ability to express relations as values. Indeed a value of type {{1/b1 * b2}}
is a pair of a \emph{constraint} that can only be satisfied by some 
{{v1 : b1}} and a value {{v2 : b2}}. In other words, it corresponds to a
function or relation which when ``given'' a value {{v1 : b1}} ``releases''
the value {{v2 : b2}}. To emphasize this view, we introduce the abbreviation:

%subcode{bnf} include main
%  b1 :-* b2 &::=& 1/b1 * b2

\noindent which suggests a function-like behavior for the pair of a fractional value
and a regular value. What is remarkable is that we can almost trivially turn
any combinator {{c : b1 <-> b2}} into a (constant) value of 
type {{b1 :-* b2}} as shown below on the left:

%subcode{opsem} include main
%! columnStyle = rcl
% name &:& (b1 <-> b2) -> (1 <-> (b1 :-* b2))
% name c &=& eta*_{b1} (;) (id (*) c)

\begin{multicols}{2}
  \dgm{function-times}

  \dgm{delimc-times}
\end{multicols}
\noindent Dually, as illustrated above on the right, we also have a
{{coname}} combinator which when given an answer {{v2 : b2}}, and a request
for the corresponding input {{1/b1}} for {{c : b1 <-> b2}}, eliminates this
as a un-needed computation.

More generally any combinator manipulating values of type {{b}} can be turned
into a combinator manipulating values of type {{1/b}} and vice-versa. This is
due to the fact that fractionals types satisfy a self-dual involution
relating {{b}} and {{1/(1/b)}}:

%subcode{opsem} include main
%! columnStyle = rcl
% doubleDiv &:& b <-> 1/(1/b)
% doubleDiv &=& uniti (;) (eta*_{1/b} (*) id) (;) assocr* (;) (id (*) eps*_b) (;) swap* (;) unite

% jacques: elide for now, to save space, as this diagram is backwards.
% \dgm{involution-times}

% As a simple example, consider {{swap+}} denoting boolean
% negation.  It is straightforward to check that the combinator 
% {{name~swap+ : 1 <-> (bool :-* bool)}} denotes the following relation:

% %subcode{proof} include main
% %@ ~
% %@@ () ~~(name~swap+)~~ (1/false,true)
% %
% %@ ~
% %@@ () ~~(name~swap+)~~ (1/true,false)

% \noindent The relation {{name~swap+}} maps {{()}} to a value which can be used
% to convert any {{false}} to {{true}} and vice-versa. We formalize this
% ``apply'' combinator below.

%%%%%%%%%%%%
\subsection{Higher-Order Relations}

We are now a small step from implementing various higher-order combinators
that manipulate functions or relations. In particular, we can \emph{apply},
\emph{compose}, \emph{curry}, and \emph{uncurry} values representing
relations.  We first show the realization of \emph{apply}:

%subcode{opsem} include main
%! columnStyle = rcl
% apply &:& (b1 :-* b2) * b1 <-> b2
% apply &=& swap* (;) assocl* (;) (swap* (*) id) (;) (eps*_{b1} (*) id) (;) unite

\begin{multicols}{2}
  \dgm[1.2]{apply-times1}

  \dgm{apply-times2}
\end{multicols}

\noindent Intuitively, we simply match the incoming argument of type {{b1}} with
the constraint encoded by the function. If they match, they cancel
each other and the value of type {{b2}} is exposed with no
constraints. Otherwise, the result is undefined.

\noindent A flipped variant is also useful:
%subcode{opsem} include main
%! columnStyle = rcl
% apply' &:& b1 * (b1 :-* b2) <-> b2
% apply' &=& assocl* (;) (swap* (*) id) (;) (eps*_{b1} (*) id) (;) unite

\noindent Function or relation composition is now straightforward:

%subcode{opsem} include main
%! columnStyle = rcl
% compose &:& (b1 :-* b2) * (b2 :-* b3) -> (b1 :-* b3)
% compose &=& assocr* (;) (id (*) apply')

\begin{multicols}{2}
  \dgm[1.0]{compose-times1}

  \dgm{compose-times2}
\end{multicols}

% \dgm{compose-times3}

\noindent
We can also derive currying (and dually, uncurrying) combinators. Observe that
the type of {{curry}} needs to be {{(b1 * b2 :-* b3) <-> (b1 :-* (b2 :-* b3))}},
which can be written in mathematical notation as 
{{1/(b1 * b2) * b3 = (1/b1) * ((1/b1) * b3)}}. This means {{curry}} can be
written using {{recip}}, the implementation of the mathematical identity
{{1/(b1 * b2) = 1/b1 * 1/b2}}:

%subcode{opsem} include main
%! columnStyle = rcl
% recip &:& 1 / (b1 * b2) <-> 1 / b1 * 1 / b2
% recip &=& (uniti) (;) (uniti) (;) (assocl*) (;) ((eta*_{b1} (*) eta*_{b2}) (*) id) (;)
%     && (reorder (*) id) (;) (assocr*) (;) (id (*) swap*) (id (*) eps*) (;) swap* (;) (unite)

\noindent
where {{reorder}} is the obvious combinator of type {{b1 * (b2 * b3) * b4 <-> b1 * (b3 * b2) * b4}}.
With {{recip}} out of the way, we can easily write {{curry}}:

%subcode{opsem} include main
%! columnStyle = rcl
% curry &:& b1 * b2 :-* b3 <-> b1 :-* (b2 :-* b3)
% curry &=& (id * recip) (;) (id * swap*) (;) assocl*

\noindent That {{recip}} is the heart of currying seems quite remarkable.

%%%%%%%%%%%%
\subsection{Feedback, Iteration, and Trace Operators}

Mathematically speaking, recursion and iteration can be expressed using
categorical trace
operators~\cite{joyal1996traced,Hasegawa:1997:RCS:645893.671607}.  In a
language like {{Pi}} there are two natural families of trace operators that
can be defined, an additive family (explored in detail in our previous
work~\cite{rc2011}) and a multiplicative family, which is expressible using
fractionals.

The idea of the multiplicative trace operator is as follows. We are given a
computation {{c : b1 * b2 <-> b1 * b3}}, and we build a ``looping'' version
which feeds the output value of type {{b1}} back as an input. Effectively,
this construction cancels the common type {{b1}}, to produce a new combinator
{{trace*~c:b2<->b3}}.

With fractionals, {{trace*}} becomes directly expressible:

%subcode{opsem} include main
%! columnStyle = rcl
% trace*_b &:& ((b * b1) <-> (b * b2)) -> (b1 <-> b2)
% trace*_b c &=& uniti (;) (eta*_b (*) id) (;) assocr* (;) 
%            && (id (*) c) (;) assocl* (;) (eps*_b (*) id) (;) unite

As an example, we can use the operational semantics (outlined in the previous
section) to calculate the result of applying {{trace*_{bool} (swap+ (*) id)}}
to {{false}}.
%subcode{opsem} include main
%! columnStyle = ll
% {[ ((),false) ]}  & (uniti)
% {[ ((1/false,false),false), ((1/true,true),false) ]} & (eta*_{bool} * id)
% {[ (1/false, (false,false)), (1/true,(true,false)) ]} & (assocr*)
% {[ (1/false, (true,false)), (1/true,(false,false)) ]} & (id (*) not )
% {[ ((1/false,true),false), ((1/true,false),false) ]} & (assocl*)
% emptyset & (eps*_{bool} * id)

\noindent The computation with {{true}} gives the same result. This confirms
that although our semantics is formally reversible, it does not result in
isomorphisms.

More abstractly, the evaluation of {{trace*~c : b2 <-> b3}} must ``guess'' a
value of type {{b1}} to be provided to to the inner combinator 
{{c : b1 * b2 <-> b1 * b3}}. 
This value {{v1 : b1}} cannot be arbitrary: it must be such that it is the value
produced as the first component of the result.  In general, there may be
several such fixed-point values, or none. Indeed with a little ingenuity 
(see the next section), one can 
express any desired relation by devising a circuit which keeps the desired pairs as 
valid fixed-points.

% \begin{example}
% Consider the combinator {{c : bool * 1 <-> bool * 1}}, with {{c = id}}.

% \noindent Using {{trace*}}, we construct the
% combinator {{trace* c : 1 <-> 1}}. Applying this to 
% {{() : 1}} requires us to find a value {{b : bool}} such that 
% {{(b,()) ~~d~~ (b,())}}. Given that the type {{bool}} 
% has two values, there are two values {{false}} and {{true}} that satisfy the 
% constraint. In other words, {{() ~~(trace* c)~~() }}.
% \end{example}

% \begin{example}
% \label{ch3:ex:annihilate}
% Consider a small variation on the combinator {{c}} above:

% {{c : bool * 1 <-> bool * 1}}

% {{c = swap+ (*) id }}

% \noindent which negates the boolean component of its incoming
% pair. Using the multiplicative trace operator, we can construct the
% combinator {{trace* c : 1 <-> 1}} as before. But now, applying this
% combinator to {{() : 1}} requires us to find a value {{b : bool}} such
% that {{(b,()) ~~swap+~~ (b,())}} which is impossible since {{swap+}} 
% has no fixed points.  Operationally, the evaluation of
% such a combinator should produce no value (were it to even terminate).
% \end{example}

%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{(Finite) Relational Programming}
\label{ch3:sec:lp}

Relational programming leverages set-theoretic relations and their
composition to express computational tasks in a declarative way. 
% For instance, consider this example:
% {{parent = {[ (A,B), (B,C), (A,D) ]} }}
% {{grandparent = parent (o) parent}}
% \noindent The example defines a relation {{parent}} specified using a
% set of tuples and another relation {{grandparent}} specified using the
% composition of two parent relations. If we wrote this example in a
% relational language (e.g., Prolog) and we executed the query
% {{grandparent(A)}}, we would get the answer {{ {[C]} }}.
It turns out that with the addition of the multiplicative trace, and
the move to relations motivated in the previous section, we can
express relational programming.

Consider the relation {{R}} on booleans given by:

{{ {[(false,false), (false,true), (true,false) ]}. }}

\noindent We can define a combinator {{c_R}} whose denotation is {{R}}, by
defining a combinator {{cInner : (a * bool) <-> (a * bool)}} for some type
{{a}} such that {{c_R = trace* cInner}}. The basic requirement of {{cInner}}
is that for each desired pair {{(v1,v2)}} in {{R}}, it maps {{(a0,v1)}} to
{{(a0,v2)}} for some value {{a0}} and that for each pair {{(v1,v2)}} that is
\emph{not} in {{R}}, it maps {{(a1,v1)}} to {{(a2,v2)}} for \emph{different}
values {{a1}} and {{a2}}.

% %subcode{proof} include main
% %@ ~
% %@@ ((false,false),false) ~~cInner~~ ((false,false),false) 
% %----
% %@ ~
% %@@ ((false,true),false) ~~cInner~~ ((false,true),true) 
% %----
% %@ ~
% %@@  ((false,true),true) ~~cInner~~ ((false,true),false) 
% %----
% %@ ~
% %@@ ((false,false),true) ~~cInner~~ ((true,false),true) 
% %----
% %@ ~
% %@@ ((true,false),false) ~~cInner~~ ((true,true),false)  
% %----
% %@ ~
% %@@ ((true,false),true) ~~cInner~~ ((true,true),true)  
% %----
% %@ ~
% %@@ ((true,true),false) ~~cInner~~ ((true,false),false) 
% %----
% %@ ~
% %@@ ((true,true),true) ~~cInner~~ ((false,false),true) 
% In the first three lines, the first argument is a fixed point and hence the
% multiplicative trace would map the second input to the second output
% producing the desired relation. In the remaining five cases, the first
% argument is not a fixed point and hence all these cases would be rejected as
% solutions to the constraint imposed by the multiplicative trace. It simply
% remains to find the actual realization of {{cInner}} that would produce the
% behavior above.

% %subcode{opsem} include main
% %! columnStyle = rcl
% % cInner &:& (bool * bool) * bool <-> (bool * bool) * bool
% % cInner &=& sym (swap* (;) assocl* (;) (cnot (*) id) (;) assocr* (;) swap* (;)  toffoli (;)
% %        &&  ((swap* (;) cnot (;) swap*) (*) id) (;) toffoli (;)
% %        &&  (assocr* (;) swap* (;) toffoli (;) swap* (;) assocl*) (;) toffoli (;) (cnot (*) id))

% \noindent where {{toffoli}} and {{cnot}} are the conventional reversible gates (their definition in {{Pi}} is in~\cite{rc2011}).
% %     
% % c_R &:& bool <-> bool
% % c_R &=& trace* cInner

% The above example should convince the reader that the language {{Pi}}
% with multiplicative trace is expressive enough to model finite relational
% programming. 
% %%%%%%%%%%%%%%%
% \subsection{Solving Constraints}
% \label{ch3:sec:constraints}

% A large class of constraint satisfaction problems can be expressed
% using multiplicative traces. We illustrate the main ideas with the
% implementation of a SAT solver.

For small examples, finding such combinators by trial and error is a
relatively straightforward but tedious task. It is, however, possible to
automate this task by expressing what is essentially a reversible SAT solver
as follows. In the usual setting, an instance of SAT is a function~{{f}}
which, when given some boolean inputs, returns {{true}} or {{false}}. The
function returns {{true}} when the inputs satisfy the constraints imposed by
the structure of {{f}} and a solution to the SAT problem is the set of all
inputs on which {{f}} produces {{true}}. The basic idea of our construction
is to use {{trace*}} to annihilate values that fail to satisfy the
constraints represented by the SAT instance {{f}}.

% In the usual setting, an instance of SAT is a function~{{f}} which,
% when given some boolean inputs, returns {{true}} or {{false}}. The
% function returns {{true}} when the inputs satisfy the constraints
% imposed by the structure of {{f}} and a solution to the SAT problem is
% the set of all inputs on which {{f}} produces {{true}}. The basic idea
% of our construction is to generalize the annihilation circuit from
% Ex.~\ref{ch3:ex;annihilation} to only annihilate values that fail to
% satisfy the constraints represented by the SAT instance {{f}}. To
% achieve this goal, we must however deal with several important
% details.

% First, because we are in a reversible world, our instance of SAT must be
% expressed as an isomorphism: this is easily achieved by the construction
% which embeds any boolean function {{f}} into a reversible one {{iso_f}} with
% a larger domain and range.  Given such a reversible function {{iso_f}} which
% represents a SAT instance, we first construct the circuit below:

% \begin{center}
% \scalebox{1.2}{
%   \includegraphics{diagrams/sat2.pdf}
% }
% \end{center}

% As shown in the circuit, the reversible SAT instance {{iso_f}} takes
% two sets of values and produces two outputs. The incoming values
% labeled \textsf{inputs} are the inputs we need to test for
% satisfiability. The other incoming values labeled \textsf{heap} are
% the additional inputs needed to embed the original SAT instance {{f}}
% into a reversible function. If these \textsf{heap} values are all
% initialized to {{false}}, the output wire \textsf{satisfied?}
% corresponds to the output that {{f}} would have produced on
% \textsf{inputs}. The other outputs labeled \textsf{garbage} are not
% needed for their own sake but they are important because they are used
% as inputs to the adjoint of {{iso_f}} to reproduce the inputs exactly,
% in anticipation of closing the loop with {{trace*}}.

% To summarize, the top half of the circuit is the identity function
% except that we have also managed to produce a boolean wire labeled
% \textsf{satisfied?} that tells us if the inputs satisfy the desired
% constraints. We can take this boolean value and use it to decide
% whether to negate the bottom wire (labeled \textsf{control
%   wire}). Specifically, if the inputs do \emph{not} satisfy {{f}}, the
% control wire is negated. The last wire labeled \textsf{heap control
%   wire} is negated if the heap values do not have the right initial
% values, i.e., are not all {{false}}.

% Let us call the above construction {{sat_f}}. If we now close the loop
% using {{trace*}}, two things should happen:
% \begin{itemize}
% \item configurations in which the \textsf{heap} values are not all
%   {{false}} will be annihilated;
% \item configurations in which the \textsf{inputs} do not satisfy {{f}}
%   will cause the \textsf{satisfied?} wire to be negated and hence will
%   also be annihilated.
% \end{itemize}
% In other words, the only configurations that will survive are the ones in
% which the \textsf{inputs} satisfy {{f}}. We simply need to arrange to
% \emph{clone} these values and produce them as the output of the whole
% circuit. The final construction is therefore:

% \begin{center}
% \scalebox{1.5}{
%   \includegraphics{diagrams/sat3.pdf}
% }
% \end{center}

% To make the previous discussion concrete, we present a small, but
% complete, example. In our example, the SAT instance {{f}} is tiny: it
% takes two inputs. This function is embedded into a reversible function
% {{iso_f}} of type 
% {{((bool * bool) * bool) <-> ((bool * bool) * bool)}} where the last
% input represents the heap and the first two outputs represent the garbage. 
% The realization of {{sat_f}} given below is parametrized by such 
% a function {{iso_f}}. The inputs to {{sat_f}} are 
% \textsf{heap control}, \textsf{control}, \textsf{heap}, \textsf{input-1}, and 
% \textsf{input-2}. Its operation is simple: if the \textsf{heap} is {{true}}, 
% \textsf{heap control} is negated, and if the last output of {{iso_f}} 
% is {{false}}, \textsf{control} is negated:

% %subcode{opsem} include main
% %! columnStyle = rcl
% % sat_f &:& ((((bool * bool) * bool) * bool) * bool) <-> ((((bool * bool) * bool) * bool) * bool)
% % sat_f &=& ((swap* (*) id) (*) id) (;) 
% % && ((assocl* (*) id) (*) id) (;) 
% % &&   (((cnot (*) id) (*) id) (*) id) (;) 
% % &&   assocr* (;) 
% % &&   (assocr* (*) id) (;) 
% % &&   (swap* (*) id) (;) 
% % &&   assocr* (;) 
% % &&   (id (*) assocl*) (;) 
% % &&   (id (*) isof) (;) 
% % &&   swap* (;) 
% % &&   assocr* (;) 
% % &&   (id (*) (id (*) swap*)) (;) 
% % &&   (id (*) assocl*) (;) 
% % &&   (id (*) ((inot (*) id) (*) id)) (;)
% % &&   (id (*) (cnot (*) id)) (;) 
% % &&   (id (*) ((inot (*) id) (*) id)) (;)
% % &&   (id (*) assocr*) (;) 
% % &&   (id (*) (id (*) swap*)) (;) 
% % &&   assocl* (;)
% % &&   swap* (;)
% % &&   (id (*) Sym isof) (;) 
% % &&   (id (*) assocr*) (;) 
% % &&   assocl* (;) 
% % &&   assocl*

% Given the construction of {{sat_f}} we can build the full solver as
% follows. The overall input is the cloning heap. The combinator given
% to {{trace*}} takes the cloning heap and the inputs flowing around the
% loop and produces two copies of these inputs. One copy is produced as
% the overall output and another is fed back around the loop.

% %subcode{opsem} include main
% %! columnStyle = rcl
% % solve_f &:& bool * bool <-> bool * Bool
% % solve_f &=& trace* (
% % && (assocr* (*) id) (;) 
% % && assocr* (;)
% % && (id (*) swap*) (;)
% % && assocl* (;)
% % && (assocr* (*) id) (;)
% % && (clone2 (*) id) (;)
% % && swap* (;)
% % && (swap* (*) id) (;)
% % && ((swap* (*) id) (*) id) (;)
% % && assocl* (;)
% % && assocl* (;)
% % && (sat_f (*) id) (;)
% % && (assocr* (*) id) (;)
% % && (swap* (*) id) (;) 
% % && ((id (*) swap*) (*) id) (;)
% % && (assocl* (*) id) (;)
% % && ((id (*) swap*) (*) id))

% We can test our {{solve_f}} combinator using several SAT
% instances. Here are two possible instances. The first instance is
% satisfied by {{(false,false)}} and the
% second is satisfied by {{(false,true)}} and {{(true,true)}}.

% %subcode{opsem} include main
% %! columnStyle = rcl
% % iso_{f_1} &:& ((bool * bool) * bool) <-> ((bool * bool) * bool)
% % iso_{f_1} &=& (assocr* (;) swap* (;) toffoli (;) swap* (;) assocl*) (;)
% % && (((swap+ (*) id) (*) id) (;) toffoli (;) ((swap+ (*) id) (*) id)) (;) 
% % && (id (*) swap+)
% % 
% % iso_{f_2} &:& ((bool * bool) * bool) <-> ((bool * bool) * bool)
% % iso_{f_2} &=& toffoli

% It can indeed be verified using the semantics that the {{solve_f}}
% combinators instantiated with the SAT instances produce the expected
% results.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
\label{sec:cat}

We have introduced the idea of \emph{fractional types} in the context of a
reversible language founded on type isomorphisms and preservation of
information. Values of fractional types represent \emph{negative
  information}, a concept which is difficult to introduce in a conventional
language that allows arbitrary creation and deletion of information but which
is much simpler to deal with when the surrounding language infrastructure
guarantees preservation of information. Fractional types and values can be
used to express a simple and elegant notion of higher-order functions: a
function {{b1 :-* b2}} is a first-class value consisting of negative
information drawn from {{b1}} and positive information drawn from {{b2}}.

The interpretation of our language {{PiTF}} in the category of sets and
relations is adequate in the sense that it produces a language in which every
program is a reversible relation and in which relations are first-class
values. This relational model is however, unsatisfactory, for several
reasons:

\begin{itemize}
\item the type {{1/b}} is interpreted in the same way as {{b}}, which gives no
  insight into the ``true meaning'' of fractionals;
\item the interpretation is inconsistent with the view of types as algebraic
  structures; for example, {{Pi}} with the empty type, is the
  categorification of a semiring but although fractional types in {{PiTF}}
  syntactically ``look like'' rational numbers, there is no such formal
  connection to the rational numbers;
\item finally, we have lost some delicate structure moving from {{Pi}} to
  {{PiTF}} as we can express arbitrary relations between types and not just
  \emph{isomorphisms}.
\end{itemize}

For these reasons, it is interesting to consider other possible semantic
interpretations of fractionals. A natural alternative is to use the
``canonical'' compact closed category, that is finite dimensional vector
spaces and linear
maps~\cite{Selinger:2011:FDH:1942319.1942398,Hasegawa:2008:FDV:1805839.1805859}
over fields of characteristic $0$ (or even the category of finite dimensional
Hilbert spaces). Let us fix an arbitrary field $k$ of characteristic $0$.
Then each type {{b}} in {{PiTF}} is interpreted as a finite dimensional
vector space $\mathbf{V}_b$ over the field $k$. In particular:
\begin{itemize}
  \item every vector space contains a zero vector which means that the type
    {{0}} (if included in {{PiTF}}) would not be the ``empty'' type.
    Furthermore all combinators would be
    strict as they would have to map the zero vector to itself. 
  \item the type {{1}} is interpreted as a 1-dimensional vector space and hence
    is isomorphic to the underlying field.
  \item the fractional type {{1/b}} is interpreted as the dual vector space
    to the vector space $\mathbf{V}_b$ representing {{b}} consisting of all
    the \emph{linear functionals} on $\mathbf{V}_b$.
  \item one can then validate certain desirable properties: {{1/(1/b)}} is
    isomorphic to~{{b}}; and {{eps*_b}} corresponds to a bilinear form which
    maps a dual vector and a vector to a field element.
\end{itemize}
In such categories, the fractional type is given a non-trivial
interpretation. Indeed, while the space of column vectors is isomorphic to
the space of row vectors, they are nevertheless quite different, being
$(1,0)$ and $(0,1)$ tensors (respectively).  In other words, the category
provides a more refined model in which isomorphic negative-information values
and positive-information are not identified.

Although this semantics appears to have ``better'' properties than the
relational one, we argue that it is not yet \emph{the} ``perfect''
semantics. By including the zero vector, the language has morphisms that do
not correspond to isomorphisms (in particular it allows partial morphisms by
treating the {{0}} element of the zero-dimensional vector space as a
canonical ``undefined'' value).  It is also difficult to reconcile the
interpretation with the view that the types correspond to the (positive)
rational numbers, something we are actively seeking. What would really be a
``perfect'' interpretation is one in which we can only express
{{Pi}}-isomorphisms as first class values and in which the types are
interpreted in a way that is consistent with the rational numbers.

Fortunately, there is promising significant work on the groupoid
interpretation of type theory~\cite{Hofmann96thegroupoid} and on the 
categorification of the rational numbers~\cite{math/9802029} that may well give
us the model we desire.  The fundamental idea in both cases is that groupoids
(viewed as sets with explicit isomorphisms as morphisms) naturally have a
\emph{fractional cardinality}.  Types would be interpreted as groupoids, and
terms would be (invertible) groupoid actions.  The remaining challenge is to
identify those groupoid actions which are proper generalizations of
isomorphisms \emph{and} can be represented as groupoids, so as to obtain a
proper interpretation for a higher-order language.  As the category of
groupoids is cartesian closed, this appears eminently feasible.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

{\small
\bibliographystyle{splncs03} 
\bibliography{cites}
}
\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\appendix
\section{Categorical Background}
\label{app:cat}

We recall that a \emph{symmetric monoidal category} is a category together
with a bifunctor $\otimes$, a distinguished object $I$, and natural
isomorphisms $\alpha_{A,B,C} : (A \otimes B) \otimes C \rightarrow A \otimes
(B \otimes C)$, $\lambda_A : A \rightarrow I \otimes A$, and $\sigma_{A,B} :
A \otimes B \rightarrow B \otimes A$ subject to standard coherence
conditions~\cite{nla.cat-vn1051288}. Following common practice
(e.g.,~\cite{Selinger:2007:DCC:1229185.1229207}), we write $\rho_A =
\sigma_{I,A} \circ \lambda_A : A \rightarrow A \otimes I$.

A \emph{compact closed category} is a symmetric monoidal category where each
object $A$ is assigned a dual object $A^*$, together with a unit map $\eta_A
: I \rightarrow A^* \otimes A$ and a counit map $\epsilon_A : A \otimes A^*
\rightarrow I$, such that:
\[\begin{array}{rcl}
\lambda_A^{-1} \circ (\epsilon_A \otimes A) \circ \alpha_{A,A^*,A}^{-1} \circ (A \otimes \eta_A) \circ \rho_A &=& \mathit{id}_A \\
\rho_{A^*}^{-1} \circ (A^* \otimes \epsilon_A) \circ \alpha_{A^*,A,A^*} \circ (\eta_A \otimes A^*) \circ \lambda_A &=& \mathit{id}_{A^*}
\end{array}\]

A \emph{dagger category} is a category together with an involutive,
identity-on-objects, contravariant functor $\dagger$. Concretely, this means
that to every morphism $F : A \rightarrow B$, one associates a morphism
$f^{\dagger} : B \rightarrow A$, called the \emph{adjoint} of $f$, such that
for all $f : A \rightarrow B$ and $g : B \rightarrow C$, we have:
\[\begin{array}{rcl}
\mathit{id}^\dagger_A &=& \mathit{id}_A \\
(g \circ f)^\dagger &=& f^\dagger \circ g^\dagger \\
(f^\dagger)^\dagger &=& f
\end{array}\]

A \emph{dagger symmetric monoidal category} is a symmetric monoidal category
with a dagger structure such that the contravariant functor $\dagger$
coherently preserves the symmetric monoidal structure. Concretely, this
requirement means that for all $f : A \rightarrow B$ and $g : C \rightarrow
D$, we have:
\[\begin{array}{rcl}
(f \otimes g)^\dagger &=& f^\dagger \otimes g^\dagger \\
\alpha^\dagger_{A,B,C} &=& \alpha^{-1}_{A,B,C} \\
\lambda^\dagger_A &=& \lambda^{-1}_A \\
\sigma^\dagger_{A,B} &=& \sigma^{-1}{A,B}
\end{array}\]

\begin{definition}[Dagger Compact Closed Category]
\label{def:cat}
A \emph{dagger compact closed category} is a dagger symmetric monoidal
category that is also compact closed and such that for all $A$:
\[
\eta_A = \sigma_{A,A^*} \circ \epsilon^\dagger_A
\]
\end{definition}

\todo{biproducts}

\begin{verbatim}
homotopy equivalence???

axioms for field or meadow; actually we probably a semifield

definition of logical reversibility with relations

definition of information preservation in the case of relations; fanout as an
example which we can write

explain in detail the size of 2 * 1/2; it should have exactly one element;
there are two elemens in 2 but the 1/2 identifies them somehow; be precise

we can get the empty relation (eta ; id * swap; eps) what does that mean?

connection to vector spaces; projectors; inner product. Given dual vectors,
we can have a standalone |0><0| which is a piece of a function or a projector
if viewed by itself; we can also have |v><v'| which produces the scalars YES
or NO. We have an isomorphism between matrices and vectors by viewing 
|0><0|+|1><1| as |00>+|11>

Perhaps if restrict the ``top-level'' to non-fractional types, we can never
observe a function built from from fractionals; we must apply it and in that
case, the system overall behaves like Pi. (no empty relation for example)???
Idea suggested by Zach is to change eps to (1/a) * a * a <-> a

The code implementing the 16 relations... all 16 relations have the same
information content as () ??? So we should treat (1/bool * bool) as opaque
things; we can only extract information from them by applying them.

Include dneg, name (inv cnot) as example of programs which when fed
non-fractional values produce fractional values

Also include recipT which shows that 1/ distributes over products and tens
which shows that we can manipulate fractional types in interesting ways

Using relational semantics is good because it's clear; but it's also bad
unless we spend time to explain the operational model

Explain 'name' in detail; it IS a faithful representation of the input
relation as a value; we don't get to see the full relation though; we
interact with it by giving it inputs and observing the related ouptut come
out. If we make two copies and use each in a different context (applying each
to a different value, we can ``see'' that it's really a full relation and not
just one branch.)

total functions have size 1; partial functions have fractional sizes!!!

denotation of type in Pi is a set; in Pi/ we get a set with an equivalence
relation (Id by default) but fractionals introduce first-class equivalence
relations that can be used to restrict the sets.

---

Having re-read Baez-Dolan, I now feel better.

On 13-02-08 08:53 AM, Amr Sabry wrote:
> I think all the observations in this email are due to the fact that we
> haven't formalized the size of fractionals using homotopy equivalence. I
> started doing this but abandoned it for now but the bottom line is that
> the size of b * 1/b is 1.

Again, agree - more deeply so now.  In more detail, my current thinking:
- v \in b has size 1.  So |b| counts how many elements it has, whenever 
b is a type of Pi.  The sum and product rules apply.
- for b * 1/b to have size 1, then 1/b must have size 1/|b| (assuming 
these are independent types, which we have been assuming all along)

Attempt #1
- there are |b| elements in 1/b, and if size is additive, ( 1/|b| = 
sum_{1/v \in 1/b} |1/v| ) should hold; since |1/v| should be constant, 
this sum is |b|*c, so c = 1/|b|^2.  Weird.

Let us look at a single 'function' in bool -o bool, namely the identity 
function.  It should be 'the same' as { (1/true, true), (1/false, false) 
}.  The cardinality of that set is (1/(2^2)*1 + 1/(2^2)*1) = 1/4 + 1/4 = 
1/2.   From weird to probably-wrong.

Attempt #2
- the elements of 1/b are not distinguishable from each other (i.e. they 
are all isomorphic).  So each element 1/v has b automorphisms, so |1/b| 
= sum_{aut classes of 1/v \in 1/b} 1/|aut 1/v| = sum_{singleton} 1/|b| = 
1/|b|.  [as per p.14 of Baez-Dolan]

The size of the set { (1/true, true), (1/false, false) } is now 1/2*1 + 
1/2*1 = 1.  So this 'set' indeed represents a single "concept".

Note that in attempt #2, we get a new phenomenon.  Consider the 
(partial) relation {(1/true, false)}, also of type bool -o bool.  It has 
size 1/2 !  So, if I have not erred somewhere, only total functions have 
size 1.  Which, I must admit, I rather like.  It's not that partial 
functions are outlawed, they just don't pull as much weight [pun intended].

So 1/b is very much like a collection of |b| constraints, all of which 
are "the same" (externally).  While they may have internal structure, it 
is not visible to any of our combinators, so up to iso, there is only a 
single element of 1/b, of size 1/|b|.

Yet another way to look at it: let's assume we are looking at FinSet_0, 
where in fact our universe U has finitely (say N) things in it.  Then 
the singleton set 1 is actually the isomorphism class of {u_1, ..., 
u_N}.  There are N of them, but they have N automorphisms in a single 
orbit, so the quotient has size 1 (i.e. |S| = N but |G| = N too, so |S 
// G| = 1 ).

So an element of 1/b is closer to an action of b on b; we call it a 
constraint, or a pattern-match.  I think thinking of it as 'actions' is 
definitely the right way to go, as we will be able to have different 
kinds of actions on a type b that just those that come from the elements.

> I don't know how much of a priority this is but if it is we should try
> to work it out in detail. The good and bad news is that we will have a
> model much richer and much more complicated than relations. My thought
> when writing sec. 3 was that we would use the simple but slightly
> inadequate model of relations (which for one thing confuses v and 1/v)
> and simply mention that the right abstraction is that of 'biproduct
> dagger compact closed categories' which presumably would include the
> richer model.

I am as convinced that the right model is 'biproduct dagger compact 
closed categories' as I am that the right model for the type level is 
'categorified field'.

\end{verbatim}

\begin{comment}
Each type in {{PiTF}} is interpreted as a \emph{groupoid}. A groupoid can be
equivalently viewed as an oriented graph, a generalization of a group, or a
special category. We present the categorical view below.

A groupoid is defined by:
\begin{itemize}
\item a set of objects $x,y,\ldots$;
\item for each pair of objects, a (possibly empty) set $G(x,y)$ of morphisms
  $x \rightarrow y$; 
\item for each object $x$, there exists an identity morphism {{id_x in G(x,x)}};
\item for each triple of objects $x,y,z$, we have a function 
{{comp_{x,y,z} : G(x,y) * G(y,z) -> G(x,z)}}
\item there exists a function {{inv_{x,y} : G(x,y) -> G(y,x)}}
\end{itemize}
such that for all {{f : x -> y}}, {{g : y -> z}}, and {{h : z -> w}}, we have:
%subcode{bnf} include main
% comp_{x,x,y}(id_x,f) &=& f \\
% comp_{x,y,y}(f,id_y) &=& f \\
% comp_{x,y,z}(f, comp_{y,z,w}(g,h)) &=& comp_{x,z,w}(comp_{x,y,z}(f,g),h) \\
% comp_{y,x,y}(inv_{x,y}(f),f) &=& id_y \\
% comp_{x,y,x}(f,inv_{x,y}(f)) &=& id_x

\noindent We now proceed to explain how each {{PiTF}} type denotes a groupoid. First,
each {{Pi}} type denotes a set as follows.

\begin{definition}[Denotation of {{Pi}} Value Types {{ [[ b ]] }}]
\label{chx:def:denot}
Each {{Pi}}type denotes a finite set of values as follows:
%subcode{opsem} include main
%! <- = \leftarrow
%! union = \cup
% [[1]] '= {[ () ]}
% [[b1 + b2]] '= {[ left v ~|~ v <- [[b1]] ]} union {[ right v ~|~ v <- [[b2]] ]}
% [[b1 * b2]] '= {[ (v1, v2) ~|~ v1 <- [[b1]], v2 <- [[b2]] ]}
\end{definition}

For each of the types {{b}} above, the corresponding groupoid consists of the
set {{ [[b]] }} of elements with only the trivial identity morphisms. The
interesting groupoid is the one associated with {{1/b}}. It is defined as
follows.

Fractional types and values add considerable expressiveness to our
language. 
\end{comment}

If it were not for the cases of {{eta*}} and {{eps*}}, each combinator 
{{c : b1 <-> b2}} would define a total one-to-one function between the sets 
{{ [[ b1 ]] }} and {{ [[ b2 ]] }}. 
This is indeed the situation for the language {{Pi}} 
defined in the previous section. In the presence of fractionals, this simple 
interpretation of combinators as one-to-one functions is not sufficient%
\footnote{The interpretation of combinators as one-to-one functions is also 
not valid in the presence of recursive 
types~\cite{rc2011,James:2012:IE:2103656.2103667}.}.
The intuitive reason, is that the fission point combinator 
{{eta*_b : 1 <-> 1/b * b}} must be
able to map the element {{() : 1}} to any pair of the form {{(1/v,v)}} with
{{v:b}}. Similarly, the combinator {{eps*_b : 1/b * b <-> 1}} must be able to
accept pairs {{(1/v1,v2)}} in which the negative information {{1/v1}} does
not match the positive information {{v2}}. On such pairs, {{eps*_b}} is
undefined, i.e., it does not produce {{():1}}.

discuss two generalizations of the semantics

vector spaces (more refined because it doesn't collapse 1/b to b; also each
vector space must have a zero vector; so we have an explicit notion of
``error'' which we can use and track at run time. Adding this error
eliminates at RUN TIME relations which are not isomorphisms I think.

a potentially better and more semantics is based on groupoids; mention that
we can have sets (with structure) who cardinality is fractionals and that this
would formalize nicely the negative information