### 2013-10-13 f.tex
started a new monster paper

\documentclass[11pt]{article}
\usepackage{fullpage}
\usepackage{url}
\usepackage{color, colortbl}
\usepackage{courier}

\begin{document}
\begin{titlepage}
\begin{center}
\vfill
{\LARGE Isomorphisms as Groupoids: \\ 
  Homotopy Type Theory with Fractional Types} \\[1.5cm]
{\Large Jacques Carette${}^*$ \qquad Amr Sabry${}^{\dagger}$}\\[0.5cm]
{\Large $({}^*)$~McMaster University \qquad $({}^\dagger)$~Indiana University}
\vfill
\end{center}
\vfill
\noindent\textbf{Abstract.}

\vfill
\vfill
\end{titlepage}

\title{Isomorphisms as Groupoids: \\ 
  Homotopy Type Theory with Fractional Types}
\author{Jacques Carette \and Amr Sabry}
\date{\today}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

HoTT cannot really deal with functions. The starting point is that paths
explain and witness identities. This works for all the usual types but not
for functions or universes. For functions, the HoTT treatment is to allow
arbitrary functions, single out the isos via classical extensional methods,
and then assert via an axiom that the class of singled out functions is
equivalent to paths. This is a convoluted way. Why not start with functions
that are, by construction, isos. This was the programme we started earlier,
inspired by physical considerations. Because physics requires various
conservation principles (including conservation of information) and because
computation is fundamentally a physical process, the argument (originally due
to Feynman etc.) was that computation should be based on reversible
processes, or in abstract terms, on isomorphisms. That fits very well with
the HoTT philosophy that makes isos first-class. Except that HoTT does not
deal exclusively with isos, the functions are generally not isos. So let's
investigate the radical idea that all functions are isos in the HoTT
framework. What we get is a notion of first-class functions that by
construction are isos and obey the usual path rules etc. The technical device
is the notion of fractional types which are dual to products. A function
(iso) $A \rightarrow B$ is represented as an element of the type
$(\frac{1}{A} \times B)$. The construction shows the essential nature of the
$\infty$-groupoid structure. The more higher-order functions we have the
higher we have to go in the groupoid structure. We also have something
significant in the realm of reversible computing. We finally have
higher-order functions. Of course we need to show that we have constructed a
monoidal closed category (a concrete one) and then we can inherit all the
fancy constructions. Finally we have a very interesting operational
interpretation of paths and we outline various applications.

Need a clear explanation of why we can't deal with the usual functions that
are not isomorphisms. Regular function spaces obey $1^A = 1$ for all $A$. In
terms of fractionals this gives $\frac{1}{A} = 1$ for all $A$ which is not
true. However if we restrict to isomorphisms then the only functions we can
write are $A \rightarrow B$ where $A$ and $B$ are isomorphic types and hence
the only functions we can write in the space $1^A$ are to types $A$ that are
isomorphic to 1 and in that case we do indeed get that $\frac{1}{A} =
1$. See~\cite{fiore-remarks} for more details. The benefit is that we fit
squarely in the physically motivated notion that isomorphisms are the
foundational computational mechanism as well as the HoTT approach that is
essentially all about isomorphisms. Of course we don't lose anything by
restricting our attention to isomorphisms because we know that general
(irreversible) functions can be obtained from isomorphisms via
\emph{information effects}~\cite{James:2012:IE:2103656.2103667}.

We are witnessing a convergence of ideas from several distinct research
communities (physics, mathematics, and computer science) towards replacing
\emph{equalities} by \emph{isomorphisms}. The combined programme has sparked
a significant amount of research that unveiled new and surprising connections
between geometry, algebra, logic, and computation (see~\cite{baez2011physics}
for an overview of some of the connections).

In the physics community, Landauer~\cite{Landauer:1961,Landauer},
Feynman~\cite{springerlink:10.1007/BF02650179}, and others have interpreted
the laws of physics as fundamentally related to computation. The great
majority of these laws are formulated as equalities between different
physical observables which is unsatisfying: \emph{different} physical
observables should not be related by an \emph{equality}. It is more
appropriate to relate them by an \emph{isomorphism} that witnesses, explains,
and models the process of transforming one observable to the other.

In the mathematics and logic community, Martin-L\"of developed an extension
of the simply typed $\lambda$-calculus originally intended to provide a
rigorous framework for constructive
mathematics~\cite{citeulike:7374951}. This theory has been further extended
with \emph{identity types} representing the proposition that two terms are
``equal.'' (See~\cite{streicher,warren} for a survey.) Briefly speaking,
given two terms $a$ and $b$ of the same type $A$, one forms the type
$\texttt{Id}_A(a,b)$ representing the proposition that~$a$ and~$b$ are equal:
in other words, a term of type $\texttt{Id}_A(a,b)$ witnesses, explains, and
models the process of transforming $a$ to $b$ and vice-versa.

In the computer science community, the theory and practice of type
isomorphisms is well-established. Originally, such type isomorphisms were
motivated by the pragmatic concern of searching large libraries of functions
by providing one of the many possible isomorphic types for the desired
function~\cite{Rittri:1989:UTS:99370.99384}. More recently, type isomorphisms
have taken a more central role as \emph{the} fundamental computational
mechanism from which more conventional, i.e., irreversible computation, is
derived. In our own previous
work~\cite{James:2012:IE:2103656.2103667,rc2011,rc2012} we started with the
notion of type isomorphism and developed from it a family of programming
languages, $\Pi$ with various superscripts, in which computation is an
isomorphism preserving the information-theoretic entropy.

A major open problem remains, however: a higher-order extension of
$\Pi$. This extension is of fundamental importance in all the originating
research areas. In physics, it allows for quantum states to be viewed as
processes and processes to be viewed as states, such as with the
Choi-Jamiolkowski
isomorphism~\cite{choi1975completely,jamiolkowski1972linear}.  In mathematics
and logic, it allows the equivalence between different proofs of type
$\texttt{Id}_A(a,b)$ to itself be expressed as an isomorphism (of a higher
type) $\texttt{Id}_{\texttt{Id}_A(a,b)}(p,q)$. Finally, in computer science,
higher-order types allow code to abstract over other code fragments as well
as the manipulation of code as data and data as code.

Technically speaking, obtaining a higher-order extension requires the
construction of a \emph{closed category} from the underlying monoidal
category for~$\Pi$. Although the general idea of such a construction is
well-understood, the details of adapting it to an actual programming language
are subtle.  Our main novel technical device to achieving the higher-order
extension is a \emph{fractional type} which represents \emph{negative
information} and which is so named because of its duality with conventional
product types.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Intuitive Ideas and Examples} 
\label{sec:intuition}

Talk about groupoid cardinality and give an example of how the cardinality of
a group of order $n$ is $\frac{1}{n}$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Finite Types as Discrete (0-) Groupoids}

A review of pi and slight reformulation of the semantics to use the groupoid
terminology. In more detail, motivate and introduce pointed types here, add
an element $\bot_A$ to every type $A$ and express the groupoids using a
pointed type and an explicit collection of discrete paths (not indexed by
anything at this point).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Fractional Types as Groupoids}

In the previous section, every type was modeled by a collection of values
(including a special ``unusable'' value $\bot$) and a collection of trivial
paths. In this section, non-trivial paths will emerge to model ``fractional
values.'' The intuition is that two values that that are connected by a path
should not be counted twice as the path induces an equivalence between these
two values. As illustrated in Sec.~\ref{sec:intuition}, this idea generalizes
to conclude that a value connected to iself using several paths corresponds
to a ``fraction.''

We begin by investigating the special situation of a type $\frac{1}{A}$ where
$A$ is a 0-type (i.e., where $A$ does not contain any fractional components).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Rank-1 Isomorphisms as 1-Groupoids}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Rank-$n$ Isomorphisms as $n$-Groupoids}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Operational Semantics via Backtracking Constraints}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Connections to Type Theory}

\begin{quote} 
As usual, something funny happens at the left of the arrow. John Reynolds,
Aarhus, June 27, 2000.~\cite{danvypage} 
\end{quote}

No more! We finally have an inductive characterization of functions
(isomorphisms). No negative occurrences of types. No positivity
checks. Nothing funny.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Connections to Semantics}

Now that functions are a proper inductive type, we can use conventional
induction to reason about functions. There are also probably connections to
logical relations and parametricity that are essentially some kind of
isomorphism which might be expressible as a higher-order path.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Connections to Logic}

Similar to double negation proof of $A \vee \neg A$. Make a guess and if
wrong backtrack and choose the other branch.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Connections to Programming Practice}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Connections to Resource Interpretation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Connections to Category Theory}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Connections to Homotopy Theory}

The connection between type theory and homotopy theory raises some
interesting questions. What do the various shapes (spheres, tori, etc.) have
to go with programming? Our answer is that they represent a class
of \emph{functions}: isomorphisms.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Connections to Quantum Computing}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Connections to Physics, Information Theory, and Reversible Computing}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions and Future Work}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgments}

This material is based upon work supported by the National Science Foundation
under Grant No. 1217454. We would like to specially thank Roshan James and
Zachary Sparks for their contributions to early stages of this work.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{plain}
\bibliography{cites}
\end{document}