### 2013-10-13 f.tex
small updates to f.tex

\documentclass[11pt]{article}
\usepackage{fullpage}
\usepackage{url}
\usepackage{color, colortbl}
\usepackage{courier}
\usepackage{graphicx}

\begin{document}
\begin{titlepage}
\begin{center}
\vfill
{\LARGE Isomorphisms as Groupoids: \\ 
  Homotopy Type Theory with Fractional Types} \\[1.5cm]
{\Large Jacques Carette${}^*$ \qquad Amr Sabry${}^{\dagger}$}\\[0.5cm]
{\Large $({}^*)$~McMaster University \qquad $({}^\dagger)$~Indiana University}
\vfill
\end{center}
\vfill
\noindent\textbf{Abstract.}

\vfill
\vfill
\end{titlepage}

\title{Isomorphisms as Groupoids: \\ 
  Homotopy Type Theory with Fractional Types}
\author{Jacques Carette \and Amr Sabry}
\date{\today}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

HoTT cannot really deal with functions. The starting point is that paths
explain and witness identities. This works for all the usual types but not
for functions or universes. For functions, the HoTT treatment is to allow
arbitrary functions, single out the isos via classical extensional methods,
and then assert via an axiom that the class of singled out functions is
equivalent to paths. This is a convoluted way. Why not start with functions
that are, by construction, isos. This was the programme we started earlier,
inspired by physical considerations. Because physics requires various
conservation principles (including conservation of information) and because
computation is fundamentally a physical process, the argument (originally due
to Feynman etc.) was that computation should be based on reversible
processes, or in abstract terms, on isomorphisms. That fits very well with
the HoTT philosophy that makes isos first-class. Except that HoTT does not
deal exclusively with isos, the functions are generally not isos. So let's
investigate the radical idea that all functions are isos in the HoTT
framework. What we get is a notion of first-class functions that by
construction are isos and obey the usual path rules etc. The technical device
is the notion of fractional types which are dual to products. A function
(iso) $A \rightarrow B$ is represented as an element of the type
$(\frac{1}{A} \times B)$. The construction shows the essential nature of the
$\infty$-groupoid structure. The more higher-order functions we have the
higher we have to go in the groupoid structure. We also have something
significant in the realm of reversible computing. We finally have
higher-order functions. Of course we need to show that we have constructed a
monoidal closed category (a concrete one) and then we can inherit all the
fancy constructions. Finally we have a very interesting operational
interpretation of paths and we outline various applications.

Need a clear explanation of why we can't deal with the usual functions that
are not isomorphisms. Regular function spaces obey $1^A = 1$ for all $A$. In
terms of fractionals this gives $\frac{1}{A} = 1$ for all $A$ which is not
true. However if we restrict to isomorphisms then the only functions we can
write are $A \rightarrow B$ where $A$ and $B$ are isomorphic types and hence
the only functions we can write in the space $1^A$ are to types $A$ that are
isomorphic to 1 and in that case we do indeed get that $\frac{1}{A} =
1$. See~\cite{fiore-remarks} for more details. The benefit is that we fit
squarely in the physically motivated notion that isomorphisms are the
foundational computational mechanism as well as the HoTT approach that is
essentially all about isomorphisms. Of course we don't lose anything by
restricting our attention to isomorphisms because we know that general
(irreversible) functions can be obtained from isomorphisms via
\emph{information effects}~\cite{James:2012:IE:2103656.2103667}.

We are witnessing a convergence of ideas from several distinct research
communities (physics, mathematics, and computer science) towards replacing
\emph{equalities} by \emph{isomorphisms}. The combined programme has sparked
a significant amount of research that unveiled new and surprising connections
between geometry, algebra, logic, and computation (see~\cite{baez2011physics}
for an overview of some of the connections).

In the physics community, Landauer~\cite{Landauer:1961,Landauer},
Feynman~\cite{springerlink:10.1007/BF02650179}, and others have interpreted
the laws of physics as fundamentally related to computation. The great
majority of these laws are formulated as equalities between different
physical observables which is unsatisfying: \emph{different} physical
observables should not be related by an \emph{equality}. It is more
appropriate to relate them by an \emph{isomorphism} that witnesses, explains,
and models the process of transforming one observable to the other.

In the mathematics and logic community, Martin-L\"of developed an extension
of the simply typed $\lambda$-calculus originally intended to provide a
rigorous framework for constructive
mathematics~\cite{citeulike:7374951}. This theory has been further extended
with \emph{identity types} representing the proposition that two terms are
``equal.'' (See~\cite{streicher,warren} for a survey.) Briefly speaking,
given two terms $a$ and $b$ of the same type $A$, one forms the type
$\texttt{Id}_A(a,b)$ representing the proposition that~$a$ and~$b$ are equal:
in other words, a term of type $\texttt{Id}_A(a,b)$ witnesses, explains, and
models the process of transforming $a$ to $b$ and vice-versa.

In the computer science community, the theory and practice of type
isomorphisms is well-established. Originally, such type isomorphisms were
motivated by the pragmatic concern of searching large libraries of functions
by providing one of the many possible isomorphic types for the desired
function~\cite{Rittri:1989:UTS:99370.99384}. More recently, type isomorphisms
have taken a more central role as \emph{the} fundamental computational
mechanism from which more conventional, i.e., irreversible computation, is
derived. In our own previous
work~\cite{James:2012:IE:2103656.2103667,rc2011,rc2012} we started with the
notion of type isomorphism and developed from it a family of programming
languages, $\Pi$ with various superscripts, in which computation is an
isomorphism preserving the information-theoretic entropy.

A major open problem remains, however: a higher-order extension of
$\Pi$. This extension is of fundamental importance in all the originating
research areas. In physics, it allows for quantum states to be viewed as
processes and processes to be viewed as states, such as with the
Choi-Jamiolkowski
isomorphism~\cite{choi1975completely,jamiolkowski1972linear}.  In mathematics
and logic, it allows the equivalence between different proofs of type
$\texttt{Id}_A(a,b)$ to itself be expressed as an isomorphism (of a higher
type) $\texttt{Id}_{\texttt{Id}_A(a,b)}(p,q)$. Finally, in computer science,
higher-order types allow code to abstract over other code fragments as well
as the manipulation of code as data and data as code.

Technically speaking, obtaining a higher-order extension requires the
construction of a \emph{closed category} from the underlying monoidal
category for~$\Pi$. Although the general idea of such a construction is
well-understood, the details of adapting it to an actual programming language
are subtle.  Our main novel technical device to achieving the higher-order
extension is a \emph{fractional type} which represents \emph{negative
information} and which is so named because of its duality with conventional
product types.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Intuitive Ideas and Examples} 
\label{sec:intuition}

Talk about groupoid cardinality and give an example of how the cardinality of
a group of order $n$ is $\frac{1}{n}$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Finite Types as Discrete (0-)Groupoids}

A review of pi and slight reformulation of the semantics to use the groupoid
terminology. In more detail, motivate and introduce pointed types here, add
an element $\bot_A$ to every type $A$ and express the groupoids using a
pointed type and an explicit collection of discrete paths (not indexed by
anything at this point).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Fractional Types as Groupoids}

In the previous section, every type was modeled by a collection of values
(including a special ``unusable'' value $\bot$) and a collection of trivial
paths. In this section, non-trivial paths will emerge to model ``fractional
values.'' The intuition is that two values that are connected by a path
should not be counted twice as the path induces an equivalence between these
two values: ``they are the same.'' As illustrated in
Sec.~\ref{sec:intuition}, this idea generalizes to conclude that a value
connected to itself using several paths corresponds to a ``fraction.''

We begin by investigating the special situation of a type $\frac{1}{A}$ where
$A$ is a 0-type (i.e., where $A$ does not contain any fractional components).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Rank-1 Isomorphisms as 1-Groupoids}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Rank-$n$ Isomorphisms as $n$-Groupoids}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Operational Semantics via Backtracking Constraints}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Connections to Type Theory}

\begin{quote} 
As usual, something funny happens at the left of the arrow. John Reynolds,
Aarhus, June 27, 2000.~\cite{danvypage} 
\end{quote}

No more! We finally have an inductive characterization of functions
(isomorphisms). No negative occurrences of types. No positivity
checks. Nothing funny.

There are also deep connections with algebraic types, the seven trees in one,
and the Fiore papers.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Connections to Semantics}

Now that functions are a proper inductive type, we can use conventional
induction to reason about functions. There are also probably connections to
logical relations and parametricity that are essentially some kind of
isomorphism which might be expressible as a higher-order path.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Connections to Logic}

Similar to double negation proof of $A \vee \neg A$. Make a guess and if
wrong backtrack and choose the other branch.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Connections to Programming Practice}

Our construction can be interpreted in the context of speculative concurrent
programming. A thread can at any time create, out of nothing, a value of type
$A$ and proceed with it, generating in the process a value of type
$\frac{1}{A}$ that a concurrent process can be manipulate. When an actual
value of type $A$ is produced, the fractional value is used to coordinate a
backtracking mechanism that communicates the actual value of type $A$ to the
original process. This is very similar to how we use logical variables and
unification.

Because we are \emph{not} in the context of the full $\lambda$-calculus,
which allows arbitrary duplication and erasure of information, values of
fractional types are first-class values that can flow anywhere. The
information-preserving computational infrastructure guarantees that, in a
complete program, every constraint imposed by a fractional value will also be
satisfied exactly once. This property is shared with systems that are based
on linear logic.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Connections to Reversible Computing}

We have higher-order reversible computations.

Consider a similar algebraic manipulation involving fractional types.
\begin{center}
\scalebox{1.1}{
\includegraphics{diagrams/thesis/algebra-wire2.pdf}
}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Connections to Category Theory}

We have constructed a concrete closed monoidal category. We did not use the
Int-construction which requires an existing trace operator. We inherit a lot
of constructions for free: name, coname, trace, etc.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Connections to Resource Interpretation}

We can interpret $\times$ as parallel computations (i.e., as computations
occurring simultaneously in different regions in space). So a value of type
$A \times B$ consists of two values occupying two different space locations,
e.g., two different memory locations. With that perspective, a value of type
$\frac{1}{A}$ consists of a small dictionary that stores all the values of
type $A$ and can shuffle them around.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Connections to Homotopy Theory}

The connection between type theory and homotopy theory raises some
interesting questions. What do the various shapes (spheres, tori, etc.) have
to go with programming? Our answer is that they represent a class
of \emph{functions}: isomorphisms.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Connections to Physics and Quantum Computing}

The dual vector space in quantum mechanics and its role in quantum
computing. The fact that $\eta$ and $\epsilon$ are dual is essentially the
teleportation algorithm. We can give it a computational interpretation in
terms of backtracking.

One understanding of quantum computing is that it exploits the laws of physics
to build faster machines (perhaps). Another more foundational understanding
is that it provides a computational interpretation of physics, and in
particular directly addresses the question of interpretation of quantum
mechanics. In a little known document, Rozas~\cite{Rozas:1987:CMO:889539}
uses continuations to implement the transactional interpretation of quantum
mechanics~\cite{transactional} which includes as its main ingredient a
fixpoint calculation between waves or particles traveling forwards and
backwards in time. Our work sheds no light on whether this interpretation is
the ``right one'' but it is interesting that we can directly realize it using
the primitives of our language.

The multiplicative structure has a direct connection to entangled quantum
particles, or perhaps entangled particles and anti-particles.  The idea of
entanglement, that an action on one particle is ``instantaneously''
communicated to the other, is analogous to how unifying one value affects its
dual pair which is possibly in another part of the computation.  Again our
model sheds no light on whether this is related to how nature computes but it
is again interesting that we can directly realize the idea.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Connections to Information Theory}

We have given classic computational interpretations to negative
information~\cite{negative-information} and negative
entropy~\cite{negative-entropy}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Speculative Connections}

Negative types, square root types, and imaginary types have also appeared in
the literature: we proceed with a simple explanation. We have so far extended
the set of types to be the positive rational numbers. Now we will push this
and extend the set of types to algebraic numbers. In other words, we will
allow datatypes defined by arbitrary polynomials and allow the roots of such
polynomials to be types.

Consider first an example in which we want to compute with the sides of a
rectangle whose area is 91 and whose length is longer than its width by 6
units. One can solve the quadratic equation to determine that the sides are 7
and 13 and proceed. This however prematuraly forces us to globally solve the
constraint. Instead we can let the two sides of the rectangle be $x$ and
$x+6$ and use the following equation to capture the desired constraint:
\[
x^2 + 6x - 91 = 0
\]
The equation introduces an isomorphism between the type $x^2 + 6x - 91$ and
the type $0$. We can now proceed to compute with the unknown $x$, being
assured that in a closed program, our computation will eventually be
consistent with the solution of the algebraic equation.

Previously the most famous example of a similar nature is the puzzling
isomorphism that one can establish between seven binary trees and one.
A binary tree is defined by the datatype
\[
x = 1 + x * x 
\]
which can be rearranged to the polynomial equation $x^2 - x + 1 = 0$. By
algebraic manipulation we can reason as follows:
\[\begin{array}{rclclclcl}
x^3 &=& x^2 x &=& (x-1) x &=& x^2 - x &=& -1 \\
x^6 &=& 1 \\
x^7 &=& x^6 x &=& x
\end{array}\]
Fiore poses the question of why such algebraic manipulation would make sense
type theoretically but states that even though some of the intermediate steps
make no sense, the final equivalence is valid and can be used to actually
construct an isomorphism between $x^7$ the type of seven binary trees 
and $x$ the type of binary trees.

Discussion of possible polynomials:
\begin{itemize}
\item $b=1+1$. Boring.
\item $b=b+1$. That introduces the natural numbers. No solution for this
  polynomial over the algebraic numbers. We must extend the numbers with
  $\omega$ to get a solution. We reject this in this paper and prefer to
  stick to algebraic numbers. The advantage is that all the isomorphisms are
  valid numerically. With the above type we could subtract $b$ from both
  sides to show that $0=1$ which is nonsense. We can however have infinite
  types as long as they have algebraic solutions.
\item $b^2=2$ or $b = \pm \sqrt{2}$. We have introduced the square root of a
  boolean! If we had superpositions, we could then write a function that
  performs the square root of negation in the sense that applying it twice
  would be equivalent to boolean negation.
\end{itemize}

\paragraph*{Geometry of Interaction (GoI).}
Geometry of Interaction was developed by Girard~\cite{girard1989geometry} as
part of the development of linear logic. It was given a computational
interpretation by Abramsky and
Jagadeesan~\cite{Abramsky:1994:NFG:184662.184664}, and was developed into a
reversible model of computing by
Mackie~\cite{Mackie2011,DBLP:conf/popl/Mackie95}. Preliminary investigations
suggest that many of the GoI machine constructions can be simulated by
treating Mackie's bi-directional wires as pairs of wires and replacing the
machine's global state with a typed value on the wire that captures the
appropriate state. This connection is exciting because when viewed through a
Curry-Howard lens it suggests that the logical interpretation of would be a
linear-like logic with a notion of resource preservation and with a natural
computational interpretation.

\paragraph*{Computing in the Field of Algebraic Numbers.}
\label{sec:algebraic-field}
Algebraically, the addition of fractional types corresponds to a move from a
ring-like structure to a full field. Our language captures the structure of
one particular field: that of the rational numbers. As we have seen,
computation in this field is quite expressive and interesting and yet, it has
two fundamental limitations. First it cannot express any recursive type, and
second it cannot express any datatype definitions. We believe these to be two
orthogonal extensions: recursive types were considered in our previous
paper~\cite{James:2012:IE:2103656.2103667}; arbitrary datatypes are however
even more exciting that plain rationals as each datatype definition can be
viewed as a polynomial (see below) which essentially means that we start
computing in the field of algebraic numbers, which includes square roots and
imaginary numbers. As crazy as it might seem, the type $\sqrt{2}$ and even
the type $(1/2)+i(\sqrt{3}/2)$ ``make sense.''  In fact the latter type is
the solution to the polynomial $x^2-x+1=0$ which if re-arranged looks like
$x=1+x*x$ and perhaps more familiarly as the datatype of binary trees $\mu
x.(1+x*x)$. These types happen to have been studied extensively following a
paper by Blass~\cite{seventrees} which used the above datatype of trees to
infer an isomorphism between seven binary trees and one!

We have confirmed that we can extend our language with the datatype
declaration for binary trees and build a witness for this isomorphism that
works as expected. However not every isomorphism constructed from algebraic
manipulation is computationally meaningful. To understand the issue in more
detail, consider the following algebraically valid proof of the isomorphism
in question:
\[\begin{array}{rclclclcl}
x^3 &=& x^2 x &=& (x-1) x &=& x^2 - x &=& -1 \\
x^6 &=& 1 \\
x^7 &=& x^6 x &=& x
\end{array}\]
The question is why such an algebraic manipulation makes sense type
theoretically, even though the intermediate step asking for an isomorphism
between $x^6$ and $1$ has no computational context. In our setting, this
isomorphism can be constructed but it diverges on all inputs (in both
ways). This suggests that, in the field of algebraic numbers, some algebraic
manipulations are somehow ``more constructive'' than others.

A related issue is that not all meaningful recursive types are meaningful
polynomials. For instance $\textsf{nat}=\mu x.(1+x)$ implies the polynomial
$x=1+x$ which has no algebraic solutions without appeal to more complex
structures with limits etc.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions and Future Work}

This is cool and deep.

\paragraph*{Declarative Continuations.} 
In his Masters thesis~\cite{Filinski:1989:DCI:648332.755574}, Filinski
proposes that continuations are a \emph{declarative} concept. He,
furthermore, introduces a symmetric extension of the $\lambda$-calculus in
which call-by-value is dual to call-by-name and values are dual to
continuations. In more detail, the symmetric calculus contains a ``value''
fragment and a ``continuation'' fragment which are mirror images. Pairs and
sums are treated as duals in the sense that the ``value'' fragment includes
pairs whose mirror image in the ``continuation'' fragment are sums. In
contrast, our language includes pairs and sums in the value fragment and two
symmetries: one that maps the pairs to fractions and another that maps the
sums to subtractions.

\paragraph*{The Duality of Computation.}
The duality between call-by-name and call-by-value was further investigated
by Selinger using control
categories~\cite{Selinger:2001:CCD:966910.966911}. Curien and
Herbelin~\cite{Curien:2000} also introduce a calculus that exhibits
symmetries between values and continuations and between call-by-value and
call-by-name. The calculus includes the type $A-B$ which is the dual of
implication, i.e., a value of type $A-B$ is a context expecting a function of
type $A \rightarrow B$. Alternatively a value of type $A-B$ is also explained
as a \emph{pair} consisting of a value of type $A$ and a continuation of type
$B$. This is to be contrasted with our interpretation of a value of that type
as \emph{either} a value of $A$ or a demand for a value of type $B$. This
calculus was further analyzed and extended by
Wadler~\cite{Wadler:2003,DBLP:conf/rta/Wadler05}. The extension gives no
interpretation to the subtraction connective and like the original symmetric
calculus of Filinski, introduces a duality that relates sums to products and
vice-versa.

\paragraph*{Subtractive Logic.} 
Rauszer~\cite{springerlink:10.1007/BF02120864,rauszer,rauszer2} introduced a
logic which contains a dual to implication. Her work has been distilled in
the form of \emph{subtractive logic}~\cite{Crolard01} which has recently been
related to coroutines~\cite{Crolard01082004} and delimited
continuations~\cite{Ariola:2009:TFD:1743339.1743381}.  In more detail,
Crolard explains the type $A-B$ as the type of \emph{coroutines} with a local
environment of type $A$ and a continuation of type $B$. The description is
complicated by what is essentially the desire to enforce linearity
constraints so that coroutines cannot access the local environment of other
coroutines.

\paragraph*{Negation in Classical Linear Logic} 
Filinski~\cite{Filinski92} uses the negative types of linear logic to model
continuations. Reddy~\cite{Reddy91} generalizes this idea by interpreting the
negative types of linear logic as \emph{acceptors}, which are like
continuations in the sense that they take an input and return no
output. Acceptors however are also similar in flavor to logic variables:
they can be created and instantiated later once their context of use is
determined. Although a formal connection is lacking, it is clear that, at an
intuitive level, acceptors are entities that combine elements of our negative
and fractional types.

\paragraph*{The Lambek-Grishin Calculus.} The ``parsing-as-deduction'' style
of linguistic analysis uses the Lambek-Grishin calculus with the following
types: product, left division, right division, sum, right difference, and
left difference~\cite{Bernardi:2010:CSL:1749618.1749689}. The division and
difference types are similar to our types but because the calculus lacks
commutativity and associativity and only has limited notions of
distributivity, each connective needs a left and right version. The
Lambek-Grishin exhibits two notions of symmetry but they are unrelated to our
notions. In particular, the first notion of symmetry expresses commutativity
and the second relates products to sums and divisions to subtractions. In
contrast, our two symmetries relate sums to subtractions and products to
divisions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgments}

This material is based upon work supported by the National Science Foundation
under Grant No. 1217454. We would like to specially thank Roshan James and
Zachary Sparks for their contributions to early stages of this work.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{plain}
\bibliography{cites}
\end{document}