### 2014-02-14 Univalence/tex/theorems.tex
added explanation of theorems in slightly more legible LaTeX form

\documentclass[11pt]{article}
\usepackage{fullpage}
\usepackage{latexsym}
\usepackage{verbatim}
\usepackage{amsthm,amssymb,amsmath,stmaryrd}
\usepackage{ifthen}
\usepackage{url}

\input{defs.tex}

\title{Univalence Theorems with $\Pi$}

\author{Zach Sparks}

%% \newcommand{\code}[1]{\texttt{#1}}

\begin{document}
\maketitle

We are trying to prove theorems that relate syntactic functions in our system
(ie, combinators) with semantic functions (ie, vectors). That is:

For all finite types $a$ and $b$ and all combinators $c : a \combarrow b$, there
exists some semantic function $f : \tonat{a} \to \tonat{b}$ such that for all $x
: a$, $\tonat{c\,x} \eq v\,\tonat{x}$.

For all finite types $a$ and $b$ and all semantic functions $f : \tonat{a} \to
\tonat{b}$, there exists some combinator $c : a \combarrow b$ such that for all
$i : \tonat{a}$, $c\,\fromnat{a}{i} \eq \fromnat{b}{v\,i}$.

In our case, we already know the transformations we are using (called $compile$
and $decompile$), so we can rewrite these theorems as follows:

For all finite types $a$ and $b$, combinators $c : a \combarrow b$, and all $x :
a$, $\tonat{c\,x} \eq compile\,c\,\tonat{x}$.

For all finite types $a$ and $b$, all semantic functions $f : \tonat{a} \to
\tonat{b}$, and all $i : \tonat{a}$, $decompile\,v\,\fromnat{a}{i} \eq
\fromnat{b}{v\,i}$.

As an intermediate step, we are proving a slightly different theorem about
combinators over \emph{normalized types}, since these are closer to the vector
representation we are compiling to. Thus, the process looks something like:

\begin{verbatim}
combinator <===> combinator over normalized types <===> vector
\end{verbatim}

Right now, we are working on the arrow on the right; the arrow on the left
should be somewhat easier.

\end{document}