### 2014-05-06 negatives.tex
intro to numbers

\documentclass[authoryear,preprint]{sigplanconf}
\usepackage{amsmath}
\usepackage{listings} 
\usepackage{stmaryrd}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{xcolor}
\usepackage{courier}
\usepackage{thmtools}
\usepackage{bbold}

%%

\newcommand{\game}[2]{\langle #1 ~|~ #2 \rangle}
\newcommand{\zerog}{$\mathbb{0}$}
\newcommand{\oneg}{$\mathbb{1}$}
\newcommand{\twog}{$\mathbb{2}$}
\newcommand{\threeg}{$\mathbb{3}$}
\newcommand{\fourg}{$\mathbb{4}$}

%%

\begin{document}
\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\newcommand{\alt}{~|~}
\lstnewenvironment{code}{\lstset{basicstyle={\sffamily\footnotesize}}}{}

\lstset{frame=none,
         language=Haskell,
         basicstyle=\sffamily, 
         numberstyle=\tiny,
         numbersep=5pt,
         tabsize=2,    
         extendedchars=true,
         breaklines=true,   
         breakautoindent=true,
         keywordstyle=\color{black},
         captionpos=b,
         stringstyle=\color{black}\ttfamily,
         showspaces=false,  
         showtabs=false,    
         framexleftmargin=2em,
         framexbottommargin=1ex,
         showstringspaces=false
         basicstyle=\sffamily,
         columns=[l]flexible,
         flexiblecolumns=true,
         aboveskip=\smallskipamount,
         belowskip=\smallskipamount,
         lineskip=-1pt,
         xleftmargin=1em,
         escapeinside={/+}{+/},
         keywords=[1]{Monad,Just,Nothing,type,data,right,left,id,where,do,
                     if,then,else,let,in},
         literate={+*}{{$+*$}}1 {+}{{$\;+\;$}}1 {/}{{$/$}}1 {*}{{$\;*\;$}}1
           {=}{{$=\ $}}1 {/=}{{$\not=$}}2
           {[]}{$[\;]$}2
           {<}{{$<$}}1 {>}{{$\rangle$}}1 
           {++}{{$+\!\!\!+\;$}}1 {::}{{$:\mkern -2.5mu:\;$}}1
           {&&}{{$\&\!\!\!\&$}}2
           {:=:}{{$:\mkern -2mu=\mkern -2mu:\;$}}2
           {:+:}{{$:\mkern -5mu+\mkern -5mu:\;$}}2
           {:-:}{{$:\mkern -5mu-\mkern -5mu:\;$}}2
           {:*:}{{$:\mkern -5mu*\mkern -5mu:\;$}}2
           {$}{{\texttt{\$}\hspace{0.5em}}}1
           {`}{$^\backprime$}1
           {==}{{$=\!=\;$}}2
           {===}{{$\equiv\;$}}2
           {->}{{$\rightarrow\;$}}2 {>=}{{$\geq$}}2 {<-}{{$\leftarrow$}}2
           {<=}{{$\leq$}}2 {=>}{{$\Rightarrow\;$}}2
           {<<}{{$\ll$}}2 {>>}{{$\gg\;$}}2
           {>>>}{{$\ggg\;$}}3 {<<<}{{$\lll\;$}}3
           {>>=}{{$\gg\mkern -2.5mu=\;$}}3
           {=<<}{{$=\mkern -2.5mu\ll\;$}}3
           {<|}{$\lhd\;$}2
           {<||}{$\unlhd\;$}2
           {\ ||\ }{$\|$}1
           {\\}{$\lambda$}1
           {:>}{{$\rhd$}}2
           {||>}{{$\unrhd$}}2
           {_}{{$\_$}}1
           {_B}{{$_b$}}2
           {forall}{{$\forall$}}1
}

\lstset{postbreak=\raisebox{0ex}[0ex][0ex]
        {\ensuremath{\hookrightarrow}}}
\lstset{breaklines=true, breakatwhitespace=true}
\lstset{numbers=none, numbersep=5pt, stepnumber=2, numberstyle=\scriptsize}
\lstset{rangeprefix=/*!\ , rangesuffix=\ !*\/, includerangemarker=false}

%% double-blind reviewing...
\title{Negative Types}
\authorinfo{}{}{}
\maketitle

\begin{abstract}
\ldots
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conway Numbers}

A Conway number is a game consisting of left and right options where each
option models a move to another game. Players alternate taking options and
the player with no available options loses. In Haskell, one might define the
datatype of Conway numbers as:
\begin{code}
data ConwayNumber = CN [ConwayNumber] [ConwayNumber]
\end{code}
The simplest Conway number is \lstinline|CN [] []| with empty left and right
options. We call this number \zerog: 
\begin{code}
/+ \zerog +/  = CN  [] []
\end{code}
Once we have defined \zerog, we can also define the following:
\begin{code}
/+ \oneg +/  = CN  [ /+ \zerog +/ ]  []
/+ \twog +/  = CN  [ /+ \oneg +/ ]  []
/+ \threeg +/  = CN  [ /+ \twog +/ ]  []
\end{code}
and so on. Intuitively, the number $\mathbb{n}$ represents a game in which
the left player has an $n$-move advantage over the right player. Dually, we
can define the following numbers where the left player has $n$-move
\emph{disadvantage} over the right player:
\begin{code}
/+ -\oneg +/  = CN  [] [ /+ \zerog +/ ]
/+ -\twog +/  = CN  [] [ /+ -\oneg +/ ]
/+ -\threeg +/  = CN  [] [ /+ -\twog +/ ]
\end{code}
More generally, the \emph{unary negation} of a number is defined as follows:
\begin{code}
neg :: ConwayNumber -> ConwayNumber
neg (CN xls xrs) = 
  CN [ neg xr | xr <- xrs ] [ neg xl | xl <- xls ]
\end{code}

Conway numbers also come equipped with addition and multiplication operations
defined as follows:

\begin{code}
(:+:) :: ConwayNumber -> ConwayNumber -> ConwayNumber
x@(CN xls xrs) :+: y@(CN yls yrs) = 
  CN 
    ([ xl :+: y | xl <- xls ] `union`
     [ x :+: yl | yl <- yls ])
    ([ xr :+: y | xr <- xrs ] `union`
     [ x :+: yr | yr <- yrs ])

(:-:) :: ConwayNumber -> ConwayNumber -> ConwayNumber
x :-: y = x :+: (neg y)

(:*:) :: ConwayNumber -> ConwayNumber -> ConwayNumber
x@(CN xls xrs) :*: y@(CN yls yrs) = 
  CN 
    ([ (xl :*: y) :+: (x :*: yl) :-: (xl :*: yl)
     | xl <- xls, yl <- yls] `union`
     [ (xr :*: y) :+: (x :*: yr) :-: (xr :*: yr)
     | xr <- xrs, yr <- yrs])
    ([ (xl :*: y) :+: (x :*: yr) :-: (xl :*: yr)
     | xl <- xls, yr <- yrs] `union`
     [ (xr :*: y) :+: (x :*: yl) :-: (xr :*: yl)
     | xr <- xrs, yl <- yls])
\end{code}

Conway numbers form a \emph{ring} under the semantic notion of equality
\lstinline$:=:$ defined below:
\begin{code}
geq/+\zerog+/, leq/+\zerog+/, eq/+\zerog+/ :: ConwayNumber -> Bool
geq/+\zerog+/ (CN _ xrs)  = not $ or (map leq/+\zerog+/ xrs) 
leq/+\zerog+/ (CN xls _)  = not $ or (map geq/+\zerog+/ xls) 
eq/+\zerog+/ x            = geq/+\zerog+/ x && leq/+\zerog+/ x

(:=:) :: ConwayNumber -> ConwayNumber -> ConwayNumber
x :=: y = eq/+\zerog+/ (x :-: y)
\end{code}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{abbrvnat}
\softraggedright
\bibliography{cites}

\end{document}