### 2014-05-06 negatives.tex
addition and comparison

\documentclass[authoryear,preprint]{sigplanconf}
\usepackage{amsmath}
\usepackage{listings} 
\usepackage{stmaryrd}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{xcolor}
\usepackage{courier}
\usepackage{thmtools}
\usepackage{bbold}

%%

\newcommand{\geqz}{\geq\zerog}
\newcommand{\leqz}{\leq\zerog}
\newcommand{\zerog}{\mathbb{0}}
\newcommand{\oneg}{\mathbb{1}}
\newcommand{\twog}{\mathbb{2}}
\newcommand{\threeg}{\mathbb{3}}
\newcommand{\fourg}{\mathbb{4}}
\newcommand{\fiveg}{\mathbb{5}}

%%

\begin{document}
\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\newcommand{\alt}{~|~}
\lstnewenvironment{code}{\lstset{basicstyle={\sffamily\footnotesize}}}{}

\lstset{frame=none,
         language=Haskell,
         basicstyle=\sffamily, 
         numberstyle=\tiny,
         numbersep=5pt,
         tabsize=2,    
         extendedchars=true,
         breaklines=true,   
         breakautoindent=true,
         keywordstyle=\color{black},
         captionpos=b,
         stringstyle=\color{black}\ttfamily,
         showspaces=false,  
         showtabs=false,    
         framexleftmargin=2em,
         framexbottommargin=1ex,
         showstringspaces=false
         basicstyle=\sffamily,
         columns=[l]flexible,
         flexiblecolumns=true,
         aboveskip=\smallskipamount,
         belowskip=\smallskipamount,
         lineskip=-1pt,
         xleftmargin=1em,
         escapeinside={/+}{+/},
         keywords=[1]{Monad,Just,Nothing,type,data,right,left,id,where,do,
                     if,then,else,let,in},
         literate=
           {+}{{$\;+\;$}}1 
           {/}{{$/$}}1 
           {*}{{$\;*\;$}}1
           {=}{{$=\ $}}1 
           {/=}{{$\not=$}}1
           {[]}{$[\;]$}2
           {<}{{$<$}}1 
           {>}{{$>$}}1 
           {++}{{$+\!\!\!+\;$}}1 
           {::}{{$:\mkern -2.5mu:\;$}}1
           {&&}{{$\&\!\!\!\&$}}2
           {:=:}{{$:\mkern -2mu=\mkern -2mu:\;$}}3
           {:+:}{{$:\mkern -5mu+\mkern -5mu:\;$}}3
           {:-:}{{$:\mkern -5mu-\mkern -5mu:\;$}}3
           {:*:}{{$:\mkern -5mu*\mkern -5mu:\;$}}3
           {$}{{\texttt{\$}\hspace{0.5em}}}1
           {`}{$^\backprime$}1
           {==}{{$=\!=\;$}}2
           {===}{{$\equiv\;$}}2
           {->}{{$\rightarrow\;$}}2 
           {>=}{{$\geq$}}2 
           {<=}{{$\leq$}}2 
           {>=0}{{$\geq_\zerog\;$}}2 
           {<=0}{{$\leq_\zerog\;$}}2 
           {==0}{{$=_\zerog\;$}}2 
           {>0}{{$>_\zerog\;$}}2 
           {<0}{{$<_\zerog\;$}}2 
           {<-}{{$\leftarrow$}}2
           {=>}{{$\Rightarrow\;$}}2
           {<<}{{$\ll$}}2 
           {>>}{{$\gg\;$}}2
           {>>>}{{$\ggg\;$}}3 
           {<<<}{{$\lll\;$}}3
           {>>=}{{$\gg\mkern -2.5mu=\;$}}3
           {=<<}{{$=\mkern -2.5mu\ll\;$}}3
           {<|}{$\lhd\;$}2
           {<||}{$\unlhd\;$}2
           {\ ||\ }{$\|$}1
           {\\}{$\lambda$}1
           {:>}{{$\rhd$}}2
           {||>}{{$\unrhd$}}2
           {_}{{$\_$}}1
           {_B}{{$_b$}}2
           {forall}{{$\forall$}}1
}

\lstset{postbreak=\raisebox{0ex}[0ex][0ex]
        {\ensuremath{\hookrightarrow}}}
\lstset{breaklines=true, breakatwhitespace=true}
\lstset{numbers=none, numbersep=5pt, stepnumber=2, numberstyle=\scriptsize}
\lstset{rangeprefix=/*!\ , rangesuffix=\ !*\/, includerangemarker=false}

%% double-blind reviewing...
\title{Negative Types}
\authorinfo{}{}{}
\maketitle

\begin{abstract}
\ldots
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conway Numbers}

A Conway number is a game consisting of left and right options where each
option models a move to another game. There is another technical condition
given below that constrains every left option to be ``less than'' every right
option but that can be ignored for the moment. Players alternate taking
options and the player with no available options loses. In Haskell, ignoring
the additional technical condition, one might define the datatype of Conway
numbers as:
\begin{code}
data ConwayNumber = CN [ConwayNumber] [ConwayNumber]
\end{code}
The simplest Conway number is \lstinline|CN [] []| with empty left and right
options. We call this number $\zerog$: 
\begin{code}
/+ $\zerog$ +/  = CN  [] []
\end{code}
Once we have defined $\zerog$, we can also define the following:
\begin{code}
/+ $\oneg$ +/  = CN  [ /+ $\zerog$ +/ ]  []
/+ $\twog$ +/  = CN  [ /+ $\oneg$ +/ ]  []
/+ $\threeg$ +/  = CN  [ /+ $\twog$ +/ ]  []
\end{code}
and so on. Intuitively, the number $\mathbb{n}$ represents a game in which
the left player has an $n$-move advantage over the right player. Dually, we
can define the following numbers where the left player has $n$-move
\emph{disadvantage} over the right player:
\begin{code}
/+ -$\oneg$ +/  = CN  [] [ /+ $\zerog$ +/ ]
/+ -$\twog$ +/  = CN  [] [ /+ -$\oneg$ +/ ]
/+ -$\threeg$ +/  = CN  [] [ /+ -$\twog$ +/ ]
\end{code}
More generally, the \emph{unary negation} of a number or its \emph{dual} is
defined as follows:
\begin{code}
neg :: ConwayNumber -> ConwayNumber
neg (CN xls xrs) = 
  CN [ neg xr | xr <- xrs ] [ neg xl | xl <- xls ]
\end{code}

\textbf{Comparing numbers.} There are three outcomes to games represented by
Conway numbers $\mathbb{n}$:
\begin{itemize}
\item $\mathbb{n} > \zerog$ which indicates that the \emph{left} player can
  enforce a win, no matter who starts;
\item $\mathbb{n} < \zerog$ which indicates that the \emph{right} player can
  enforce a win, no matter who starts;
\item $\mathbb{n} = \zerog$ which indicates that the \emph{second}
  player can enforce a win, no matter who.
\end{itemize}
For example, we have $\threeg > \zerog$ because the right player would have
no options (i.e., would lose) on the first or second turn depending on who
starts. Similarly, we have $-\threeg < \zerog$ because the left player would
have no options on on the first or second turn depending on who starts. For
$\zerog$, it is evident that the first player to take a turn must lose, i.e.,
that the second player wins. More interestingly, consider the game
$\textsf{CN}~[-\oneg]~[\oneg]$ which we claim is $= \zerog$. Indeed, if the
left player starts, the only option is to move to the game $-\oneg$ which the
right player wins. If however the right player moves first, the only option
is to move to the game $\oneg$ which the left player wins. In both cases, the
second player wins. To ensure that these three outcomes are the only possible
outcomes, Conway numbers have an additional restriction: for every left
option $x$ and right option $y$, we require that $x < y$. If that restriction
is lifted we could construct games such that $\textsf{CN}~[\zerog]~[\zerog]$
for which none of the relations $> \zerog$, $< \zerog$, or $= \zerog$
holds. Instead for this game, the first player always wins, which means that
the left player cannot enforce a win (in case the right player starts) and
vice-versa.

The formal definitions of the comparison operators are:
\begin{code}
>=0 (CN _ xrs)  = not $ or (map <=0 xrs)
<=0 (CN xls _)  = not $ or (map >=0 xls) 
==0 x           = >=0 x && <=0 x
<0 x            = <=0 x && not (>=0 x)
>0 x            = >=0 x && not (<=0 x)
\end{code}
Only $\geq_\zerog$ and $\leq_\zerog$ are fundamental. The predicate
$\geq_\zerog$ means that the left player can win as second player and
$\leq_\zerog$ means that the right player can win as second player. If the
left player can win as second player, it means that the right player has no
good opening move, i.e., that there is no right option that is $\leq_\zerog$.

Conway numbers also come equipped with addition, subtraction, and
multiplication operations defined as follows.

\noindent\textbf{Addition and subtraction.} Addition of two games intuitively
gives each player the choice of selecting an option from either
game. Subtraction is simply the addition of the dual of a game. Formally:
\begin{code}
(:+:) :: ConwayNumber -> ConwayNumber -> ConwayNumber
x@(CN xls xrs) :+: y@(CN yls yrs) = 
  CN 
    ([ xl :+: y | xl <- xls ] `union`
     [ x :+: yl | yl <- yls ])
    ([ xr :+: y | xr <- xrs ] `union`
     [ x :+: yr | yr <- yrs ])

(:-:) :: ConwayNumber -> ConwayNumber -> ConwayNumber
x :-: y = x :+: (neg y) 
\end{code}
It is easy to check that $\zerog$ is the unit of addition and that addition
is commutative. Furthermore, adding two positive numbers like~$\twog$
and~$\threeg$ gives $\fiveg$ as desired, and similarly for two negative
numbers. When mixing positive and negative numbers, e.g., adding $\threeg$
and $-\twog$, the result appears much more complicated:
\begin{code}
CN [CN  [CN  [-/+ $\twog$ +/] 
             [CN [-/+ $\oneg$ +/] [/+ $\oneg$ +/]]]
        [CN  [CN [-/+ $\oneg$ +/] [/+ $\oneg$ +/]] 
             [/+ $\twog$ +/]]] 
   [CN  [CN  [CN [-/+ $\oneg$ +/] [/+ $\oneg$ +/]] 
             [/+ $\twog$ +/]] 
        [/+ $\threeg$ +/]]
\end{code}
We will formalize the sense in which the game above is equivalent to the game
$\oneg$, but for now we can intuitively reason as follows. We have already
seen that the game $\textsf{CN}~[-\oneg]~[\oneg]$ is equivalent to $\zerog$
which simplifies the above to:
\begin{code}
CN [CN  [CN  [-/+ $\twog$ +/] [/+ $\zerog$ +/]]
        [CN  [/+ $\zerog$ +/] [/+ $\twog$ +/]]]
   [CN  [CN  [/+ $\zerog$ +/] [/+ $\twog$ +/]]
        [/+ $\threeg$ +/]]
\end{code}
Now consider the game $\textsf{CN}~[\zerog]~[\twog]$. If the left player
starts, the game proceeds to $\zerog$ and the right player loses. If the
right player starts, the game proceeds to $\twog$ which is an even stronger
position for the left player. In other words, the right option is useless and
the right player always loses: the game is equivalent to
$\textsf{CN}~[\zerog]~[\;]$, i.e., to~$\oneg$. By a similar reasoning,
$\textsf{CN}~[-\twog]~[\zerog]$ simplifies to $-\oneg$, and the whole
expression becomes:
\begin{code}
CN [CN  [/+ -$\oneg$ +/] [/+ $\oneg$ +/]]
   [CN  [/+ $\oneg$ +/] [/+ $\threeg$ +/]]
\end{code}
The left option is equivalent to $\zerog$ and the right one is equivalent to
$\twog$, and the entire expression is equivalent to $\oneg$.

\noindent\textbf{Multiplication.} The most complicated operation on Conway
numbers is multiplication. It is defined as follows:
\begin{code}
(:*:) :: ConwayNumber -> ConwayNumber -> ConwayNumber
x@(CN xls xrs) :*: y@(CN yls yrs) = 
  CN 
    ([ (xl :*: y) :+: (x :*: yl) :-: (xl :*: yl)
     | xl <- xls, yl <- yls] `union`
     [ (xr :*: y) :+: (x :*: yr) :-: (xr :*: yr)
     | xr <- xrs, yr <- yrs])
    ([ (xl :*: y) :+: (x :*: yr) :-: (xl :*: yr)
     | xl <- xls, yr <- yrs] `union`
     [ (xr :*: y) :+: (x :*: yl) :-: (xr :*: yl)
     | xr <- xrs, yl <- yls])
\end{code}

Conway numbers form a \emph{ring} under the semantic notion of equality
\lstinline$:=:$ defined below:
\begin{code}
(:=:) :: ConwayNumber -> ConwayNumber -> ConwayNumber
x :=: y = ==0 (x :-: y)
\end{code}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{abbrvnat}
\softraggedright
\bibliography{cites}

\end{document}