### 2014-05-20 ring.tex
something about values

\documentclass[authoryear,preprint]{sigplanconf}
\usepackage{amsmath}
\usepackage{listings} 
\usepackage{stmaryrd}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{xcolor}
\usepackage{courier}
\usepackage{thmtools}
\usepackage{bbold}
\usepackage{tikz}
\usepackage{proof}

\newcommand{\negp}[1]{\textit{neg}(#1)}
\newcommand{\inl}[1]{\textsf{inl}~#1}
\newcommand{\inr}[1]{\textsf{inr}~#1}
\newcommand{\lolli}{\multimap} 
\newcommand{\cubt}{\mathbb{T}}
\newcommand{\den}[1]{\llbracket #1 \rrbracket}
\newcommand{\denc}[1]{\llbracket #1 \rrbracket^c}
\newcommand{\nodet}[2]{\fcolorbox{black}{white}{$#1$}\fcolorbox{black}{gray!20}{$#2$}}

\begin{document}
\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\newcommand{\alt}{~|~}
\lstnewenvironment{code}{\lstset{basicstyle={\sffamily\footnotesize}}}{}

\lstset{frame=none,
         language=Haskell,
         basicstyle=\sffamily, 
         numberstyle=\tiny,
         numbersep=5pt,
         tabsize=2,    
         extendedchars=true,
         breaklines=true,   
         breakautoindent=true,
         keywordstyle=\color{black},
         captionpos=b,
         stringstyle=\color{black}\ttfamily,
         showspaces=false,  
         showtabs=false,    
         framexleftmargin=2em,
         framexbottommargin=1ex,
         showstringspaces=false
         basicstyle=\sffamily,
         columns=[l]flexible,
         flexiblecolumns=true,
         aboveskip=\smallskipamount,
         belowskip=\smallskipamount,
         lineskip=-1pt,
         xleftmargin=1em,
         escapeinside={/+}{+/},
         keywords=[1]{Monad,Just,Nothing,type,data,right,left,id,where,do,
                     if,then,else,let,in},
         literate=
           {+}{{$\;+\;$}}1 
           {/}{{$/$}}1 
           {*}{{$\;*\;$}}1
           {=}{{$=\ $}}1 
           {/=}{{$\not=$}}1
           {[]}{$[\;]$}2
           {<}{{$<$}}1 
           {>}{{$>$}}1 
           {++}{{$+\!\!\!+\;$}}1 
           {::}{{$:\mkern -2.5mu:\;$}}1
           {&&}{{$\&\!\!\!\&$}}2
           {:=:}{{$:\mkern -2mu=\mkern -2mu:\;$}}3
           {:+:}{{$:\mkern -5mu+\mkern -5mu:\;$}}3
           {:-:}{{$:\mkern -5mu-\mkern -5mu:\;$}}3
           {:*:}{{$:\mkern -5mu*\mkern -5mu:\;$}}3
           {$}{{\texttt{\$}\hspace{0.5em}}}1
           {`}{$^\backprime$}1
           {==}{{$=\!=\;$}}2
           {===}{{$\equiv\;$}}2
           {->}{{$\rightarrow\;$}}2 
           {>=}{{$\geq$}}2 
           {<=}{{$\leq$}}2 
           {>=0}{{$\geq_\zerog\;$}}2 
           {<=0}{{$\leq_\zerog\;$}}2 
           {==0}{{$=_\zerog\;$}}2 
           {>0}{{$>_\zerog\;$}}2 
           {<0}{{$<_\zerog\;$}}2 
           {<-}{{$\leftarrow$}}2
           {=>}{{$\Rightarrow\;$}}2
           {<<}{{$\ll$}}2 
           {>>}{{$\gg\;$}}2
           {>>>}{{$\ggg\;$}}3 
           {<<<}{{$\lll\;$}}3
           {>>=}{{$\gg\mkern -2.5mu=\;$}}3
           {=<<}{{$=\mkern -2.5mu\ll\;$}}3
           {<|}{$\lhd\;$}2
           {<||}{$\unlhd\;$}2
           {\ ||\ }{$\|$}1
           {\\}{$\lambda$}1
           {:>}{{$\rhd$}}2
           {||>}{{$\unrhd$}}2
           {_}{{$\_$}}1
           {_B}{{$_b$}}2
           {forall}{{$\forall$}}1
}

\lstset{postbreak=\raisebox{0ex}[0ex][0ex]
        {\ensuremath{\hookrightarrow}}}
\lstset{breaklines=true, breakatwhitespace=true}
\lstset{numbers=none, numbersep=5pt, stepnumber=2, numberstyle=\scriptsize}
\lstset{rangeprefix=/*!\ , rangesuffix=\ !*\/, includerangemarker=false}

%% double-blind reviewing...
\title{Cubical Polarized Types}
\authorinfo{}{}{}
\maketitle

\begin{abstract}
\ldots
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

The \textbf{Int} construction or the $\mathcal{G}$ construction are neat. As
Neel K. explains, given first-order types and feedback you get higher-order
functions. But if you do the construction on the additive structure, you lose
the multiplicative structure. It turns out that this is related to a deep
open problem in algebraic topology and homotopy theory that was recently
solved. We ``translate'' that solution to a computational type-theoretic
world. This has evident connections to homotopy (type?) theory that remain to
be investigated.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The \textbf{Int} Construction}

Explain in detail perhaps with Haskell embedding and type functions etc. The
key insight it to add ``negative'' types representing demand for values. This
is how you get functions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Problem and the Intuitive Solution}

We lose the multiplicative structure. No evident way to define the
multiplication functor. Perhaps there is a clever way. But this turns out to
be a well-known open problem. Review the problem and tell the story from the
algebraic topology perspective.

Regular first-order types are viewed as 0-dimensional cubes. Generalize to
$n$-dimensional cubes.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Type Structure}

We first define the syntax and semantics of types and then discuss type
isomorphisms.

%%%%%%%%%%%%%%%%%%%%
\subsection{Negative and Cubical Types}

We use $t$ for regular first-order types which include the empty type, the
unit type, and conventional sum and product types.  We use $\tau$ for cubical
types: these have the same formal structure as regular types but also include
\emph{negative} types. The syntax of the types is summarized below:
\[\begin{array}{rcl}
t &::=& 0 \alt 1 \alt t_1 + t_2 \alt t_1 * t_2 \\
\tau &::=& t \alt 
      \alt \tau_1 + \tau_2
      \alt \tau_1 * \tau_2
      \alt - \tau
\end{array}\]
We use $\tau_1 - \tau_2$ to abbreviate $\tau_1 + (- \tau_2)$ and more
interestingly $\tau_1 \lolli \tau_2$ to abbreviate $(- \tau_1) + \tau_2$. At
the level of cubical types, we can write types such as
$(t_1-t_2)-(t_3-t_4*t_5)$ which mix negative types with the usual type
constructors.

It is easy to find a model for regular types: finite sets are such a
model. In more detail:
\[\begin{array}{rcl}
\den{0} &=& \emptyset \\
\den{1} &=& \{ \bullet \} \\
\den{t_1 + t_2} &=& \den{t_1} \uplus \den{t_2} \\
\den{t_1 * t_2} &=& \den{t_1} \times \den{t_2} 
\end{array}\]
The type 0 maps to the empty set. The type 1 maps to a singleton set. The sum
of types maps to disjoint union of sets. The product of types maps to the
cartesian product of sets. We usually use $S$ for the denotation of a
conventional first-order type.

For cubical types, their denotations, usually written as $\cubt$, are
$n$-dimensional cubical sets with polarities. An $n$-dimensional cubical set
$\nodet{\cubt_1}{\cubt_2}$ consists of a positive subspace $\cubt_1$ and a
negative subspace $\cubt_2$ (shaded in gray). Each of the subspaces are
cubical set of a lower dimension; the 0-dimensional case reduces to the
denotation of conventional first-order types. As we explain below, cubical
sets are closed under sums, differences, products.

Examples of 0-dimensional cubical sets include all the finite sets denoting a
regular first-order type. A 1-dimensional cubical set looks like
$\nodet{S_1}{S_2}$ which intuitively corresponds to the difference $t_1 -
t_2$ of the two types whose denotations are $S_1$ and $S_2$ respectively. A
2-dimensional cubical type looks like
$\nodet{(\nodet{S_1}{S_2})}{(\nodet{S_3}{S_4})}$ which intuitively
corresponds to the iterated difference of the appropriate types
$(t_1-t_2)-(t_3-t_4)$ where the product of the successive signs from the
outermost box encodes the sign.

We now give the denotation of cubical types:
\[\begin{array}{rcl}
\denc{t} &=& \den{t} \\
\denc{\tau_1 + \tau_2} &=& \denc{\tau_1} \oplus \denc{\tau_2} \\
\denc{\tau_1 * \tau_2} &=& \denc{\tau_1} \otimes \denc{\tau_2} \\
\denc{- \tau} &=& \ominus \denc{\tau}
\end{array}\]
where the operations $\oplus$, $\otimes$, and $\ominus$ are defined below.
It is possible for cubical sets not be balanced in general but it is always
possible to embed an $n$-dimensional set into a larger dimension using the
fact (to be justified later) that $t$ is equivalent to $t-0$ and to $(t-0)-0$
and so on, i.e.,
\[
S = \nodet{S}{\phantom{S}} = 
  \nodet{\nodet{S}{\phantom{S}}}
        {\phantom{{\nodet{S}{\phantom{S}}}}} = \ldots \\
\]
In the recursive definitions below, we assume that the cubical sets have been
extended to be balanced as needed:
\[\begin{array}{rcl}
\ominus~S &=& \nodet{\phantom{S}}{S} \\
\ominus~\nodet{\cubt_1}{\cubt_2} &=& \nodet{\ominus~\cubt_2}{\ominus~\cubt_1} \\
\\
S_1 \oplus S_2 &=& S_1 \uplus S_2 \\
(\nodet{\cubt_1}{\cubt_2}) \oplus (\nodet{\cubt_3}{\cubt_4}) &=& 
  \nodet{\cubt_1 \oplus \cubt_3}{\cubt_2 \oplus \cubt_4} \\
\\
S_1 \otimes S_2 &=& S_1 \times S_2 \\
S \otimes (\nodet{\cubt_1}{\cubt_2}) &=& 
  \nodet{S \otimes \cubt_1}{S \otimes \cubt_2} \\
(\nodet{\cubt_1}{\cubt_2}) \otimes \cubt &=& 
  \nodet{\cubt_1 \otimes \cubt}{\cubt_2 \otimes \cubt}
\end{array}\]

\begin{figure*}
\[\begin{array}{c}
\nodet{S_1}{S_2}
\quad\otimes\quad
\nodet{(\nodet{S_3}{S_4})}{(\nodet{S_5}{S_6})} \quad= \\
\\
\nodet{(\nodet{{(\nodet{S_1 \times S_3}{S_1 \times S_4})}}
              {{(\nodet{S_1 \times S_5}{S_1 \times S_6})}})}
      {(\nodet{{(\nodet{S_2 \times S_3}{S_2 \times S_4})}}
              {{(\nodet{S_2 \times S_5}{S_2 \times S_6})}})}
\end{array}\]
\caption{\label{mult}Example of multiplication of two cubical sets.}
\end{figure*}

\noindent The definition of negation is natural: it recursively swaps the
positive and negative subspaces. We note that the general definitions of sum
and product reduce to the conventional sums and products for 0-dimensional
types. Also the sum of 1-dimensional types reduces to the sum used in the
\textbf{Int} construction and that addition and negation do not increase the
dimensions. The interesting new development is the product definition. The
example in Fig.~\ref{mult} illustrates a general pattern: the product of an
$n$-dimensional type with an $m$-dimensional type gives an
$(n+m)$-dimensional type. For low dimensions such as the ones used in the
example above, it is possible to visualize what is happening in coordinate
space (see Fig.~\ref{multv}).

\begin{figure*}
\begin{tikzpicture}
\node[left] at (0,0) {$S_1^+$};
\draw[fill] (0,0) circle [radius=0.05];
\node[right] at (0.6,0) {$S_2^-$};
\draw[fill] (0.6,0) circle [radius=0.05];
\draw[-] (0,0) -- (0.6,0);
\node at (1.6,0) {$\otimes$};

%%
\node[below left] at (2.5,-0.5) {$S_3^+$};
\draw[fill] (2.5,-0.5) circle [radius=0.05];
\node[below right] at (3.5,-0.5) {$S_4^-$};
\draw[fill] (3.5,-0.5) circle [radius=0.05];
\draw[-] (2.5,-0.5) -- (3.5,-0.5);
\draw[-] (2.5,-0.5) -- (2.5,0.5);
\node[above left] at (2.5,0.5) {$S_5^-$};
\draw[fill] (2.5,0.5) circle [radius=0.05];
\node[above right] at (3.5,0.5) {$S_6^+$};
\draw[fill] (3.5,0.5) circle [radius=0.05];
\draw[-] (2.5,0.5) -- (3.5,0.5);
\draw[-] (3.5,-0.5) -- (3.5,0.5);
%% 
\node at (5,0) {$=$};
%% 
\node[left] at (7.5,0.75) {$(S_2 \times S_3)^-$};
\draw[fill] (7.5,0.75) circle [radius=0.05];
\node[right] at (9.5,0.75) {$(S_2 \times S_4)^+$};
\draw[fill] (9.5,0.75) circle [radius=0.05];
\node[above right] at (10.2,1.2) {$(S_2 \times S_6)^-$};
\draw[fill] (10.2,1.2) circle [radius=0.05];
\node[above left] at (8.2,1.2) {$(S_2 \times S_5)^+$};
\draw[fill] (8.2,1.2) circle [radius=0.05];
%%
\node[left] at (7.5,-0.75) {$(S_1 \times S_3)^+$};
\draw[fill] (7.5,-0.75) circle [radius=0.05];
\node[right] at (9.5,-0.75) {$(S_1 \times S_4)^-$};
\draw[fill] (9.5,-0.75) circle [radius=0.05];
\node[above right] at (10.2,-0.3) {$(S_1 \times S_6)^+$};
\draw[fill] (10.2,-0.3) circle [radius=0.05];
\node[above] at (8.2,-0.3) {$(S_1 \times S_5)^-$};
\draw[fill] (8.2,-0.3) circle [radius=0.05];
%%
\draw[-] (7.5,0.75) -- (9.5,0.75);
\draw[-] (9.5,0.75) -- (10.2,1.2);
\draw[-] (10.2,1.2) -- (8.2,1.2);
\draw[-] (8.2,1.2) -- (7.5,0.75);
%%
\draw[-] (7.5,-0.75) -- (9.5,-0.75);
\draw[-] (9.5,-0.75) -- (10.2,-0.3);
\draw[-,dashed] (10.2,-0.3) -- (8.2,-0.3);
\draw[-,dashed] (8.2,-0.3) -- (7.5,-0.75);
%%
\draw[-] (7.5,0.75) -- (7.5,-0.75);
\draw[-] (9.5,0.75) -- (9.5,-0.75);
\draw[-] (10.2,1.2) -- (10.2,-0.3);
\draw[-,dashed] (8.2,1.2) -- (8.2,-0.3);
\end{tikzpicture}
\caption{\label{multv}Visualization of multiplication of two cubical sets.}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%
\subsection{Type Isomorphisms}

\begin{verbatim}
To complete the story we need to 
define morphisms. (More on this below.)
Once we have a notion of morphism we 
can check whether X + 0 is the same
as X etc. i.e., we can check all the 
ring equivalences.

Ok so what are the morphisms between 
these cubical objects? We know what
they are for 1-dimensional cubes: 
they are the pi combinabors. We also
know what they are for the 2-dimensional 
cubes: a maps (A-B) ==> (C-D) 
is a Pi map between A+D <=> C+B. 
How to generalize this?

Why is there no trace in the ring 
completion paper??? What are 
the morphisms in that paper?

The ring completion paper produces
a simplicial category.

p. 3 talks about the group cancellation
as subcubes along the diagonal...

We shouldn't focus on denotations. We want
an operational semantics for pi with 
negatives. The same way that Neel turned
the G construction into code that does
something neat; we want to turn that 
ring completion construction into code
that produces h.o. functions without
losing the multiplicative structure.

Show how to embed a square in 2D into
each of the faces of the 3D cube.

1D into 2D:

(a-b) => (a-b)-(0-0)
(a-b) => (a-0)-(b-0)
(a-b) => (0-b)-(0-a)
(a-b) => (0-0)-(b-a)

\end{verbatim}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{A Reversible Language with Cubical Types}

Now that the type structure is defined, we turn our attention to the notion
of values. For conventional first-order types, the values are also
conventional:
\[
\infer{() : 1}{} 
\qquad
\infer{\inl{v} : t_1 + t_2}{v : t_1}
\qquad
\infer{\inr{v} : t_1 + t_2}{v : t_2}
\qquad
\infer{(v_1,v_2) : t_1 * t_2}{v_1 : t_1 & v_2 : t_2}
\]
For cubical types, the values are more involved. Intuitively, a value of the
$n$-dimensional type $\tau$ is an element of one of the sets located in a
corner of the $n$-dimensional cube denoted by $\tau$. Thus to specify a
value, we must specify one of the corners of the cube which can easily be
done using a sequence $p$ of $+$ and $-$ polarities indicating how to
navigate the cube in each successive dimension starting from a fixed origin
to reach the desired corner. We denote such values as $v^{p}$ and use
$\epsilon$ for the empty sequence of polarities. One can think of the
sequence of signs as denoting a path in the binary tree representation of
cubical types. Formally:
\[\begin{array}{c}
\infer[\textit{base}]{v^{\epsilon} : t}{v : t} \\
\\
\infer[\textit{left}]{(\inl{v^p})^{p} : \tau_1 + \tau_2}{v^{p} : \tau_1}
\qquad
\infer[\textit{right}]{(\inr{v^p})^{p} : \tau_1 + \tau_2}{v^{p} : \tau_2} \\
\\
\infer[\textit{prod}]{(v_1^{p_1},v_2^{p_2})^{p_1 \bullet p_2} : \tau_1 * \tau_2}
      {v_1^{p_1} : \tau_1 & v_2^{p_2} : \tau_2} \\
\\
\infer[\textit{neg}]{v^{\negp{p}} : - \tau}{v^p : \tau} 
\end{array}\]
The rule \textit{base} states that a conventional first-order value can be
viewed as trivial 0-dimensional cubical value. The rules \textit{left} and
\textit{right} reflect the fact that sums do not increase the dimension and
works pointwise. The rule \textit{prod} is the most involved one: it
increases the dimension as calculated below:
\[\begin{array}{rcl}
\epsilon \bullet p &=& p \\
(+p_1) \bullet p_2 &=& + (p_1 \bullet p_2) \\
(-p_1) \bullet p_2 &=& - (p_1 \bullet p_2) 
\end{array}\]
The rule \textit{neg} shows that a negative value $v$ is located at the
``opposite'' corner of $v$ which is calculated as follows:
\[\begin{array}{rcl}
\negp{\epsilon} &=& -\epsilon \\
\negp{+p} &=& -\negp{p} \\
\negp{-p} &=& +\negp{p}
\end{array}\]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work and Context}

A ton of stuff here.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{abbrvnat}
\softraggedright
\bibliography{cites}

\end{document}