### 2014-05-23 ring.tex
more about paths and groupoids

\documentclass[authoryear,preprint]{sigplanconf}
\usepackage{amsmath}
\usepackage{listings} 
\usepackage{stmaryrd}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{xcolor}
\usepackage{courier}
\usepackage{thmtools}
\usepackage{bbold}
\usepackage{tikz}
\usepackage{proof}
\usepackage{graphicx}

\newcommand{\refl}[1]{\textsf{refl}_{#1}}
\newcommand{\seg}[1]{\textsf{seg}_{#1}}
\newcommand{\bdim}[1]{\textsf{dim}(#1)}
\newcommand{\Rule}[4]{
\makebox{{\rm #1}
$\displaystyle
\frac{\begin{array}{l}#2\\\end{array}}
{\begin{array}{l}#3\\\end{array}}$
 #4}}
\newcommand{\proves}{\vdash}
\newcommand{\symc}[1]{\mathit{sym}~#1}
\newcommand{\jdg}[3]{#2 \proves_{#1} #3}
\newcommand{\adjoint}[1]{#1^{\dagger}}
\newcommand{\iso}{\leftrightarrow}
\newcommand{\identlp}{\mathit{identl}_+}
\newcommand{\identrp}{\mathit{identr}_+}
\newcommand{\swapp}{\mathit{swap}_+}
\newcommand{\assoclp}{\mathit{assocl}_+}
\newcommand{\assocrp}{\mathit{assocr}_+}
\newcommand{\identlt}{\mathit{identl}_*}
\newcommand{\identrt}{\mathit{identr}_*}
\newcommand{\swapt}{\mathit{swap}_*}
\newcommand{\assoclt}{\mathit{assocl}_*}
\newcommand{\assocrt}{\mathit{assocr}_*}
\newcommand{\distz}{\mathit{dist}_0}
\newcommand{\factorz}{\mathit{factor}_0}
\newcommand{\dist}{\mathit{dist}}
\newcommand{\factor}{\mathit{factor}}
\newcommand{\idc}{\mathit{id}}
\newcommand{\plus}{\raisebox{.4\height}{\scalebox{.6}{+}}}
\newcommand{\minus}{\raisebox{.4\height}{\scalebox{.8}{-}}}
\newcommand{\mm}{\texttt{-}}
\newcommand{\pp}{\texttt{+}}
\newcommand{\negp}[1]{\textit{neg}(#1)}
\newcommand{\inl}[1]{\textsf{inl}~#1}
\newcommand{\inr}[1]{\textsf{inr}~#1}
\newcommand{\lolli}{\multimap} 
\newcommand{\cubt}{\mathbb{T}}
\newcommand{\den}[1]{\llbracket #1 \rrbracket}
\newcommand{\nodet}[2]{\fcolorbox{black}{white}{$#1$}\fcolorbox{black}{gray!20}{$#2$}}
\newcommand{\hast}{:\mkern -2.5mu:\;}

\begin{document}
\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\newcommand{\alt}{~|~}
\lstnewenvironment{code}{\lstset{basicstyle={\sffamily\footnotesize}}}{}

\lstset{frame=none,
         language=Haskell,
         basicstyle=\sffamily, 
         numberstyle=\tiny,
         numbersep=5pt,
         tabsize=2,    
         extendedchars=true,
         breaklines=true,   
         breakautoindent=true,
         keywordstyle=\color{black},
         captionpos=b,
         stringstyle=\color{black}\ttfamily,
         showspaces=false,  
         showtabs=false,    
         framexleftmargin=2em,
         framexbottommargin=1ex,
         showstringspaces=false
         basicstyle=\sffamily,
         columns=[l]flexible,
         flexiblecolumns=true,
         aboveskip=\smallskipamount,
         belowskip=\smallskipamount,
         lineskip=-1pt,
         xleftmargin=1em,
         escapeinside={/+}{+/},
         keywords=[1]{Monad,Just,Nothing,type,data,right,left,id,where,do,
                     if,then,else,let,in},
         literate=
           {+}{{$\;+\;$}}1 
           {/}{{$/$}}1 
           {*}{{$\;*\;$}}1
           {=}{{$=\ $}}1 
           {/=}{{$\not=$}}1
           {[]}{$[\;]$}2
           {<}{{$<$}}1 
           {>}{{$>$}}1 
           {++}{{$+\!\!\!+\;$}}1 
           {::}{{$:\mkern -2.5mu:\;$}}1
           {&&}{{$\&\!\!\!\&$}}2
           {:=:}{{$:\mkern -2mu=\mkern -2mu:\;$}}3
           {:+:}{{$:\mkern -5mu+\mkern -5mu:\;$}}3
           {:-:}{{$:\mkern -5mu-\mkern -5mu:\;$}}3
           {:*:}{{$:\mkern -5mu*\mkern -5mu:\;$}}3
           {$}{{\texttt{\$}\hspace{0.5em}}}1
           {`}{$^\backprime$}1
           {==}{{$=\!=\;$}}2
           {===}{{$\equiv\;$}}2
           {->}{{$\rightarrow\;$}}2 
           {>=}{{$\geq$}}2 
           {<=}{{$\leq$}}2 
           {>=0}{{$\geq_\zerog\;$}}2 
           {<=0}{{$\leq_\zerog\;$}}2 
           {==0}{{$=_\zerog\;$}}2 
           {>0}{{$>_\zerog\;$}}2 
           {<0}{{$<_\zerog\;$}}2 
           {<-}{{$\leftarrow$}}2
           {=>}{{$\Rightarrow\;$}}2
           {<<}{{$\ll$}}2 
           {>>}{{$\gg\;$}}2
           {>>>}{{$\ggg\;$}}3 
           {<<<}{{$\lll\;$}}3
           {>>=}{{$\gg\mkern -2.5mu=\;$}}3
           {=<<}{{$=\mkern -2.5mu\ll\;$}}3
           {<|}{$\lhd\;$}2
           {<||}{$\unlhd\;$}2
           {\ ||\ }{$\|$}1
           {\\}{$\lambda$}1
           {:>}{{$\rhd$}}2
           {||>}{{$\unrhd$}}2
           {_}{{$\_$}}1
           {_B}{{$_b$}}2
           {forall}{{$\forall$}}1
}

\lstset{postbreak=\raisebox{0ex}[0ex][0ex]
        {\ensuremath{\hookrightarrow}}}
\lstset{breaklines=true, breakatwhitespace=true}
\lstset{numbers=none, numbersep=5pt, stepnumber=2, numberstyle=\scriptsize}
\lstset{rangeprefix=/*!\ , rangesuffix=\ !*\/, includerangemarker=false}

%% double-blind reviewing...
\title{Negative Types}
\authorinfo{}{}{}
\maketitle

\begin{abstract}
\ldots
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

The \textbf{Int} construction or the $\mathcal{G}$ construction are neat. As
Neel K. explains, given first-order types and feedback you get higher-order
functions. But if you do the construction on the additive structure, you lose
the multiplicative structure. It turns out that this is related to a deep
open problem in algebraic topology and homotopy theory that was recently
solved. We ``translate'' that solution to a computational type-theoretic
world. This has evident connections to homotopy (type) theory that remain to
be investigated.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The \textbf{Int} Construction}

Explain in detail perhaps with Haskell embedding and type functions etc. The
key insight it to add ``negative'' types representing demand for values. This
is how you get functions.

We lose the multiplicative structure. No evident way to define the
multiplication functor. Perhaps there is a clever way. But this turns out to
be a well-known open problem. Review the problem and tell the story from the
algebraic topology perspective. Regular first-order types are viewed as
$0$-dimensional cubes. The \textbf{Int} construction generalizes types to
1-dimensional cubes. Our work here generalizes types to $n$-dimensional
cubes.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Type Structure}

We first define the syntax and then present a simple semantic model of types
which is then refined.

%%%%%%%%%%%%%%%%%%%%
\subsection{Negative and Cubical Types}

Our types $\tau$ include the empty type 0, the unit type 1, conventional sum
and product types, as well as \emph{negative} types:
\[\begin{array}{rcl}
\tau &::=& 0 \alt 1 \alt \tau_1 + \tau_2 \alt \tau_1 * \tau_2 \alt - \tau
\end{array}\]
We use $\tau_1 - \tau_2$ to abbreviate $\tau_1 + (- \tau_2)$ and more
interestingly $\tau_1 \lolli \tau_2$ to abbreviate $(- \tau_1) + \tau_2$.
The \emph{dimension} of a type is defined as follows:
\[\begin{array}{rcl}
\bdim{\cdot} &\hast& \tau \rightarrow \mathbb{N} \\
\bdim{0} &=& 0 \\
\bdim{1} &=& 0 \\
\bdim{\tau_1 + \tau_2} &=& \max(\bdim{\tau_1},\bdim{\tau_2}) \\
\bdim{\tau_1 * \tau_2} &=& \bdim{\tau_1} + \bdim{\tau_2} \\
\bdim{- \tau} &=& \max(1,\bdim{\tau})
\end{array}\]
The base types have dimension 0. If negative types are not used, all
dimensions remain at 0. If negative types are used but no product types of
negative types appear anywhere, the dimension is raised to 1. This is the
situation with the \textbf{Int} or $\mathcal{G}$ construction. Once negative
and product types are freely used, the dimension can increase without bounds.

\begin{figure*}
\[\begin{array}{c}
\nodet{S_1}{S_2}
\quad\otimes\quad
\nodet{(\nodet{S_3}{S_4})}{(\nodet{S_5}{S_6})} \quad= \\
\\
\nodet{(\nodet{{(\nodet{S_1 \times S_3}{S_1 \times S_4})}}
              {{(\nodet{S_1 \times S_5}{S_1 \times S_6})}})}
      {(\nodet{{(\nodet{S_2 \times S_3}{S_2 \times S_4})}}
              {{(\nodet{S_2 \times S_5}{S_2 \times S_6})}})}
\end{array}\]
\\
\\
\begin{center}
\begin{tikzpicture}
\node[above] at (0,0) {\pp};
\node[left] at (0,0) {$S_1$};
\draw[fill] (0,0) circle [radius=0.05];
\node[above] at (0.6,0) {\mm};
\node[right] at (0.6,0) {$S_2$};
\draw[fill] (0.6,0) circle [radius=0.05];
\draw[-,dotted] (0,0) -- (0.6,0);
\node at (1.6,0) {$\otimes$};

%%
\node[left] at (2.5,-0.5) {\pp\pp};
\node[below] at (2.5,-0.5) {$S_3$};
\draw[fill] (2.5,-0.5) circle [radius=0.05];
\node[right] at (3.5,-0.5) {\pp\mm};
\node[below] at (3.5,-0.5) {$S_4$};
\draw[fill] (3.5,-0.5) circle [radius=0.05];
\draw[-,dotted] (2.5,-0.5) -- (3.5,-0.5);
\draw[-,dotted] (2.5,-0.5) -- (2.5,0.5);
\node[left] at (2.5,0.5) {\mm\pp};
\node[above] at (2.5,0.5) {$S_5$};
\draw[fill] (2.5,0.5) circle [radius=0.05];
\node[right] at (3.5,0.5) {\mm\mm};
\node[above] at (3.5,0.5) {$S_6$};
\draw[fill] (3.5,0.5) circle [radius=0.05];
\draw[-,dotted] (2.5,0.5) -- (3.5,0.5);
\draw[-,dotted] (3.5,-0.5) -- (3.5,0.5);
%% 
\node at (5,0) {$=$};
%% 
\node[left] at (7.5,0.75) {$(S_2 \times S_3)$\,\mm\pp\pp};
\draw[fill] (7.5,0.75) circle [radius=0.05];
\node[right] at (9.5,0.75) {$(S_2 \times S_4)$\,\mm\pp\mm};
\draw[fill] (9.5,0.75) circle [radius=0.05];
\node[above right] at (10.2,1.2) {$(S_2 \times S_6)$\,\mm\mm\mm};
\draw[fill] (10.2,1.2) circle [radius=0.05];
\node[above left] at (8.2,1.2) {$(S_2 \times S_5)$\,\mm\mm\pp};
\draw[fill] (8.2,1.2) circle [radius=0.05];
%%
\node[left] at (7.5,-0.75) {$(S_1 \times S_3)$\,\pp\pp\pp};
\draw[fill] (7.5,-0.75) circle [radius=0.05];
\node[right] at (9.5,-0.75) {$(S_1 \times S_4)$\,\pp\pp\mm};
\draw[fill] (9.5,-0.75) circle [radius=0.05];
\node[above right] at (10.2,-0.3) {$(S_1 \times S_6)$\,\pp\mm\mm};
\draw[fill] (10.2,-0.3) circle [radius=0.05];
\node[left] at (8.2,-0.3) {$(S_1 \times S_5)$\,\pp\mm\pp};
\draw[fill] (8.2,-0.3) circle [radius=0.05];
%%
\draw[-,dotted] (7.5,0.75) -- (9.5,0.75);
\draw[-,dotted] (9.5,0.75) -- (10.2,1.2);
\draw[-,dotted] (10.2,1.2) -- (8.2,1.2);
\draw[-,dotted] (8.2,1.2) -- (7.5,0.75);
%%
\draw[-,dotted] (7.5,-0.75) -- (9.5,-0.75);
\draw[-,dotted] (9.5,-0.75) -- (10.2,-0.3);
\draw[-,dotted,dashed] (10.2,-0.3) -- (8.2,-0.3);
\draw[-,dotted,dashed] (8.2,-0.3) -- (7.5,-0.75);
%%
\draw[-,dotted] (7.5,0.75) -- (7.5,-0.75);
\draw[-,dotted] (9.5,0.75) -- (9.5,-0.75);
\draw[-,dotted] (10.2,1.2) -- (10.2,-0.3);
\draw[-,dotted,dashed] (8.2,1.2) -- (8.2,-0.3);
\end{tikzpicture}
\end{center}
\caption{\label{mult}Example of multiplication of two cubical types.}
\end{figure*}

This point is made precise in the following tentative denotation of types (to
be refined in the next section) which maps a type of dimension $n$ to an
$n$-dimensional cube. We represent such a cube syntactically as a binary tree
of maximum depth~$n$ with nodes of the form $\nodet{\cubt_1}{\cubt_2}$. In
such a node, $\cubt_1$ is the positive subspace and $\cubt_2$ (shaded in
gray) is the negative subspace along the first dimension. Each of these
subspaces is itself a cube of a lower dimension. The $0$-dimensional cubes
are plain sets representing the denotation of conventional first-order
types. We use $S$ to denote the denotations of these plain types. A
1-dimensional cube, $\nodet{S_1}{S_2}$, intuitively corresponds to the
difference $t_1 - t_2$ of the two types whose denotations are $S_1$ and $S_2$
respectively. The type can be visualized as a ``line'' with polarized
endpoints connecting the two points~$S_1$ and $S_2$. A full 2-dimensional
cube, $\nodet{(\nodet{S_1}{S_2})}{(\nodet{S_3}{S_4})}$, intuitively
corresponds to the iterated difference of the appropriate types
$(t_1-t_2)-(t_3-t_4)$ where the successive ``colors'' from the outermost box
encode the sign. The type can be visualized as a ``square'' with polarized
corners connecting the two lines corresponding to $(t_1-t_2)$ and
$(t_3-t_4)$. (See Fig.~\ref{mult} which is further explained after we discuss
multiplication below.)

Formally, the denotation of types is as follows:
\[\begin{array}{rcl}
\den{0} &=& \emptyset \\
\den{1} &=& \{ \star \} \\
\den{\tau_1 + \tau_2} &=& \den{\tau_1} \oplus \den{\tau_2} \\
\den{\tau_1 * \tau_2} &=& \den{\tau_1} \otimes \den{\tau_2} \\
\den{- \tau} &=& \ominus \den{\tau} \\
\\
\noalign{\mbox{where:}\hfill}
\\
S_1 \oplus S_2 &=& S_1 \uplus S_2 \\
S \oplus (\nodet{\cubt_1}{\cubt_2}) &=& \nodet{S \oplus \cubt_1}{\cubt_2} \\
(\nodet{\cubt_1}{\cubt_2}) \oplus S &=& \nodet{\cubt_1 \oplus S}{\cubt_2} \\
(\nodet{\cubt_1}{\cubt_2}) \oplus (\nodet{\cubt_3}{\cubt_4}) &=& 
  \nodet{\cubt_1 \oplus \cubt_3}{\cubt_2 \oplus \cubt_4} \\
\\
S_1 \otimes S_2 &=& S_1 \times S_2 \\
S \otimes (\nodet{\cubt_1}{\cubt_2}) &=& 
  \nodet{S \otimes \cubt_1}{S \otimes \cubt_2} \\
(\nodet{\cubt_1}{\cubt_2}) \otimes \cubt &=& 
  \nodet{\cubt_1 \otimes \cubt}{\cubt_2 \otimes \cubt} \\
\\
\ominus~S &=& \nodet{\phantom{S}}{S} \\
\ominus~\nodet{\cubt_1}{\cubt_2} &=& \nodet{\ominus~\cubt_2}{\ominus~\cubt_1} 
\end{array}\]
The type 0 maps to the empty set. The type 1 maps to a singleton set. The sum
of $0$-dimensional types is the disjoint union as usual. For cubes of higher
dimensions, the subspaces are recursively added. Note that the sum of
1-dimensional types reduces to the sum used in the \textbf{Int} construction.
The definition of negation is natural: it recursively swaps the positive and
negative subspaces. The product of 0-dimensional types is the cartesian
product of sets. For cubes of higher-dimensions $n$ and $m$, the result is of
dimension $(n+m)$. The example in Fig.~\ref{mult} illustrates the idea using
the product of 1-dimensional cube (i.e., a line) with a 2-dimensional cube
(i.e., a square). The result is a 3-dimensional cube as illustrated.

%%%%%%%%%%%%%%%%%%%%
\subsection{Type Isomorphisms: Paths to the Rescue}

Our proposed semantics of types identifies several structurally different
types such as $(1+(1+1))$ and $((1+1)+1)$. In some sense, this is innocent as
the types are isomorphic. However, in the operational semantics discussed in
the next section, we make the computational content of such type isomorphisms
explicit. Some other isomorphic types like $(t_1*t_2)$ and $(t_2*t_1)$ map to
different cubes and are \emph{not} identified: explicit isomorphisms are
needed to mediate between them. We therefore need to enrich our model of
types with isomorphisms connecting types we deem equivalent.

So far, our types are modeled as cubes which are really sets indexed by
polarities. An isomorphism between $(t_1*t_2)$ and $(t_2*t_1)$ requires
nothing more than a pair of set-theoretic functions between the spaces that
compose to the identity. What is much more interesting are the isomorphisms
involving the empty type~0. In particular, if negative types are to be
interpreted as their name suggests, we must have an isomorphism between
$(t-t)$ and the empty type 0. Semantically the former denotes the ``line''
$\nodet{\cubt}{\cubt}$ and the latter denotes the empty set. Their
denotations are different and there is no way, in the world of plain sets, to
express the fact that these two spaces should be identified. What is needed
is the ability to \emph{contract} the \emph{path} between the endpoints of
the line to the trivial path on the empty type. This is, of course, where the
ideas of homotopy (type) theory enter the development.

Consider the situation above in which we want to identify the spaces
corresponding to the types $(1-1)$ and the empty type:
\begin{center}
\begin{tikzpicture}
\node[above] at (0,0) {\pp};
\node[left] at (-0.2,0) {$1$};
\draw[fill] (0,0) circle [radius=0.05];
\node[above] at (2.6,0) {\mm};
\node[right] at (2.8,0) {$1$};
\draw[fill] (2.6,0) circle [radius=0.05];
\draw[->,red] (0.1,0) -- (2.5,0);
\draw[->,red] (-0.1,-0.1) arc (135:405:0.2);
\node[below,red] at (0,-0.4) {$\refl{1}$};
\draw[->,red] (2.5,-0.1) arc (135:405:0.2);
\node[below,red] at (2.6,-0.4) {$\refl{1}$};
\node[above] at (1.3,0) {$\seg{1}$};
\node[right] at (1.3,-0.5) {$q$};
%%
\node[below] at (1.3,-1.5) {$\emptyset$};
\draw[fill] (1.3,-1.5) circle [radius=0.05];
\draw[->,red] (1.4,-1.4) arc (-45:225:0.2);
\node[right,red] at (1.5,-1.3) {$\refl{0}$};
%% 
\draw[->,double,blue] (1.3,-0.1) -- (1.3,-1);
\end{tikzpicture}
\end{center}
The top of the figure is the 1-dimensional cube representing the type $(1-1)$
as before except that we now add a path $\seg{1}$ to connect the two
endpoints. (Note that previously, the dotted lines in the figures were a
visualization aid and were \emph{not} meant to represent paths.) We also make
explicit the trivial identity paths from every space to itself.  The bottom
of the figure is the 0-dimensional cube representing the empty type. The path
$\seg{1}$ identifies the two occurrences of 1. To express the
equivalence of $(1-1)$ and 0, we add a 2-path, i.e. a path between paths,
that asserts connects the path $\seg{1}$ to the trivial path
$\refl{0}$ on the empty type. That effectively makes the two points
``disappear.''  Surprisingly, that is everything that we need. The extension
to higher dimensions just ``works'' because paths in homotopy type theory
have a rich structure. We explain the details below.

%%%%%%%%%%%%%%%%%%%%
\subsection{The Universe of Types}

We describe the construction of our universe of types. Our starting point is
the construction in the previous section which we summarize again. We start
with all finite sets built from the empty set, a singleton set, disjoint
unions, and cartesian products. All these sets are thought of being indexed
by the empty sequence of polarities $\epsilon$. We then constructs new spaces
that consist of pairs of finite sets $\nodet{S_1}{S_2}$ indexed by positive
and negative polarities. These are the 1-dimensional spaces. We iterate this
construction to the limit to get $n$-dimensional cubes for all natural
numbers $n$.

At this point, we switch from viewing the universe of types as an
unstructured collection of spaces to a viewing it as a \emph{groupoid} with
explicit \emph{paths} connecting spaces that we want to consider as
equivalent. We itemize the paths we add next:
\begin{itemize}
\item The first step is to identify each space with itself by adding trivial
  identity paths $\refl{\cubt}$ for each space $\cubt$;
\item Then we add paths $\seg{\cubt}$ that connect all occurrences of the
  same type $\cubt$ in various positions in $n$-dimensional cubes. For
  example, the 1-dimensional space corresponding to $(1-1)$ would now include
  a path connecting the two endpoints at the end of the previous section;
\item We then add paths for witnessing the usual type isomorphisms between
  finite types such as associativity and commutativity of sums and
  products. The complete list of these isomorphisms is given in the next
  section.
\item Finally, we add \emph{2-paths} between $\refl{0}$ and any path
  $\seg{\cubt}$ whose endpoints are of opposite polarities, i.e., of
  polarities $s$ and $\negp{s}$ where:
  \[\begin{array}{rcl}
  \negp{\epsilon} &=& \epsilon \\
  \negp{+s} &=& -\negp{s} \\
  \negp{-s} &=& +\negp{s}
  \end{array}\]
\item The groupoid structure forces other paths to be added as described in
  detail in the recent book on Homotopy Type Theory~\cite{hottbook}. In
  particular, for every path $p$, the groupoid structure requires an inverse
  path $p^{-1}$ in the opposite direction; and for every two paths with
  common ending and starting point, a path corresponding to the composition
  of the two paths is included and so on. These paths obey various coherence
  conditions that ensure for example that composing a path and its inverse is
  equivalent to the trivial path, and that composition is associative, up to
  path equivalences at higher levels.
\end{itemize}

\begin{verbatim}
be careful about that 2-path. Are we 
allowed to do that given that the 
empty type is, well, empty.

identical to a single point. It should 
then follow that a square where one edge 
contracts to a point would itself contract 
to a line, and a 3-dimensional cube 
in which one face contracts to a point 
would itself contract to a square, etc.

So the story is that as we are building
the denotations of types, we add these
paths making up the cubes. The only
case where we add new paths is multiplication
(see Fig.1 again). We take the paths p in S1
and the paths q in S2 and build paths (p,q)
in the product space.

Once we have this structure, we do this 
homotopy completion which turns our inductive
type into a higher-inductive type. We 
start adding 2-paths between any path 
connecting (S-S) and the refl path on 0.

Technically this is a functor between (S-S) 
and the empty type mapping S to 0, the 
other S to 0, and the path between them
to refl.

Now consider (a-b)-(c-c). We have a 
functor to (a-b)-0 that contracts the
path between c and -c. We want to 
reduce the dimension of that space
and map to (a-b).

original space

    c ------ c
    |        |  
p1  |        | p2
    |        |  
    a ------ b
       p3

after collapsing the two c's

      0
     /\
    /  \
   a -- b

or because path compose

  a ---- b
    ----

where the top path is p1;p2
and the bottom path is p3

I think we should have a 2-path
between them to say that these
paths are equivalent which will 
then collapse the square to a 
line as desired.

A critical isomorphism that is 
not reflected at all in the semantics
is that 1-1 = 0

The denotation of 1-1 is a line;
the denotation of 0 is the empty set

[ {*} | {*} ] --> []

Something like the simplification
of Conway games here perhaps?

There are two values of type
1-1

() 
and 
()^-

epsilon should take one and map
to the other because it can
never produce something of type 0

In general, we have several direction
not just left and right like in the
Int construction. In 3D we have
8 directions: +++, ++-, ..., ---
We have an epsilon that 
takes ++- and flips it to --+
and another that flips
+++ to --- and so on

So we should have 8 interpreters
for 3D one for each direction!!!

The ++- interpreter would 
manipulates values like
(v1,v2,v3) and would apply
the normal combinators to v1, v2, 
and the adjoint to v3 and so on.
\end{verbatim}

\begin{verbatim}
To complete the story we need to 
define morphisms. (More on this below.)
Once we have a notion of morphism we 
can check whether X + 0 is the same
as X etc. i.e., we can check all the 
ring equivalences.

Ok so what are the morphisms between 
these cubical objects? We know what
they are for 1-dimensional cubes: 
they are the pi combinabors. We also
know what they are for the 2-dimensional 
cubes: a maps (A-B) ==> (C-D) 
is a Pi map between A+D <=> C+B. 
How to generalize this?

Why is there no trace in the ring 
completion paper??? What are 
the morphisms in that paper?

The ring completion paper produces
a simplicial category.

p. 3 talks about the group cancellation
as subcubes along the diagonal...

We shouldn't focus on denotations. We want
an operational semantics for pi with 
negatives. The same way that Neel turned
the G construction into code that does
something neat; we want to turn that 
ring completion construction into code
that produces h.o. functions without
losing the multiplicative structure.

Show how to embed a square in 2D into
each of the faces of the 3D cube.

1D into 2D:

(a-b) => (a-b)-(0-0)
(a-b) => (a-0)-(b-0)
(a-b) => (0-b)-(0-a)
(a-b) => (0-0)-(b-a)

\end{verbatim}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{A Reversible Language with Cubical Types}

We first define values then combinators that manipulate the values to witness
the type isomorphisms.

%%%%%%%%%%%%%%%%%%
\subsection{Values}

Now that the type structure is defined, we turn our attention to the notion
of values. Intuitively, a value of the $n$-dimensional type $\tau$ is an
element of one of the sets located in one of the corners of the
$n$-dimensional cube denoted by $\tau$. Thus to specify a value, we must
first specify one of the corners of the cube (or equivalently one of the
leaves in the binary tree representation) which can easily be done using a
sequence $s$ of $+$ and $-$ polarities indicating how to navigate the cube in
each successive dimension starting from a fixed origin to reach the desired
corner. We write $v^{s}$ for the value $v$ located at corner $s$ of the cube
associated with its type. We use $\epsilon$ for the empty sequence of
polarities and identify $v$ with $v^\epsilon$. Note that the polarities
doesn't completely specify the type since different types like $(1+(1+1))$
and $((1+1)+1)$ are assigned the same denotation. What the path $s$ specifies
is the \emph{polarity} of the value, or its ``orientation'' in the space
denoted by its type. Formally:
\[\begin{array}{c}
\infer{() : 1}{} \\
\qquad
\infer[\textit{neg}]{v^{\negp{s}} : - \tau}{v^s : \tau} 
\\
\infer[\textit{left}]{(\inl{v})^{s} : \tau_1 + \tau_2}{v^{s} : \tau_1}
\qquad
\infer[\textit{right}]{(\inr{v})^{s} : \tau_1 + \tau_2}{v^{s} : \tau_2} 
\\
\infer[\textit{prod}]{(v_1,v_2)^{s_1 \cdot s_2} : \tau_1 * \tau_2}
      {v_1^{s_1} : \tau_1 & v_2^{s_2} : \tau_2} 
\end{array}\]
The rules \textit{left} and \textit{right} reflect the fact that sums do not
increase the dimension. Note that when $s$ is $\epsilon$, we get the
conventional values for the 0-dimensional sum type. The rule \textit{prod} is
the most involved one: it increases the dimension by \emph{concatenating} the
two dimensions of its arguments. For example, if we pair $v_1^{\pp}$ and
$v_2^{\mm\pp}$ we get $(v_1,v_2)^{\pp\mm\pp}$. (See Fig.~\ref{mult} for the
illustration.) Note again that if both components are 0-dimensional, the pair
remains 0-dimensional and we recover the usual rule for typing values of
product types. The rule \textit{neg} uses the function below which states
that the negation of a value $v$ is the same value $v$ located at the
``opposite'' corner of the cube.

\begin{verbatim}
We only have -
which flips ALL the directions simultaneously.
So from ++-+--+ you can only go to
        --+-++-
What if you wanted to go to
        ++++--+
I think we use the * functor to flip the 
direction we want. 
\end{verbatim}

%%%%%%%%%%%%%%%%%%
\subsection{Combinators: Example}

Consider the following simple function on 0-dimensional sum types:
\[\begin{array}{rcl}
\textsf{swapP} &\hast& (\tau_1 + \tau_2) \rightarrow (\tau_2 + \tau_1) \\
\textsf{swapP} (\inl{v}) &=& inr{v} \\
\textsf{swapP} (\inr{v}) &=& inl{v}
\end{array}\]
Given our setup, this just works $n$-dimensional types. And so do most of the
$\Pi$ combinators for that matter. Only $\eta$ and $\epsilon$ seem to need
thinking. How and why would $1-1$ which is a 1-dimensional line be the same
as the empty type which is a 0-dimensional thing. And how do we generalize
for arbitrary group identities at higher dimensions. We need a mechanism for
cubes with subspaces that ``cancel'' to map to equivalent smaller subcubes.

%%%%%%%%%%%%%%%%%%
\subsection{$\Pi$ Combinators}

The terms of $\Pi$ witness type isomorphisms of the form $b \iso b$. They
consist of base isomorphisms, as defined below, and their composition. Each
line of the above table introduces a pair of dual constants\footnote{where
  $\swapp$ and $\swapt$ are self-dual.} that witness the type isomorphism in
the middle.  These are the base (non-reducible) terms of the second,
principal level of $\Pi$. Note how the above has two readings: first as a set
of typing relations for a set of constants. Second, if these axioms are seen
as universally quantified, orientable statements, they also induce
transformations of the (traditional) values. The (categorical or homotopical)
intuition here is that these axioms have computational content because they
witness isomorphisms rather than merely stating an extensional equality.

The isomorphisms are extended to form a congruence relation by adding the
following constructors that witness equivalence and compatible closure:

\begin{table*}[t]
\[\begin{array}{cc}
\begin{array}{rrcll}
\identlp :&  0 + b & \iso & b &: \identrp \\
\swapp :&  b_1 + b_2 & \iso & b_2 + b_1 &: \swapp \\
\assoclp :&  b_1 + (b_2 + b_3) & \iso & (b_1 + b_2) + b_3 &: \assocrp \\
\identlt :&  1 * b & \iso & b &: \identrt \\
\swapt :&  b_1 * b_2 & \iso & b_2 * b_1 &: \swapt \\
\assoclt :&  b_1 * (b_2 * b_3) & \iso & (b_1 * b_2) * b_3 &: \assocrt \\
\distz :&~ 0 * b & \iso & 0 &: \factorz \\
\dist :&~ (b_1 + b_2) * b_3 & \iso & (b_1 * b_3) + (b_2 * b_3)~ &: \factor \\
\eta :&~ 0 & \iso & b + (-b)~ & : \epsilon
\end{array}
& 
\begin{minipage}{0.5\textwidth}
\begin{center} 
\Rule{}
{}
{\jdg{}{}{\idc : b \iso b}}
{}
\qquad\qquad
\Rule{}
{\jdg{}{}{c : b_1 \iso b_2}}
{\jdg{}{}{\symc{c} : b_2 \iso b_1}}
{}
\\ \bigskip
\Rule{}
{\jdg{}{}{c_1 : b_1 \iso b_2} \quad c_2 : b_2 \iso b_3}
{\jdg{}{}{c_1 \fatsemi c_2 : b_1 \iso b_3}}
{}
\\ \bigskip
\Rule{}
{\jdg{}{}{c_1 : b_1 \iso b_2} \quad c_2 : b_3 \iso b_4}
{\jdg{}{}{c_1 \oplus c_2 : b_1 + b_3 \iso b_2 + b_4}}
{}
\\ \bigskip
\Rule{}
{\jdg{}{}{c_1 : b_1 \iso b_2} \quad c_2 : b_3 \iso b_4}
{\jdg{}{}{c_1 \otimes c_2 : b_1 * b_3 \iso b_2 * b_4}}
{}
\end{center}
\end{minipage}
\end{array}\]
\caption{Combinators\label{pi-combinators}}
\end{table*}

It is important to note that ``values'' and ``isomorphisms'' are completely
separate syntactic categories which do not intermix. The semantics of the
language come when these are made to interact at the ``top level'' via
\emph{application}: 
\[\begin{array}{lrcl}
\textit{top level term}, l &::=& c~v
\end{array}\]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work and Context}

A ton of stuff here.

Connection to our work on univalence for finite types. We didn't have to rely
on sets for 0-dimensional types. We could have used groupoids again.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{abbrvnat}
\softraggedright
\bibliography{cites}

\end{document}