### 2015-03-23 Univalence/outline.tex
added new file FinSetCategory to eventually replace CauchyEquiv

\documentclass{article}
\usepackage{fullpage}
\usepackage{amssymb}
\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Basic Utilities}

\paragraph*{LeqLemmas.} A few lemmas about natural numbers.

\paragraph*{FinNatLemmas.} A few lemmas about $\texttt{Fin}~n$ which
are the numbers used to index into vectors.

\paragraph*{SubstLemmas.} A few lemmas about compositions of
propositional equalities..

\paragraph*{VectorLemmas.} A few lemmas about vectors, lookups,
mapping functions over vectors, etc.

\paragraph*{FiniteFunctions.} Proves extensionality for finite functions.

\paragraph*{Proofs.} Collects all the above and re-exports them along
with a couple of other general utilities for managing Agda proofs.

\paragraph*{DivModUtils.} External library for reasoning about
uniqueness of \texttt{divMod}. This is used extensively to mediate
between $\texttt{Fin}~(m*n)$ and
$\texttt{Fin}~m \times \texttt{Fin}~n$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Structures}

\paragraph*{SetoidUtils.} Any discrete type $A$ can be viewed as a
setoid with propositional equality $\equiv$ as the equivalence
relation.

\paragraph*{Groupoid.} A definition of 1-groupoids and operations on
them.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Equivalences}

\paragraph*{Equiv.} Defines extensional equivalence of functions
$\sim$ and shows that it is an equivalence relation. Defines
equivalence between sets $\simeq$ using two functions that go back and
forth and whose compositions are extensionally equivalent to the
identity, and shows that this equivalence is indeed an equivalence
relation. Finally shows that equivalences are injective and form a
congruence that respects $\uplus$ and $\times$.

\paragraph*{TypeEquiv.} Establishes that the Agda types $\bot$,
$\top$, $\uplus$, and $\times$ form a commutative semiring using
$\simeq$ as the underlying equivalence relation.

\paragraph*{FinEquiv.} Establishes that $\texttt{Fin}~n$ also forms a
commutative semiring with $\simeq$ as the underlying equivalence
relation. In particular, we have:
\[\begin{array}{rcll}
\texttt{Fin}~0 &\simeq& \bot \\
\texttt{Fin}~1 &\simeq& \top \\
\texttt{Fin}~(m+n) &\simeq& \texttt{Fin}~m \uplus \texttt{Fin}~n \\
\texttt{Fin}~(m*n) &\simeq& \texttt{Fin}~m \times \texttt{Fin}~n
\end{array}\]
and then we have all the commutative semiring axioms, e.g.,
$\texttt{Fin}~(0+m) \simeq \texttt{Fin}~m$. The actual proof says that
0, 1, $+$, and $*$ for a commutative semiring structure under the
equivalence that equates $m$ and $n$ if
$\texttt{Fin}~m \simeq \texttt{Fin}~n$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Bimonoidal Categories}

\paragraph*{FinSetCategory.} Defines the category of finite sets and
all functions between them. As stated on the nLab page, ``it is easy
(and thus common) to make \texttt{FinSet} skeletal; there is one
object for each natural number $n$ (including $n=0$), and a morphism
from $m$ to $n$ is an $m$-tuple $(f_0,\ldots,f_{m−1})$ of numbers
satisfying $0 \leq f_i < n$. This amounts to identifying $n$ with the
set $\{0,\ldots,n−1\}$.'' This structure will be a building block for
permutations (defined in \texttt{ConcretePermutation}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{To do}

PiLevel0
ConcretePermutation
VecOps
PiPerm

Enumeration ??
SEequivSCPermEquiv ??

\paragraph*{Cauchy representation.} A permutation on $n$ elements is
represented by a vector \texttt{v : Vec (Fin n) n}. If the $n$
elements are indexed by positions, the element at position $i$ is
mapped to position \texttt{v !! i} by the permutation. There is always
a trivial permutation called \texttt{1C} that maps each position to
itself. The Cauchy representation does not enforce that the vector
entries are disjoint. This is enforced by the definition of ``concrete
permutations'' below.

\paragraph*{Concrete permutation.} A concrete permutation consists of
two Cauchy vectors and two proofs that their compositions is the
identity permutation \texttt{1C}. Concrete permutations are an
equivalence relation. Concrete permutations actually have more
structure: a sum, a unit for the sum, etc. We can also build setoids
whose carriers are concrete permutations under the standard $\equiv$
propositional equality.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}