### 2015-09-19 Univalence/response.txt
expand response

A general theme of the reviews appears to request a better presentation of the
theorems and how they all relate to each other. A compressed and intuitive
presentation would be as follows. First we have three independent
constructions:

  A : Set    <------- HoTT-style equivalences -------> B : Set

  A : Fin m  <------------ Permutations -------------> B : Fin n

  A : Pi     <------------ Pi-combinators -----------> B : Pi

Each construction relates a particular representation of types via appropriate
relations; each construction has a rig structure (additive and multiplicative
units, commutative and associative sums and products, and distributivity of
products over sums). We then have a second level:

  e : HoTT-Equivalence <--------- Equiv. of equivalences --------> f : HoTT-Equivalence

  p : Permutations     <--------- Equiv. of permutations --------> q : Permutations

  c : Pi-combinators   <-------------- Pi-optimizations ---------> d : Pi-combinators

The two levels of constructions are connected, and form categories. For
example, we have a category whose objects are Pi types, whose morphisms are
Pi-combinators, and whose equality of morphisms is Pi-optimizations; this
category is a (free) symmetric rig groupoid. We finally have theorems that
connect some of the constructions vertically allowing us to move across
representations.

These are the "raw" results without interpretation, context, or applications.
Perhaps unfortunately, the current presentation of the paper mixes these "raw"
results with theoretical connections to HoTT, with Agda formalization concerns,
and with practical connections to reversible circuits and their optimizations.
We acknowledge that these connections can be perceived as distracting, and that
their presentation could be much improved (which we are working on), BUT these
connections are all critical and essential -- and are why this is being sent to
POPL rather than (say) Theory and Applications of Category Theory.

Indeed some of the connections above are trivial in "conventional mathematics";
however some of those same connections hide implicit isomorphisms (changes of
representations) that are important computationally.  Our construction shows,
for example, that transporting properties across equivalences of finite types
can be done via executing permutations, something which has a clear
computational cost and which itself depends on the choice of representations of
the permutations. Since this is POPL, the presentation emphasizes turning these
constructions into actual executable programming constructs of a programming
language, complete with executable optimizer, and with an eye towards
higher-order extensions. To our knowledge, there are no programming languages
for representing --and reasoning about-- a rich notion of "permutations"
(including sums and products).  Furthermore, the higher-order extensions are
crucial for abstractions over such permutations by allowing the permutations
themselves to be treated as data that can be constructed and examined
dynamically.

Specific (and short with apologies) responses:

* I think our excitement with all the connections apparently left a "hint of
  arrogance." We sincerely apologize for that, and will fix.

* The supplementary material gives one small example of rewriting circuits
  using the 2-axioms. We do not have larger experiments at this point but we
  have done enough to notice that the rewriting of one circuit to another does
  not necessarily take an exponential number of steps: the diagram on p.10
  shows an equivalence that is independent of the sizes of c1 and c2 and the
  sizes of A, B, C, and D.

* Our axioms vs. Laplaza diagrams: Yes, the level 2 combinators for 
  distributivity properties and the interaction of combinators involving 0
  are the same.  We apparently were not clear enough about what changed:

  The general process of incorporating these diagrams as combinators is to give
  judicious names to each natural transformation of each diagram representing
  a natural isomorphism, and then proceed to prove equivalence of the two
  expressions (for all 3 constructions). In some cases this was achieved by
  adding a 2-axiom that directly relates the terms but in others it was possible
  to simplify the two sides using other 2-axioms. The process led us to refine
  the original 1-axioms of Pi which implicitly collapsed certain terms that
  should only been related by explicit non-trivial paths. Thus the most 
  apparent changes were to the programming language Pi itself, and not the
  combinators derived from Laplaza's diagrams.  In other words, we let the
  ``proper'' categorical semantics guide us to define the syntax of a new
  version of Pi which had better meta-properties.

  So it is not the coherences which are more "small step", it is the 
  underlying programming language.

* Proofs: All the theorems (except Thm 2) are in our Agda formalization. This
  code was not presentable as an artifact at the time of submission and
  required some serious refactoring which is under progress. Thm 2 is indeed
  wrong and the informal statement was included by mistake; the core of what
  we meant to convey is in the (correct) Thm 9.

* The 3rd axiom on p9 says c;(!c)=id which proves that swap;swap=id. (The
  bottom of p4 notes that every combinator has a dual according to figure 1
  with swap being its own inverse.) The existence of inverses for 1-combinators
  is a meta-property (not a constructor of the inductive datatype of
  combinators but a function mapping combinators to combinators). The same
  applies to inverses for 2-combinators.

* Completeness in Thm 11. *** BRIEFLY EXPLAIN HOW WE GET COMPLETENESS FROM LAPLAZA ***

* lifting of the 0d combinators: The problem is when given a : T1 <=> T2,
  b : T3 <=> T4, to construct something of type T1 x T3 <=> T2 x T4.  If
  b is the identity, yes it works.  But it does not work in general.
  There is even an (easy!) general definition for what it ought to be.
  The point is that that definition is not functorial.  If it were, that
  would contradiction Thomasson's result.

* We would like to remark that Thomasson believed strongly enough that a 
  particular construction (which had been published several times) was in 
  fact wrong, that he published a paper point exactly that, and only that.
  Most PL theorists would not understand that this result actually applies to
  constructions they are familiar with: the Int-construction, which applies so
  nicely to language built atop linear logic.  We believe, just as strongly 
  as Thomasson, that pointing out that the ``obvious'' generalization to
  the setting with two monoidal constructions (such as ours) does not work,
  for the same reasons.  Yes, we do realize that including a negative result
  in paper makes it harder to publish; we nevertheless feel that we can save
  others a lot of fruitless pain by attempting to do so.

* Referee 4 thinks that "The main result of the paper is that the category of
  finite sets and bijections, with the two monoidal structures given by sum and
  product, is the free rig groupoid." And then further that " A serious
  weakness is that the main result has no proof, nor even a proof sketch."  This
  would be because we did not actually claim this result as a theorem!  [We
  believe it to be true, but as of this writing, that proof is still 
  incomplete.]
  Our main result is that weak rig categories are exactly the setting for the
  semantics for a programming language of reversible circuits.  Furthermore, the
  coherence laws of such a category exactly give complete rules for deciding
  equivalence of such circuits.  In other words, theorems 10 and 11.  We also
  consider theorem 9 (that *all* types, not just finite ones, equivalences and
  equivalences-of-equivalences), form a weak rig groupoid to be quite important
  for PL theory.