### 2019-04-04 fractionals/p.tex
thoughts on fractionals based on discussions with Chao-Hong

Reference cells have fractional types

1 -> A * 1/A
From no information, create a ref cell with a getter (A) and a setter (1/A)
Usually type of getter is 1 -> A and setter A -> 1
Using that A -> B is [not A + B] and hence [1/A * B] in this case then:
1 -> A is [1/1 * A] which is A, and
A -> 1 is [1/A * 1] which is 1/A

Inspiration from lenses
Particle/Anti-particle
Entanglement (side channel between setter and getter)

Forget 1/0 for now; we will fix later with meadow trick

Also we will index the types with a monad or something to make the effect of creating the cell explicit

Denotational semantics already kind of known: groupoid with fractional cardinality

If A and 1/A are left loose, we get errors (reading uninitialized cells, race conditions, etc.). But if information is conserved and we do meadow trick, no errors! That should be the main theorem.

Can get functions A -> B as [1/A * B] , which has a hidden cell for parameter passing, a setter to pass the parameter and a value of type B that is blocked until the cell is full.

Now do interesting programs. Start with all the constructions for closed monoidal categories.

Check 1/(1/A) = A
1 -> 1/A * 1/(1/A)
Se here we are creating a cell that holds the setter for another cell; ho-functions!!!