### 2019-11-22 fracGC/trace1.tex

\resizebox{\columnwidth}{!}{%
\begin{tikzpicture}
\node[draw,align=left] at (0,0) {%
⟪ dist ⊚ (id⟷ ⊕ (id⟷ ⊗ swap₊)) ⊚ factor ∥ [{\color{gray}(𝔽 , 𝔽) , (𝔽 , 𝕋) , (𝕋 , 𝔽) ,} (𝕋 , 𝕋)] [ 3 ]⟫\\\\
⟪ id⟷ ⊚ (id⟷ ⊕ (id⟷ ⊗ swap₊)) ⊚ factor ∥ [{\color{gray}inj₁ (tt , 𝔽) , inj₁ (tt , 𝕋) , inj₂ (tt , 𝔽) ,} inj₂ (tt , 𝕋)] [ 3 ]⟫\\\\
⟪ (id⟷ ⊕ (id⟷ ⊗ swap₊)) ⊚ factor ∥ [{\color{gray}inj₁ (tt , 𝔽) , inj₁ (tt , 𝕋) , inj₂ (tt , 𝔽) ,} inj₂ (tt , 𝕋)] [ 3 ]⟫\\\\
⟪ (id⟷ ⊕ (id⟷ ⊗ id⟷)) ⊚ factor ∥ [{\color{gray}inj₁ (tt , 𝔽) , inj₁ (tt , 𝕋) ,} inj₂ (tt , 𝔽) {\color{gray}, inj₂ (tt , 𝕋)}] [ 2 ]⟫\\\\
⟪ (id⟷ ⊕ id⟷) ⊚ factor ∥ [{\color{gray}inj₁ (tt , 𝔽) , inj₁ (tt , 𝕋) ,} inj₂ (tt , 𝔽) {\color{gray}, inj₂ (tt , 𝕋)}] [ 2 ]⟫\\\\
⟪ id⟷ ⊚ factor ∥ [{\color{gray} inj₁ (tt , 𝔽) , inj₁ (tt , 𝕋) ,} inj₂ (tt , 𝔽) {\color{gray}, inj₂ (tt , 𝕋)}] [ 2 ]⟫\\\\
⟪ factor ∥ [{\color{gray} inj₁ (tt , 𝔽) , inj₁ (tt , 𝕋) ,} inj₂ (tt , 𝔽) {\color{gray} , inj₂ (tt , 𝕋)}] [ 2 ]⟫\\\\
⟪ id⟷ ∥ [{\color{gray} (𝔽 , 𝔽) , (𝔽 , 𝕋) ,} (𝕋 , 𝔽) {\color{gray} , (𝕋 , 𝕋)}] [ 2 ]⟫
};%
\end{tikzpicture}
}%