### 2019-11-22 fracGC/trace2.tex
trace 2 & 3

\resizebox{\columnwidth}{!}{%
\begin{tikzpicture}
\node[draw,align=left] at (0,0) {%
⟪ uniti⋆r ⊚ (id↔ ⊗ η 𝔽) ⊚…∥ [{\color{gray}𝔽 ,} 𝕋] [ 1 ]⟫\\\\

⟪ (id↔ ⊗ η 𝔽) ⊚ … ∥ [{\color{gray}(𝔽 , tt) ,} (𝕋 , tt)] [ 1 ]⟫\\
  ⋮\\
⟪ (id↔ ⊗ id↔) ⊚ assocl⋆ ⊚ … ∥ [{\color{gray}(𝔽 , 𝔽 , ○) , (𝔽 , 𝕋 , ○) ,} (𝕋 , 𝔽 , ○) {\color{gray}, (𝕋 , 𝕋 , ○)}] [ 2 ]⟫\\
  ⋮\\
⟪ assocl⋆ ⊚ … ∥ [{\color{gray}(𝔽 , 𝔽 , ○) , (𝔽 , 𝕋 , ○) ,} (𝕋 , 𝔽 , ○) {\color{gray}, (𝕋 , 𝕋 , ○)}] [ 2 ]⟫\\
  ⋮\\
⟪ (((dist ⊚ …) ⊗ id↔) ⊚ …∥ [{\color{gray} ((𝔽 , 𝔽) , ○) , ((𝔽 , 𝕋) , ○) ,} ((𝕋 , 𝔽) , ○) {\color{gray}, ((𝕋 , 𝕋) , ○)}] [ 2 ]⟫\\
  ⋮\\
⟪((((id↔ ⊕ (id↔ ⊗ swap₊)) ⊚ …) ⊗ id↔) ⊚ …∥ [{\color{gray} (inj₁ (tt , 𝔽) , ○) , (inj₁ (tt , 𝕋) , ○) ,} (inj₂ (tt , 𝔽) , ○) {\color{gray}, (inj₂ (tt , 𝕋) , ○)}] [ 2 ]⟫\\
  ⋮\\
⟪((factor ⊚…) ⊗ id↔)⊚…∥ [{\color{gray}(inj₁ (tt , 𝔽) , ○) , (inj₁ (tt , 𝕋) , ○) , (inj₂ (tt , 𝔽) , ○) ,} (inj₂ (tt , 𝕋) , ○)] [ 3 ]⟫\\
  ⋮\\
⟪(((distl ⊚ …) ⊗ id↔) ⊚…∥ [{\color{gray} ((𝔽 , 𝔽) , ○) , ((𝔽 , 𝕋) , ○) , ((𝕋 , 𝔽) , ○) ,} ((𝕋 , 𝕋) , ○)] [ 3 ]⟫\\
  ⋮\\
⟪((((id↔ ⊕ (swap₊ ⊗ id↔)) ⊚ …) ⊗ id↔) ⊚ … ∥ [{\color{gray} (inj₁ (𝔽 , tt) , ○) , (inj₁ (𝕋 , tt) , ○) , (inj₂ (𝔽 , tt) , ○) ,} (inj₂ (𝕋 , tt) , ○)] [ 3 ]⟫\\
  ⋮\\
⟪ ((factorl ⊚ swap⋆) ⊗ id↔) ⊚ … ∥ [{\color{gray}(inj₁ (𝔽 , tt) , ○) , (inj₁ (𝕋 , tt) , ○) ,} (inj₂ (𝔽 , tt) , ○) , {\color{gray}(inj₂ (𝕋 , tt) , ○)}] [ 2 ]⟫\\
  ⋮\\
⟪ (swap⋆ ⊗ id↔) ⊚ assocr⋆ ⊚ … ∥ [{\color{gray}((𝔽 , 𝔽) , ○) ,} ((𝔽 , 𝕋) , ○) {\color{gray}, ((𝕋 , 𝔽) , ○) , ((𝕋 , 𝕋) , ○)}] [ 1 ]⟫\\
  ⋮\\
⟪ assocr⋆ ⊚ (id↔ ⊗ ε 𝔽) ⊚ unite⋆r ∥ [{\color{gray}(𝔽 , 𝔽) , ○) , ((𝔽 , 𝕋) , ○) ,} ((𝕋 , 𝔽) , ○) {\color{gray}, ((𝕋 , 𝕋) , ○)}] [ 2 ]⟫\\
  ⋮\\
⟪ (id↔ ⊗ ε 𝔽) ⊚ unite⋆r ∥ [{\color{gray}(𝔽 , 𝔽 , ○) , (𝔽 , 𝕋 , ○) ,} (𝕋 , 𝔽 , ○) {\color{gray}, (𝕋 , 𝕋 , ○)}] [ 2 ]⟫\\
  ⋮\\
⟪ unite⋆r ∥ [{\color{gray}(𝔽 , tt) ,} (𝕋 , tt)] [ 1 ]⟫ \\\\
⟪ id↔ ∥ [{\color{gray}𝔽 ,} 𝕋] [ 1 ]⟫
};%
\end{tikzpicture}
}%